### 2016-12-11 doc/arxivize.tex
added doc

\pdfoutput=1

\documentclass[letterpaper]{article}
\usepackage{hyperref}
\hypersetup{
  pdfinfo={
    Title={Entropy-SGD: Biasing Gradient Descent Into Wide Valleysre},
    Author={Pratik Chaudhari, Anna Choromanska, Stefano Soatto, Yann LeCun},
    Keywords={deep learning, gradient descent, optimization, machine learning, neural networks}
  }
}
\usepackage{pdfpages}
\usepackage{grffile}

\begin{document}
\includepdf[pages=1-last]{chaudhari.choromanska.ea.iclr17.pdf}
\end{document}

### 2016-12-11 doc/chaudhari.choromanska.ea.iclr17.tex
added doc

\pdfoutput = 1

\documentclass[10pt]{article}
\input{macros}
\usepackage{iclr2017_conference}
\usepackage{mathptmx}

\linespread{1.02}

\pagestyle{empty}

% save space
\setlength{\floatsep}{0.0in}
\setlength{\textfloatsep}{0.0in}
\setlength{\intextsep}{0.05in}
\setlength{\belowcaptionskip}{0.05in}
\setlength{\abovecaptionskip}{0.05in}
\setlength{\abovedisplayskip}{0.05in}
\setlength{\belowdisplayskip}{0.05in}

% notation goes here
\newcommand{\entropysgd}{\mathrm{Entropy}\textrm{-}\mathrm{SGD}}
\newcommand{\minibatch}[1]{\Xi^{#1}}
\newcommand{\mnistfc}{\textrm{mnistfc}}
\newcommand{\smallmnistfc}{\textrm{small-mnistfc}}
\newcommand{\charlstm}{\textrm{char-lstm}}
\newcommand{\lenet}{\textrm{LeNet}}
\newcommand{\allcnn}{\textrm{All-CNN-BN}}

\title{Entropy-SGD: Biasing Gradient Descent Into Wide Valleys}

\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$\\[0.05in]
$^{1}$ Computer Science, University of California, Los Angeles\\
$^{2}$ Courant Institute of Mathematical Sciences, New York University\\
$^{3}$ Facebook AI Research\\[0.05in]
{\footnotesize
Email:\ \href{mailto:pratikac@ucla.edu}{pratikac@ucla.edu},
\href{mailto:achoroma@cims.nyu.edu}{achoroma@cims.nyu.edu},
\href{mailto:soatto@cs.ucla.edu}{soatto@cs.ucla.edu},
\href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu}}
}

% macros to color text and insert margin comments
\setlength{\marginparwidth}{1in}
\newcommand{\pc}[2]{{\color{ForestGreen} #1}\marginpar{\tiny\noindent{\raggedright{\color{Sienna}[PC]}\color{Sienna}{ #2} \par}}}
\renewcommand{\ss}[2]{{\color{cyan} #1}\marginpar{\tiny\noindent{\raggedright{\color{BurntOrange}[SS]}\color{BurntOrange}{ #2} \par}}}
\newcommand{\ac}[2]{{\color{magenta} #1}\marginpar{\tiny\noindent{\raggedright{\color{magenta}[AC]}\color{magenta}{ #2} \par}}}
\newcommand{\todo}[1]{{\color{gray}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[TODO]}}}}
\newcommand{\ignore}[1]{}

% uncomment these for submission!
% \renewcommand{\pc}[2]{#1}
% \renewcommand{\ss}[2]{#1}
% \renewcommand{\ac}[2]{#1}
% \renewcommand{\todo}[1]{#1}

\graphicspath{{./fig/}}

%\iclrfinalcopy % Uncomment for camera-ready version
\begin{document}

\maketitle

\begin{abstract}
This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape at solutions found by gradient descent. Local extrema with low generalization error have a large proportion of almost-zero eigenvalues in the Hessian with very few positive or negative eigenvalues. We leverage upon this observation to construct a local entropy based objective that favors well-generalizable solutions lying in the flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
%
Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of local entropy at each update of the weights.
%
We prove that incorporating local entropy into the objective function results in a smoother energy landscape and use uniform stability to show improved generalization bounds over SGD.
%
Our experiments on competitive baselines demonstrate that $\entropysgd$ leads to improved generalization and has the potential to accelerate training.
\end{abstract}

\section{Introduction}
\label{s:intro}

This paper focuses on developing new optimization tools for deep learning that are tailored to exploit the local geometric properties of the objective function. Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the eigenspectrum of the Hessian at a local minimum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}). It is evident that:
\begin{enumerate}[(i)]
\item a large number of directions ($\approx 94\%$) have near-zero eigenvalues (magnitude less than $10^{-4}$),
\item positive eigenvalues (right inset) have a long tail with the largest one being almost $40$,
\item negative eigenvalues (left inset), which are directions of descent that the optimizer missed, have a much faster decay (the largest negative eigenvalue is only $-0.4$).
\end{enumerate}
Interestingly, this trend is not unique to this particular network. Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (refer to Sec.~\ref{s:expt} for more experiments).  Local minima that generalize well and are discovered by gradient descent lie in "wide valleys" of the energy landscape, that look more like the bottom of a bucket than sharp isolated minima like the holes on a golf-course.
%For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization. This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.

\begin{figure}[!tbh]
\vspace*{-0.1in}
\centering
\includegraphics[width=\textwidth]{lenet_hessian.pdf}
\caption{\small Eigenspectrum of the Hessian at a local minimum of a CNN on MNIST (two independent runs). \textbf{Remark:} The central plot shows the eigenvalues in a small neighborhood of zero whereas the left and right insets show the entire tails of the eigenspectrum.}
\label{fig:lenet_hessian}
\end{figure}

Based on this understanding of how the local geometry looks at the end of optimization, can we modify SGD to actively seek such regions? We show that flat regions in the energy landscape correspond to regions where the log-partition function is small, if computed locally on the Gibbs distribution corresponding to the optimization problem. Roughly, for parameters denoted by $x \in \reals^n$, we modify a loss function $f(x)$ to be
$$
	f(x) - \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx';
$$
where the second term is a log-partition function which is seen as a ``local free energy'' and measures both the depth of a valley at a location $x$ and its flatness through the entropy of $f(x')$. The $\entropysgd$ algorithm presented in this paper employs a few iterations of stochastic gradient Langevin dynamics (SGLD) to approximate the second term before each update of the parameters using the back-propagated gradient. At a conceptual level, this algorithm resembles running two nested loops of SGD. We show that the above modified loss function results in a smoother energy landscape defined by the hyper-parameter $\g$ which we can think of as a ``scope'' that seeks out valleys of specific widths. Actively biasing towards wide valleys aids generalization, in fact, we can optimize solely the free energy term to obtain similar generalization error as SGD on the original loss function. Our experiments in Sec.~\ref{s:expt} on MNIST and CIFAR-10~\citep{krizhevsky2009learning} show that $\entropysgd$ scales to deep networks used in practice and in terms of the number of updates to the parameters, trains much quicker than SGD.

\section{Related work}
\label{s:prior_work}

Our above observation about the eigenspectrum of Hessian (further discussed in Sec.~\ref{s:expt}) is similar to results on a perceptron model in~\citet{dauphin2014identifying} where the authors connect the loss function of a deep network to a high-dimensional Gaussian random field. They also relate to earlier studies such as~\citet{Baldi:1989:NNP:70359.70362,Fyodorov2007,Bray2007} which show that critical points with high training error are exponentially likely to be saddle points with many negative directions and all local minima are likely to have error that is very close to that of the global minimum. The authors also argue that convergence of gradient descent is affected by the proliferation of saddle points surrounded by high error plateaus --- as opposed to multiple local minima. One can also see this via an application of Kramer's law: the time spent by diffusion is inversely proportional to the smallest negative eigenvalue of the Hessian at a saddle point~\citep{bovier2006metastability}.

The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems. Although assumptions in these works are somewhat unrealistic in the context of deep networks used in practice, similar results are also true for linear networks which afford a more thorough analytical treatment~\citep{DBLP:journals/corr/SaxeMG13}. For instance,~\citet{soudry2016no} show that with mild over-parameterization and dropout-like noise, training error for a neural network with one hidden layer and piece-wise linear activation is zero at every local minimum. All these results suggest that the energy landscape of deep neural networks should be easy to optimize and they more or less hold in practice --- it is easy to optimize a prototypical deep network to near-zero loss \emph{on the training set}~\citep{hardt2015train,DBLP:journals/corr/GoodfellowV14}.
%Works such as which shows that for deep linear networks with more than three layers, there exist ``bad'' saddle points, i.e., where the Hessian has no negative eigenvalues; SGD will find it difficult to make progress in such regions.

Obtaining good \emph{generalization} error, however, is challenging: complex architectures are sensitive to initial conditions and learning rates~\citep{sutskever2013importance} and even linear networks~\citep{kawaguchi2016deep} may have degenerate and hard to escape saddle points~\citep{ge2015escaping,anandkumar2016efficient}. Techniques such as adaptive~\citep{duchi2011adaptive} and annealed learning rates, momentum~\citep{tieleman2012lecture}, as well as architectural modifications like dropout~\citep{srivastava2014dropout}, batch-normalization~\citep{ioffe2015batch,cooijmans2016recurrent}, weight scaling~\citep{salimans2016weight} etc. are different ways of tackling this issue by making the underlying landscape more amenable to first-order algorithms. However, the training process often requires a combination of such techniques and it is unclear beforehand to what extent each one of them helps.

Such a paradox between a seemingly benign energy landscape and delicate training algorithms is also seen in problems such as random $k$-SAT~\citep{achlioptas2006solution} and compressed sensing~\citep{krzakala2012statistical}. Introducing non-local terms in classical inference techniques for these problems has led to state-of-the-art algorithms~\citep{braunstein2005survey,maneva2007new}. Closer to deep learning, although motivated from similar ideas, the authors in~\citet{baldassi2015subdominant,baldassi2016unreasonable} show that, both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average. Using a novel out-of-equilibrium analysis, the authors show that such an ensemble recovers dense clusters in the solution space. Such dense clusters generalize better than isolated local minima that may lie close to the global minimum. A related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and encourages individual models to explore by modulating their distance to the ensemble average; this is a special case of robust ensembles.

An ensemble training procedure thus provides improved generalization but it incurs a large complexity in terms of hardware and communication across workers. In this paper, we construct an optimization algorithm for a general deep network that can exploit the same phenomenon of out-of-equilibrium dense clusters but does not require distributed training. The enabling technique in our case is that we approximate the local entropy using Langevin dynamics. \citet{baldassi2016unreasonable} instead employ belief propagation to do so; this can only be done for simpler perceptron models considered there.

The authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization. However, as they mention, the effect of hyper-parameters for these constraints is intricately tied together and they are difficult to choose even for small problems. Their formulation depends on Hessian-vector products. It is unclear how this method could be scaled to modern deep networks. Motivated by the same final goal, viz., flat local minima, our approach eliminates these extraneous and difficult to control hyper-parameters and instead results in a scalable algorithm. Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper. The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.

In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it resembles continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems. These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{,DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions. Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks. %\pc{Let us also note that motivated from continuation,~\citet{gulcehre2016mollifying} insert annealed additive noise in the activation function to approximately convolve the original objective function.}{This paper is not related to our work, it is related to continuation. I don't see any reason to add this sentence}

\section{Local entropy}
\label{s:local_entropy}

We first provide a simple intuition for the concept of local entropy of an energy landscape. Consider a cartoon energy landscape in Fig.~\ref{fig:scoping} where the x-axis denotes the configuration space of the parameters. We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\textrm{non-robust}}$. Under a Bayesian prior on the parameters, say a Gaussian of a fixed variance at locations $x_{\textrm{robust}}$ and $x_{\textrm{non-robust}}$ respectively, the wider local minimum has a higher log-likelihood than the sharp valley on the right. In discrete optimization problems, such wide local minima are dense clusters of solutions in the parameter space where nearby parameters have similar energy, say, the number of clauses violated in $k$-SAT~\citep{sipser2006introduction,achlioptas2006solution}. On the other hand, sharp local minima are isolated solutions that are far away from other solutions.

\begin{wrapfigure}{r}{0.4\textwidth}
\centering
\includegraphics[width=0.4 \textwidth]{scoping.pdf}
\caption{\small Local entropy concentrates on dense clusters in the energy landscape.}
\label{fig:scoping}
\vspace{0.2in}
\end{wrapfigure}

The above discussion suggests that parameters that lie in wider local minima like $x_{\trm{robust}}$, which may possibly have a higher loss than the global minimum, should generalize better than the ones that are simply at the global minimum. This motivates us to use a very well-known quantity from thermodynamics known as ``free energy'' (also sometimes called ``free entropy''):
$$
    \textrm{free energy} = \textrm{internal energy} - \textrm{temperature} \times \textrm{entropy}.
$$
Roughly, internal energy is the average energy of all configurations while entropy counts the number of distinct configurations in a physical system at a given temperature. In a neighborhood of $x_{\textrm{robost}}$, the entropy term term is small while internal energy is large. Conversely, near $x_{\textrm{non-robust}}$, internal energy is low while entropy is high. The ``negative free entropy'' computed locally near a particular configuration thus provides a way of picking large flat regions of landscape over sharp, narrow valleys in spite of the latter possibly having a lower loss. Quite conveniently, free energy is also the log-partition function of a system.

Formally, for a parameter vector $x \in \reals^n$, consider a Gibbs distribution corresponding to a given energy landscape $f(x)$:
\beq{
    \P(x;\ \b) = Z_\b^{-1}\ \exp\ \rbrac{-\b\ f(x)};
    \label{eq:gibbs_original}
}
where $\b$ is known as the inverse temperature and $Z_\b$ is a normalizing constant. As $\b \to \infty$, the probability distribution above concentrates on the global minimum of $f(x)$ given as
\beq{
	x^* = \argmin_x\ f(x),
    \label{eq:argmin_generic}
}
which establishes the link between the Gibbs distribution and a generic optimization problem~\eqref{eq:argmin_generic}. We would instead like the probability distribution --- and therefore the underlying optimization problem --- to focus on flat regions such as $x_{\textrm{robust}}$. With this in mind, let us construct a modified Gibbs distribution:
\beq{
    \P(x';\ x, \b, \g) = Z_{x,\b,\ \g}^{-1}\ \exp\ \rbrac{-\b\ f(x') - \b\ \f{\g}{2}\ \norm{x-x'}^2_2}.
    \label{eq:gibbs_modified}
}
The distribution in~\eqref{eq:gibbs_modified} is a function of a dummy variable $x'$ and is parameterized by the original location $x$. The parameter $\g$ biases the modified distribution~\eqref{eq:gibbs_modified} towards $x$; a large $\g$ results in a $P(x';\ x,\b,\g)$ with all its mass near $x$ irrespective of the energy landscape of $f(x')$. For small values of $\g$, the term $f(x')$ in the exponent dominates and the modified distribution is similar to the original Gibbs distribution in~\eqref{eq:gibbs_original}. We will set the inverse temperature $\b$ to $1$ because $\g$ affords us similar control on the Gibbs distribution.

\begin{definition}[Local entropy]
\label{def:local_entropy}
The local free energy of the Gibbs distribution in~\eqref{eq:gibbs_original}, colloquially called ``local entropy'' in the sequel and denoted by $F(x, \g)$, is defined as the log-partition function of modified Gibbs distribution~\eqref{eq:gibbs_modified}, i.e.,
\aeq{
    F(x, \g) &= \log Z_{x, 1, \g} \notag\\
    &= \log \int_{x'}\ \exp \Big(-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2 \Big)\ dx'.
    \label{eq:local_entropy}
}
\end{definition}
The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth. Fig.~\ref{fig:scoping} shows the negative local entropy $-F(x, \g)$ for the original energy landscape. Note that $-F(x,\g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.

Let us also note that the quantity we have defined as local entropy in Def.~\ref{def:local_entropy} is different from classical entropy which counts the number of likely configurations under a given distribution. This is given by
$
    - \int_{x'}\ \log \P(x';\ x, \g)\ d \P(x';\ x, \g)
$
for the Gibbs distribution in~\eqref{eq:gibbs_modified}. Minimizing classical entropy however does not differentiate between flat regions that have very high loss versus dense regions that lie deeper in the energy landscape. For instance, in Fig.~\ref{fig:scoping}, classical entropy is smallest in the neighborhood of $x_{\textrm{candidate}}$ which is a large region with very high loss on the training dataset and is unlikely to generalize well.
\ignore{We could minimize a modified loss function of the form $f(x) + \lambda S(x, \g)$ whose gradient can be computed to be
$$
    \nabla \Big(f(x) + \lambda S(x, \g)\Big) = \nabla f(x) - \lambda \g\ \corr (g(x'),\ x-x');
$$
where we have defined the cross-correlation as
$$
    \corr (g(x'),\ x-x') := \ag{g(x')\ (x-x')} - \ag{g(x')}\ \ag{x-x'};
$$
with $g(x') = f(x') + \f{\g}{2}\ \norm{x-x'}^2_2$. The gradient of the entropy can again be estimated using Langevin dynamics. Practically, one then has to modulate the hyper-parameter $\lambda$ during the course of the training to first use the gradient $\nabla f(x)$ to make progress and then turn on the entropy term to scope dense clusters.}

\section{Entropy-guided SGD}
\label{s:entropysgd}

We now present the $\entropysgd$ algorithm, a variant of SGD that is motivated from local entropy. Simply speaking, we modify the loss function of a deep network to include the negative local entropy from Sec.~\ref{s:local_entropy}. The gradient of the modified loss can then be computed using Langevin dynamics. The $\entropysgd$ algorithm (cf. Alg.~\ref{alg:entropysgd}) thus has a strong flavor of ``SGD-inside-SGD'': the outer SGD updates the parameters, while an inner SGD estimates the gradient of local entropy at each iteration.

Consider a typical classification setting, let $x \in \reals^n$ be the weights of a deep neural network and $\xi_k \in \Xi$ be samples from a dataset $\Xi$ of size $N$. Let $f(x; \xi_k)$ be the loss function, e.g., cross-entropy of the classifier on a sample $\xi_k$. The original optimization problem is:
\beq{
    x^* = \argmin_x\ \f{1}{N}\ \sum_{k=1}^N\ f(x;\ \xi_k);
    \label{eq:dnn_objective}
}
where the objective $f(x, \xi_k)$ is typically, a non-convex function in both the weights $x$ and the sample $\xi_k$. The $\entropysgd$ algorithm instead solves the problem
\beq{
    x^*_{\entropysgd} = \argmin_x\ \r\ \rbrac{\f{1}{N}\ \sum_{k=1}^N\ f(x;\ \xi_k)} - F(x, \gamma;\ \Xi);
    \label{eq:entropysgd_objective}
}
where we have made the dependence of local entropy $F(x,\g)$ on the dataset $\Xi$ explicit. The objective in~\eqref{eq:entropysgd_objective} thus scales the original loss by a hyper-parameter $\rho$ and adds the negative local entropy term $F(x, \g, \Xi)$ introduced in Sec.~\ref{s:local_entropy}. The parameter $\rho$ allows us to control the contribution of the original back-propagated gradient in $\entropysgd$. Sec.~\ref{ss:intuition_local_entropy} and Sec.~\ref{ss:theoretical_properties} below discuss the effect of $\rho$ and the scoping parameter $\g$.

\subsection{Computing the gradient of local entropy}
\label{ss:grad_local_entropy}

The gradient of local entropy over a randomly sampled mini-batch of $m$ samples denoted by $\xi_{\ell_i} \in \minibatch{\ell}$ for $i \leq m$ is easy to derive and is given by
\beq{
    \nabla_x F\rbrac{x, \g;\ \minibatch{\ell}} = -\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}};
    \label{eq:local_entropy_grad}
}
where the notation $\ag{\cdot}$ denotes an expectation of its arguments (we have again made the dependence on the data explicit) over a Gibbs distribution of the original optimization problem modified to focus on the neighborhood of $x$; this is given by
\beq{
    \P(x';\ x, \g) \propto\ \exp \sqbrac{-\rbrac{\f{1}{m}\ \sum_{i=1}^m\ f\rbrac{x;\ \xi_{\ell_i}}} - \f{\g}{2}\ \norm{x-x'}_2^2}.
    \label{eq:entropysgd_modified_gibbs}
}
Computationally, the gradient in~\eqref{eq:local_entropy_grad} involves estimating $\ag{x';\ \minibatch{\ell}}$ with the current weights fixed to $x$. This is an expectation over a Gibbs distribution and is hard to compute. We can however approximate it using Markov chain Monte-Carlo (MCMC) techniques. In this paper, we use stochastic gradient Langevin dynamics (SGLD)~\citep{welling2011bayesian} that is an MCMC algorithm for drawing samples from a Bayesian posterior and scales to large datasets using mini-batch updates. Please see Sec.~\ref{s:app:langevin} in the Supplement for a brief overview of SGLD. For our application, as lines 3-6 of Alg.~\ref{alg:entropysgd} show SGLD resembles a few iterations of SGD with additive gradient noise.

The gradient of the first term in~\eqref{eq:entropysgd_objective} is computed using back-propagation and the final gradient is
\beq{
    \r\ \rbrac{\f{1}{m}\ \sum_{i=1}^m\ \nabla f(x;\ \xi_{k_i})} + \g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}.
    \label{eq:entropysgd_grad}
}
Note that we use different mini-batches (denoted by $k$ and $\ell$ respectively) to compute the back-propagated gradient and gradient of the local entropy. From our experiments, we observe that using the same mini-batches for the gradients performs marginally worse in terms of convergence speed. This is explained by the fact that local entropy is really a log-partition function and depends upon the entire dataset.

\subsection{Intuition behind local entropy}
\label{ss:intuition_local_entropy}

Let us give some intuition on how local entropy works using the expression for the gradient: the term $\ag{x';\ \cdot}$ is the average over a locally focused Gibbs distribution and for two local minima in the neighborhood of $x$ roughly equivalent in loss, this term points towards the wider one because $\ag{x';\ \cdot}$ is closer to it. This results in a net gradient that takes SGD towards wider valleys. Moreover, if we unroll the SGLD steps used to compute $\rbrac{x - \ag{x'; \cdot}}$ (cf. line 5 in Alg.~\ref{alg:entropysgd}), it resembles one large step in the direction of the (noisy) average gradient around the current weights $x$ and $\entropysgd$ thus looks similar to averaged SGD in the literature~\citep{polyak1992acceleration,bottou2012stochastic}. These two phenomena intuitively explain the improved generalization performance of $\entropysgd$.

The local entropy term in the objective results in a smoother energy landscape than the original loss function. From our experience (cf.\@ experiments in Sec.~\ref{ss:expt:mnist}), we can also set $\rho$ to zero, i.e., consider the modified loss function for $\entropysgd$ to be simply $-F(x, \g)$ and completely eliminate $\nabla f(x)$ from the outer loop of $\entropysgd$. This gives similar generalization error as that of SGD for simpler architectures such as a convolutional neural network on MNIST. However, for more complex datasets and networks, the smooth energy landscape generated by $-F(x, \g)$ might have a different local minimum as compared to the original loss $f(x)$. We found that in order to fine-tune larger networks using the objective in~\eqref{eq:entropysgd_objective} with a small $\rho$ performs slightly better than setting $\rho$ to zero.

\subsection{Algorithm and Implementation details}
\label{ss:alg}

Alg.~\ref{alg:entropysgd} provides the pseudo-code for one iteration of the $\entropysgd$ algorithm. At each iteration, lines $3$-$6$ perform $L$ iterations of Langevin dynamics to estimate $\mu = \ag{x'; \minibatch{\ell}}$. The weights $x$ are updated with the modified gradient on line $9$.

We set the number of SGLD iterations $L$ to $5$-$20$ depending upon the complexity of the dataset, e.g., for experiments on convolutional networks on MNIST and CIFAR-10, we use $L = 20$ while for experiments on a fully-connected network on MNIST $L=5$ works well (cf. Sec.~\ref{s:expt}). We perform exponential averaging of the Langevin iterates $x'$ to estimate $\mu$ with a parameter $\a = 0.9$ so as to put more weight on later samples. In order to quickly transition to the Langevin dynamics phase of SGLD and ensure the we are sampling from the posterior even for small values of $L$, we set the learning rate $\eta'$ on line $5$ to be small factor of the outer SGD's learning rate: $\eta' = 0.1\ \eta$~\citep{welling2011bayesian}.

\begin{center}
\begin{minipage}{0.7 \textwidth}
\IncMargin{0.04in}
\begin{algorithm}[H]
    \SetKwInOut{Input}{Input}
    \SetKwInOut{HyperParameters}{Hyper-parameters}

    \small
    \Input{\quad $\textrm{current weights}\ x$, $\textrm{Langevin iterations}\ L$}
    \HyperParameters{\quad $\textrm{Scope}\ \g$, $\textrm{Data weight}\ \rho$, $\textrm{Learning rate}\ \eta$}
    \vspace{0.1in}
    \nonl \textrm{// SGLD iterations}\;
    \vspace{0.025in}
    $x', \mu \la x$\;
    \For{$\ell \leq L$}
    {
        $\minibatch{\ell} \la \textrm{sample mini-batch}$\;
        \vspace{0.03in}
        $dx' \la \f{1}{m}\ \sum_{i=1}^m\ \nabla_{x'} f\rbrac{x';\ \xi_{\ell_i}} - \g\ \rbrac{x - x'}$\;
        \vspace{0.03in}
        $x' \la x' - \eta'\ dx' + 2\sqrt{\eta'}\ \trm{N}(0, \trm{I})$\;
        \vspace{0.03in}
        $\mu \la (1-\a) \mu + \a\ x'$\;
    }
    \vspace{0.1in}
    \nonl \textrm{// Update weights}\;
    \vspace{0.025in}
    $\minibatch{k} \la \textrm{sample mini-batch}$\;
    \vspace{0.03in}
    $dx \la \f{1}{m}\ \sum_{i=1}^m\ \nabla_{x} f\rbrac{x;\ \xi_{k_i}}$\;
    \vspace{0.03in}
    $x \la x - \eta\ \big(\rho\ dx + \g\ (x-\mu) \big)$
    \caption{$\entropysgd$ algorithm}
    \label{alg:entropysgd}
\end{algorithm}
\DecMargin{0.04in}
\end{minipage}
\end{center}

Let us note that although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$. In our experiments, we have used both SGD with Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.

\subsection{Theoretical Properties}
\label{ss:theoretical_properties}

Our main result in this section is to show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations. Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).

With some overload of notation, we assume that the original loss $f(x)$ is $\b$-smooth, i.e., for all $x,y \in \reals^n$, we have $\norm{\nabla f(x) - \nabla f(y)} \leq \b\ \norm{x-y}$. We additionally assume for the purpose of analysis that no eigenvalue of the Hessian $\nabla^2 f(x)$ lies in the set $[-2 \g-c, c]$ for some small $c > 0$.
%
\begin{lemma}
\label{lem:smoothness_reduction}
The objective $F(x, \g;\ \Xi)$ in~\eqref{eq:entropysgd_objective} is $\f{\a}{1 + \g^{-1}\ c}$-Lipschitz and $\f{\b}{1 + \g^{-1}\ c}$-smooth.
\end{lemma}
The local entropy objective is thus smoother than the original objective while the modified objective in~\eqref{eq:entropysgd_objective} results in a factor of $\r + \f{1}{1 + \g^{-1}\ c}$ in the above lemma.

Let us now obtain a bound on the improvement in generalization error. We denote an optimization algorithm, viz., SGD or $\entropysgd$ by $A(\Xi)$, it is a function of the dataset $\Xi$ and outputs the parameters $x^*$ upon termination. Stability of the algorithm~\citep{bousquet2002stability} is then a notion of how much its output differs in loss upon being presented with two datasets $\Xi$ and $\Xi'$ that differ in at most one sample,
$$
    \sup_{\xi\ \in\ \Xi\ \cup\ \Xi'}\ \sqbrac{ f \rbrac{A(\Xi), \xi} - f \rbrac{A(\Xi'), \xi}} \leq \e.
$$
\citet{hardt2015train} connect uniform stability to generalization error and show that an $\e$-stable algorithm $A(\Xi)$ has generalization error bounded by $\e$, i.e., if $A(\Xi)$ terminates with parameters $x^*$,
$$
    \abs{ \E_{\Xi} \rbrac{R_\Xi(x^*) - R(x^*)} } \leq \e;
$$
where the left hand side is the generalization error: it is the difference between the empirical loss $R_\Xi(x) := \f{1}{N}\ \sum_{k=1}^N\ f(x, \xi_k)$ and the population loss $R(x) := \E_{\xi}\ f(x, \xi)$.

We now employ the following theorem that bounds the stability of an optimization algorithm through the smoothness of its loss function and the number of iterations on the training set.
\begin{theorem}[\citet{hardt2015train}]
\label{thm:generalization_bound_nonconvex}
For an $\a$-Lipschitz and $\b$-smooth loss function, if SGD converges in $T$ iterations on $N$ samples with decreasing learning rate $\eta_t \leq 1/t$ the stability is bounded by
\aeqs{
    \e
    %&\leq \f{1}{N-1}\ \rbrac{1 + \f{1}{\b c}}\ \rbrac{2 c \a^2}^{\f{1}{1 + \b c}}\ T^{\f{\b c}{1 + \b c}}\\
    &\lessapprox \f{1}{N}\ \a^{1/(1 + \b)}\ T^{1 - 1/(1 + \b)}.
}
\end{theorem}
%
Using Lem.~\ref{lem:smoothness_reduction} and Thm.~\ref{thm:generalization_bound_nonconvex} we have
\beq{
    \e_{\ \entropysgd} \lessapprox \rbrac{\a\ T^{-1}}^{\rbrac{1 - \f{1}{1 + \g^{-1}c}}\ \b}\ \e_{\ \textrm{SGD}},
    \label{eq:stability_bound}
}
which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$. For the objective~\eqref{eq:entropysgd_objective}, we have the constant $1 - \r - \f{1}{1 + \g^{-1}\ c}$ in the exponent instead.

As an aside, it is easy to see from the proof of Lem.~\ref{lem:smoothness_reduction} that for a convex loss function $f(x)$, adding the local entropy term does not change the minimizer of the original problem.

\section{Experiments}
\label{s:expt}

In Sec.~\ref{ss:expt:universality}, we discuss experiments that suggest that the characteristics of the energy landscape around local minimal accessible by SGD are universal to deep architectures.  This observation motivates the $\entropysgd$ algorithm considered in this paper and in Sec.~\ref{ss:expt:mnist} and~\ref{ss:expt:cifar}, we present experimental results on two standard image classification datasets, viz., MNIST and CIFAR-10.

\subsection{Universality of the Hessian at local minima}
\label{ss:expt:universality}

We use automatic differentiation\footnote{\href{https://github.com/HIPS/autograd}{https://github.com/HIPS/autograd}} to compute the Hessian at a local minimum obtained at the end of training for the following networks:

\begin{enumerate}[(i)]
\item \textbf{small-LeNet on MNIST}: This network has $47,658$ parameters and is similar to $\lenet$ but with $10$ and $20$ channels respectively in the first two convolutional layers and $128$ hidden units in the fully-connected layer. We train this with Adam to obtain a test error of $2.4\%$.
%
\item \textbf{small-$\mnistfc$ on MNIST}: A fully-connected network ($50,890$ parameters) with one layer of $32$ hidden units, ReLU non-linearities and cross-entropy loss; it converges to a test error of $2.5\%$ with momentum-based SGD.
%
\item \textbf{char-lstm for text generation}: This is a recurrent network with $48$ hidden units and Long Short-Term Memory (LSTM) architecture~\citep{hochreiter1997long}. It has $32,640$ parameters and we train it with Adam to re-generate a small piece of text consisting of $256$ lines of length $32$ each and $96$-bit one-hot encoded characters.
%
\item \textbf{$\allcnn$ on CIFAR-10}: This is similar to the All-CNN-C network~\citep{springenberg2014striving} with $\approx 1.6$ million weights (cf. Sec.~\ref{ss:expt:cifar}) which we train using Adam to obtain an error of $11.2\%$. Exact Hessian computation is in this case expensive and thus we instead compute the diagonal of the Fisher information matrix~\citep{wasserman2013all} using the element-wise first and second moments of the gradients that Adam maintains, i.e., $\trm{diag} (I) = \E(g^2) - (\E\ g)^2$ where $g$ is the back-propagated gradient. Fisher information measures the sensitivity of the log-likelihood of data given parameters in a neighborhood of a local minimum and thus is exactly equal to the Hessian of the negative log-likelihood. We will consider the diagonal of the empirical Fisher information matrix as a proxy for the eigenvalues of the Hessian, as is common in the literature.
\end{enumerate}

We choose to compute the exact Hessian and to keep the computational and memory requirements manageable, the first three networks considered above are smaller than standard deep networks used in practice. For the last network, we sacrifice the exact computation and instead approximate the Hessian of a large deep network. We note that recovering an approximate Hessian from Hessian-vector products~\citep{pearlmutter1994fast} could be a viable strategy for medium-scale networks.

Fig.~\ref{fig:lenet_hessian} in the introductory Sec.~\ref{s:intro} shows the eigenspectrum of the Hessian for small-LeNet while Fig.~\ref{fig:universality} shows the eigenspectra for the other three networks.
%
A large proportion of eigenvalues of the Hessian are very close to zero or positive with a very small (relative) magnitude. This is suggests that the local geometry of the energy landscape is almost flat at local minima discovered by gradient descent. This agrees with theoretical results such as~\citet{baldassi2016local} where the authors predict that flat regions of the landscape generalize better. Standard regularizers in deep learning such as convolutions, max-pooling and dropout seem to bias SGD towards flatter regions in the energy landscape.
%
The right tails of the eigenspectra are much longer than the left tails. Indeed, as discussed in numerous places in literature~\citep{Bray2007,dauphin2014identifying,spinglass2015}, SGD finds low-index critical points, i.e., optimizers with few negative eigenvalues of the Hessian. What is interesting and novel is that the directions of descent that SGD misses do not have a large curvature.

\begin{figure}[!tbh]
\centering
    \begin{subfigure}[t]{0.47 \textwidth}
        \includegraphics[width=1.08\textwidth]{mnistfc_hessian.pdf}
        \caption{\small $\smallmnistfc$ ($2$ runs): Peak (clipped here) at zero ($\abs{\lambda} \leq 10^{-2}$) accounts for $90\%$ of the entries.\vspace{0.15in}}
        \label{fig:mnistfc_hessian}
    \end{subfigure}
    \hspace{0.2in}
    \begin{subfigure}[t]{0.47 \textwidth}
        \includegraphics[width=\textwidth]{charlstm_hessian.pdf}
        \caption{\small $\charlstm$ ($5$ runs): Almost $95\%$ eigenvalues have absolute value below $10^{-5}$.}
        \label{fig:charlstm_hessian}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \begin{subfigure}[b]{0.47 \textwidth}
        \includegraphics[width=\textwidth]{allcnn_dm.pdf}
        \end{subfigure}
        \hspace{0.2in}
        \begin{subfigure}[b]{0.47 \textwidth}
        \includegraphics[width=\textwidth]{allcnn_dp.pdf}
        \end{subfigure}
    \caption{\small Negative and positive eigenvalues of the Fisher information matrix of $\allcnn$ at a local minimum ($4$ independent runs). The origin has a large peak with $\approx 95\%$ near-zero ($\abs{\lambda} \leq 10^{-5}$) eigenvalues (clipped here).}
    \label{fig:allcnn_hessian}
    \end{subfigure}
\caption{\small Universality of the Hessian: for a wide variety of network architectures, sizes and datasets, optima obtained by SGD are mostly flat (large peak near zero), they always have a few directions with large positive curvature (long positive tails). A very small fraction of directions have negative curvature, and the magnitude of this curvature is extremely small (short negative tails).}
\label{fig:universality}
\end{figure}

\subsection{MNIST}
\label{ss:expt:mnist}

In this section, we show that local entropy is a good measure of generalization, in fact, setting $\rho = 0$ in $\entropysgd$, i.e., completely disabling the original loss function is enough to train on MNIST. We consider two networks: the first is a fully-connected network with two hidden layers of $1024$ units each which we denote as ``$\mnistfc$'' and the second is a convolutional neural network similar to $\lenet$ but which batch-normalization~\citep{ioffe2015batch} and a dropout layer of probability $0.5$ added after each convolutional layer. We train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs and compare the average error over $5$ independent runs in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test}.

$\entropysgd$ converges to an error of $1.39\%$ in $22$ epochs on $\mnistfc$ while Adam obtains the same (best) error after $67$ epochs. For a small network such as this, with a presumably easy energy landscape, we perform $5$ SGLD iterations and set $\g = 1$. For $\lenet$, we perform $20$ SGLD iterations and set $\g = 0.1$. In this case, $\entropysgd$ converges around epoch $40$ to an average error of $0.48\%$; Adam obtains a comparable error of $0.51\%$ after $100$ epochs. For this experiment, we use a small $\rho = 10^{-4}$ to fine-tune the network; the error without this is $0.52\%$.

\textbf{Remark:} A single ``epoch'' for all algorithms considered here corresponds to the number of parameter updates necessary to run through the entire dataset once. Note that both SGD and $\entropysgd$ update parameters after each mini-batch, but $\entropysgd$ also runs $L$ iterations of SGLD between two updates of the parameters.

\begin{figure}[h!]
\centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=0.98\textwidth]{mnistfc_valid.pdf}
        \caption{\small $\mnistfc$: $\entropysgd$ with $\rho = 0$, $L = 5$.}
        \label{fig:mnistfc_test}
    \end{subfigure}
    \hspace{0.1in}
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{lenet_valid.pdf}
        \caption{\small $\lenet$: $\entropysgd$ $\rho = 10^{-4}$, $L = 20$.}
        \label{fig:lenet_test}
    \end{subfigure}
\caption{\small Comparison of $\entropysgd$ vs. Adam on MNIST\vspace*{0.15in}}
\label{fig:mnist_test}
\end{figure}

\begin{figure}[htp!]
\centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=1.03\textwidth]{allcnn_loss.pdf}
        \caption{\small $\allcnn$: Training loss}
        \label{fig:allcnn_loss}
    \end{subfigure}
    \hspace{0.1in}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{allcnn_valid.pdf}
        \caption{\small $\allcnn$: Validation error}
        \label{fig:allcnn_valid}
    \end{subfigure}
\caption{\small Comparison of $\entropysgd$ vs. Adam on CIFAR-10}
\label{fig:allcnn}
\end{figure}

\subsection{CIFAR-10}
\label{ss:expt:cifar}

We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}. As a baseline, we consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer; this is absent in the original paper. We recreate the training pipeline of the original authors with the same hyper-parameters and use SGD with Nesterov's momentum. We however decrease the initial learning rate of $0.1$ by a factor of $5$ after every $60$ epochs for SGD and after every $5$ epochs for $\entropysgd$. Probably due to batch-normalization, we obtain an average error of $8.30\%$ over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report. This gives us a very competitive baseline to compare $\entropysgd$ against. Let us note that the best result in literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error while a wide ResNet of~\citet{zagoruyko2016wide} obtains an error of $4.17\%$ with data augmentation.

%We list a few other comparable networks in the literature in Table~\ref{tab:cifar10}.
% \begin{wraptable}{r}{0.6 \textwidth}
% \centering
% \renewcommand{\arraystretch}{1.2}
% \resizebox{0.5 \columnwidth}{!}
% {
% \begin{tabular}{p{6.5cm} r}
%     Model & Test error ($\%$)\\  \hline
%     %Maxout~\citep{goodfellow2013maxout} & $11.68$\\
%     NiN~\citep{lincy13} & $10.41$\\
%     All-CNN-C~\citep{springenberg2014striving} & $9.04$\\
%     All-CNN-BN (baseline: SGD, 200 epochs) & $8.30\ \pm\ 0.08$\\
%     All-CNN-BN ($\entropysgd$, 15 epochs) & $8.65\ \pm\ 0.06$\\
%     ELU~\citep{clevert2015fast} & $6.55$\\ \hline
% \end{tabular}
% }
% \caption{\small Error rates on CIFAR-10 without data augmentation.}
% \label{tab:cifar10}
% \end{wraptable}

Fig.~\ref{fig:allcnn} compares $\entropysgd$ against SGD with $20$ SGLD iterations for $15$ epochs. Since CIFAR-10 is a more challenging dataset than MNIST, we could not obtain better test errors than $9.5\%$ while training with only the local entropy term ($\rho = 0$). We therefore set $\rho = 0.01$ to obtain an error of $8.65\%$ over $5$ independent runs. Refer to Fig.~\ref{fig:allcnn_loss}, $\entropysgd$ is designed to be an algorithm that seeks out wide local minima as opposed to sharp valleys; a side-effect of this is to smooth the energy landscape as we show in Sec.~\ref{ss:theoretical_properties}. In practice, we see this as a steady decrease in training loss for $\entropysgd$ without any plateaus that appear in the training loss for SGD. Similarly, the decrease in validation error in Fig.~\ref{fig:allcnn_valid} is also steady without any plateaus. Note that the original network $\textrm{All-CNN-C}$ was trained using SGD for $350$ epochs~\citep{springenberg2014striving}.

\section{Discussion}
\label{s:discussion}

As discussed in Sec.~\ref{ss:intuition_local_entropy}, SGLD updates in the inner loop of our algorithm compute the average gradient of the original loss in the neighborhood of the current weights. An inexpensive and approximate way of computing these updates is to average the gradient over random perturbations of weights during the forward pass itself and, on the same mini-batch as the outer loop. In our experiments, using the same mini-batch performs slightly worse than $\entropysgd$ although it is still better than SGD. Such an implementation is close to averaged SGD of~\citep{polyak1992acceleration,bottou2012stochastic} and worth further study.

In our experiments, $\entropysgd$ results in a comparable generalization error as SGD, but always has a lower cross-entropy loss. Roughly speaking, wide valleys favored by $\entropysgd$ seem to have lower empirical loss than local minima discovered by SGD. Theoretical models of deep networks (cf.\@ Sec.~\ref{s:prior_work}) suggest multiple local minima with the same loss but there is an absence of results in literature about the local geometry at these minima.

\section{Conclusions}
\label{s:conclusions}

We introduced an algorithm named $\entropysgd$ for optimization of deep networks. This was motivated from the observation that the energy landscape near a local minimum discovered by SGD is almost flat for a wide variety of deep networks irrespective of their architecture, input data or training methods. We connected this observation to the concept of local entropy which we used to bias the optimization towards flat regions that have low generalization error. Our experiments showed that this algorithm is applicable to large deep networks used in practice.

\section{Acknowledgements}
\label{s:acknowledgements}

This work was supported by the Office of Naval Research (ONR), Air Force Office of Scientific Research (AFOSR) and Army Research Office (ARO).

{
\footnotesize
\bibliographystyle{iclr2017_conference}
\bibliography{chaudhari.choromanska.ea.iclr17}
}

\begin{appendices}

\renewcommand\thetable{\thesection\arabic{table}}
\renewcommand\thefigure{\thesection\arabic{figure}}

\section{Stochastic gradient Langevin dynamics (SGLD)}
\label{s:app:langevin}

Local entropy in Def.~\eqref{def:local_entropy} is an expectation over the entire configuration space $x \in \reals^n$ and is hard to compute, we can however approximate its gradient using Markov chain Monte-Carlo (MCMC) techniques. In this section, we briefly review stochastic gradient Langevin dynamics~\citep{welling2011bayesian} that is an MCMC algorithm designed to draw samples from a Bayesian posterior and scales to large datasets using mini-batch updates.

For a parameter vector $x \in \reals^n$ with a prior distribution $p(x)$ and if the probability of generating a data item $\xi_k$ given a model parameterized by $x$ is $p(\xi_k \given x)$, the posterior distribution of the parameters based on $N$ data items can be written as
\beq{
    p\rbrac{x \given \xi_{\ k \leq N}}\ \propto\ p(x)\ \prod_{k=1}^N\ p\rbrac{\xi_k \given x}.
    \label{eq:bayesian_posterior}
}
Langevin dynamics~\citep{neal2011mcmc} injects Gaussian noise into maximum-a-posteriori (MAP) updates to prevent over-fitting the solution $x^*$ of the above equation. The updates can be written as
\beq{
    \Delta x_t = \f{\eta}{2}\ \rbrac{ \nabla \log p(x_t) + \sum_{k=1}^N\ \nabla p(\xi_k \given x_t)} + \sqrt{\eta}\ \e_t;
    \label{eq:langevin_bayesian}
}
where $\e_t \sim \trm{N}(0, 1)$ is Gaussian noise and $\eta$ is the learning rate. In this form, Langevin dynamics faces two major hurdles for applications to large datasets. First, computing the gradient $\sum_{k=1}^N\ \nabla p(\xi_k \given x_t)$ over all samples for each update $\Delta x_t$ becomes prohibitive. However, as~\citet{welling2011bayesian} show, one can instead simply use the average gradient over $m$ data samples (mini-batch) as follows:
\beq{
    \Delta x_t = \f{\eta_t}{2}\ \rbrac{ \nabla \log p(x_t) + \f{N}{m}\ \sum_{k=1}^m\ \nabla p(\xi_k \given x_t)} + \sqrt{\eta_t}\ \e_t.
    \label{eq:sgld_bayesian}
}
Secondly, Langevin dynamics in~\eqref{eq:langevin_bayesian} is the discrete-time approximation of a continuous-time stochastic differential equation~\citep{mandt2016variational} thereby necessitating a Metropolis-Hastings (MH) rejection step~\citep{roberts2002langevin} which again requires computing $p(\xi_k \given x)$ over the entire dataset. However, if the learning rate $\eta_t \to 0$, we can also forgo the MH step~\citep{chen2014stochastic}. \citet{welling2011bayesian} also argue that the sequence of samples $x_t$ generated by updating~\eqref{eq:sgld_bayesian} converges to the correct posterior~\eqref{eq:bayesian_posterior} and one can hence compute the statistics of any function $g(x)$ of the parameters using these samples. Concretely, the posterior expectation $\E\sqbrac{g(x)}$ is given by
$
    \E\sqbrac{g(x)} \approx \f{\sum_{s=1}^t \eta_t\ g(x_t)}{\sum_{s=1}^t \eta_t};
$
which is the average computed by weighing each sample by the corresponding learning rate in~\eqref{eq:sgld_bayesian}. In this paper, we will consider a uniform prior on the parameters $x$ and hence the first term in~\eqref{eq:sgld_bayesian}, viz., $\nabla \log p(x_t)$ vanishes.

Let us note that there is a variety of increasingly sophisticated MCMC algorithms applicable to our problem, e.g., Hamiltonian Monte Carlo (SGHMC) by~\citet{chen2014stochastic} based on volume preserving flows in the ``parameter-momentum'' space, stochastic annealing thermostats (Santa) by~\citet{chen2015bridging} etc. We can also employ these techniques, although we use SGLD for ease of implementation. The authors in~\citet{ma2015complete} provide an elaborate overview of these methods.

\section{Proofs}
\label{s:app:proofs}

\begin{proof}[Proof of Lemma~\ref{lem:smoothness_reduction}]
The gradient $-\nabla F(x)$ is computed in Sec.~\ref{ss:grad_local_entropy} to be $\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}$. Consider the term
\aeqs{
    x - \ag{x';\ x}
    &= x - Z_{x,\g}^{-1}\ \int_{x'}\ x'\ e^{-f(x') - \f{\g}{2}\ \norm{x-x'}^2}\ dx'\\
    &\approx x - Z_{x,\g}^{-1}\ \int_{s}\ (x + s)\ e^{-f(x) - \nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}\ ds\\
    &= x\rbrac{1 -  Z_{x,\g}^{-1}\ \int_{s}\ e^{-f(x) - \nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}\ ds} - Z_{x,\g}^{-1}\ \int_{s}\ s\ e^{-f(x) - \nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}\ ds\\
    &= -Z_{x,\g}^{-1}\ e^{-f(x)}\ \int_{s}\ s\ e^{-\nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}\ ds.
}
The above expression is the mean of a distribution $\propto e^{-\nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}$. We can approximate it using the saddle point method as the value of $s$ that minimizes the exponent to get
$$
    x - \ag{x';\ x} \approx \rbrac{\nabla^2 f(x) + \g\ I}^{-1}\ \nabla f(x).
$$
Let us denote $A(x) := \rbrac{I + \g^{-1}\ \nabla^2 f(x)}^{-1}$. Plugging this into the condition for smoothness, we have
\aeqs{
    \norm{\nabla F(x, \g) - \nabla F(y, \g)}
    &= \norm{A(x)\ \nabla f(x) - A(y)\ \nabla f(y)}\\
    &\leq \rbrac{\sup_x\ \norm{A(x)}}\ \b\ \norm{x-y}.
}
Unfortunately, we can only get a uniform bound if we assume that for a small constant $c > 0$, no eigenvalue of $\nabla^2 f(x)$ lies in the set $[-2 \g -c, c]$. This gives
$$
    \rbrac{\sup_x\ \norm{A(x)}} \leq \f{1}{1+\g^{-1}\ c}.
$$
This shows that a smaller value of $\g$ results in a smoother energy landscape, except at places with very flat directions. The Lipschitz constant also decreases by the same factor. The objective $\r f(x) - F(x,\g)$ results in a modified factor $\r + \f{1}{1+\g^{-1}\ c}$.
\end{proof}

\ignore{
\section{Sub-dominant clusters in a spin glass}
\label{s:app:spin_glass}

As an pedagogical example of a local entropy based loss function, we first consider the optimization of a well-studied model in statistical physics known as a mixed spherical spin glass~\citep{talagrand2003spin}. For $J^2_{ijk},\ J^3_{ijk} \sim \trm{N}(0,1)$, the energy of a spin glass is given by
\beq{
    H(x) = \f{1}{10\ \sqrt{n}}\ \sum_{i,j=1}^n\ J^2_{ij}\ x_i\ x_j + \f{1}{n} \sum_{i,j,k=1}^n\ J^3_{ijk}\ x_i\ x_j\ x_k
    \label{eq:sg_hamiltonian}
}
where the parameter vector $x \in \reals^n$ is called the ``spin''. A spherical spin glass is a Gaussian field on the sphere, i.e., we have an additional constraint $\norm{x}^2_2 = n$ or $x \in S^{n-1}(\sqrt{n})$. We are interested in computing
\beq{
    x^*_H = \argmin_{x \in S^{n-1}(\sqrt{n})}\ H(x).
    \label{eq:3sg}
}
This is a non-convex problem and in general, hard; the number of local minima and saddle points of $H(x)$ scale exponentially in $n$~\citep{auffinger2013random}. However, using results from statistical physics, one can show that all local minima concentrate around the global minimum~\citep{panchenko2013sherrington}:
$$
    \lim_{n \to \infty}\ \inf_{x \in S^{n-1}(\sqrt{n})}\ \f{H(x)}{n} \approx -1.8227.
$$
We have chosen a mixed spin glass in~\eqref{eq:sg_hamiltonian} to ensure that the Hamiltonian has both deep, narrow local minima belonging to the $3$-spin term $x_i x_j x_k$ and wide, shallower valleys resulting from the $2$-spin term $x_i x_j$ with a small curvature.
%Moreover, since the Hamiltonian is isotropic, all local minima have an almost equivalent local geometry of the energy landscape, in particular, the Euclidean Hessian $\nabla^2 H(x) \in \reals^{(n-1) \times (n-1)}$ belongs to the Gaussian Orthogonal Ensemble (modulo a constant). 
Our modified optimization problem with a negative local entropy term is given by
\beq{
    x^*_f = \argmin_{x \in S^{n-1}(\sqrt{n})}\ \rho H(x) - F(x, \g);
    \label{eq:3sg_modified}
}
with $\rho = 0.01$ and $\g = 10^{-4}$ and the local entropy $F(x, \g)$ is the log-partition function of
$$
    \P(x';\ x, \g) \propto \exp \rbrac{-H(x') - \f{\g}{2}\ x^\top\ x'}
$$
is estimated using Langevin dynamics with $100$ iterations in Alg.~\ref{alg:entropysgd}. In a minor deviation from the development in Sec.~\ref{s:entropysgd}, we use the dot product $x^\top x'$ as the biasing term. With the spherical constraint $\norm{x}^2_2 = n$ in play, these two methods are equivalent.

For both~\eqref{eq:3sg} and~\eqref{eq:3sg_modified} we set $n = 100$ and perform $10^3$ Monte-Carlo runs from uniformly randomly sampled initial conditions on the hyper-sphere with a learning rate $\eta = 0.01$ until the norm of the gradient is smaller than $10^{-4}$. We then contrast the eigenspectrum of the original Hessian, viz., $\nabla^2 H(x)$ (projected onto the tangent space of the sphere) for both these problems.

Fig.~\ref{fig:expt:sg} shows the results of this simulation. As expected, the solutions of~\eqref{eq:3sg} have energies close to the asymptotic value $-1.657$. The energies of the optimizers of~\eqref{eq:3sgentropy} are higher, with a mean $-1.45$. On the other hand, the Hessian for~\eqref{eq:3sg} has large positive eigenvalues suggesting deep, sharp local minima while the Hessian for~\eqref{eq:3sgentropy} has much smaller positive eigenvalues, which suggests shallower and wider valleys in the energy landscape. The effect of the local-entropy term is thus to artificially under-optimize the loss function; indeed it avoids deep and sharp valleys in the energy landscape and causes gradient descent to settle in wider --- although shallower --- local minima where nearby spins have similar values of the Hamiltonian $H(x)$. As experiments in Sec.~\ref{ss:expt:mnist} and~\ref{ss:expt:cifar} show, this is crucial to avoid over-fitting on the training set and leads to better generalization.
}

\end{appendices}

\end{document}

### 2016-12-17 doc/rebuttal.txt
wip on rebuttal

"mostly reinventing SVI"
SVI computes the posterior distribution of hidden variables given data by maximizing the ELBO. We do not have hidden variables in our formulation. Note that variable "x" are the variables of our objective f(x), not data. We are interested in maximizing the free energy F(x,\gamma) (related to the entropy of the optimization objective f(x)) as defined in Eqn. 4. This is totally unrelated to marginal likelihood or variational ELBO.

"I suspect it'd become clear that you are maximizing the marginal likelihood"
We are not maximizing the marginal likelihood.

"how is this better than SGLD", "bizzare that you compare against Adam"
SGLD is an MCMC algorithm that draws samples from a given distribution. If the step-size goes to zero slowly enough, akin to all MCMC algorithms, it converges to the maximum of the likelihood in exponentially long time-scales (see algorithms like SGHMC [CFG14], Santa [CCG15] etc. that train Bayesian neural networks using MCMC algorithms). Note, this is not the marginal likelihood because there is no notion of hidden variables in SGLD. Please see Appendix A for more details.

On the other hand, Adam is an algorithm for computing the maxmimum of a the likelihood of data given parameters or (equivalently, minimize the loss function). Entropy-SGD is desgined for minimizing the loss function f(x). It however does not explicitly do so and instead maximizes the local entropy. We therefore compare the two.

Local entropy (local free energy) is defined in Def. 1 (Eqn. 4) while the discussion (pg. 5, first line) defines the classical entropy of a Gibbs distribution. The beginning of Sec. 3 based on Fig. 2 is intended to explain things to the read at an intuitive level.

[CFG14] Chen, Tianqi, Emily B. Fox, and Carlos Guestrin. "Stochastic Gradient Hamiltonian Monte Carlo." ICML. 2014.
[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).

### 2016-12-17 dnn/preprocess_cifar10.py
get no-dropout + SGD runs perfect
minor changes in code

@@ -1,3 +1,5 @@
+'From Taco Cohens github'
+
 import os
 import numpy as np
 

### 2016-12-17 dnn/preprocess_cifar100.py

@@ -1,3 +1,5 @@
+'From Taco Cohens github'
+
 import os
 import numpy as np
 import cPickle as pickle

### 2016-12-20 doc/rebuttal.txt
minor changes to AC rebuttal

@@
+We are happy to provide clarifications to the questions posed. We enclose them below.
+
 "mostly reinventing SVI"
 SVI computes the posterior distribution of hidden variables given data by maximizing the ELBO.
-We do not have hidden variables in our formulation.
-Note that variable "x" are the variables of our objective f(x), not data.
-We are interested in maximizing the free energy F(x,\gamma) (related to the entropy of the optimization objective f(x)) as defined in Eqn.
+Please note that we do not have hidden variables in our formulation; "x" are the variables of our objective f(x), not data.
+We are interested in maximizing the free energy F(x,\gamma) as defined in Eqn.
 4.
-This is totally unrelated to marginal likelihood or variational ELBO.
+Simply speaking, this is the *free energy of the energy landscape* (defined via a Gibbs distribution).
+This is thus unrelated to marginal likelihood or variational ELBO where one can impose entropic constraints on the posterior distribution of the *hidden variables* via a prior.
 
 "I suspect it'd become clear that you are maximizing the marginal likelihood"
-We are not maximizing the marginal likelihood.
+Please note that we are not maximizing the marginal likelihood.
 
-"how is this better than SGLD", "bizzare that you compare against Adam"
+"how is this better than SGLD", "bizarre that you compare against Adam instead of SGLD"
 SGLD is an MCMC algorithm that draws samples from a given distribution.
@@
 that train Bayesian neural networks using MCMC algorithms).
-Note, this is not the marginal likelihood because there is no notion of hidden variables in SGLD.
-Please see Appendix A for more details.
+Note that SGLD does not optimize the marginal likelihood because there is no notion of hidden variables in vanilla SGLD.
+We provide a brief review of SGLD in Appendix A of our paper.
 
-On the other hand, Adam is an algorithm for computing the maxmimum of a the likelihood of data given parameters or (equivalently, minimize the loss function).
-Entropy-SGD is desgined for minimizing the loss function f(x).
+We do not know of any results in literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD.
+In our experience, vanilla SGLD performs poorly due to non-energy preserving dynamics while more complex algorithms like [CFG14 or CCF15] have too many sensitive hyper-parameters.
+We would like to emphasize that Entropy-SGD simply uses MCMC sampling to estimate Eqn.
+7, it is unrelated to SGLD otherwise.
+
+On the other hand, Adam is an algorithm for computing the maximum of a the likelihood of data given parameters or (equivalently, minimize the loss function).
+Entropy-SGD is an algorithm designed for minimizing the loss function f(x), in particular, it is not an MCMC algorithm that draws samples from the likelihood.
 It however does not explicitly do so and instead maximizes the local entropy.
-We therefore compare the two.
+We therefore compare Entropy-SGD with state-of-the-art algorithms for training deep networks like Adam and SGD.
 
-Local entropy (local free energy) is defined in Def.
+"frustrating that you discuss free energy and entropy without precise definitions"
+Local entropy (local free energy) is formally defined in Def.
 1 (Eqn.
-4) while the discussion (pg.
-5, first line) defines the classical entropy of a Gibbs distribution.
+4) but it is introduced as early as pg.
+2 in the Introduction.
+The discussion towards the end of Sec.
+3 (pg.
+5 first line) defines the classical entropy of a Gibbs distribution.
 The beginning of Sec.
 3 based on Fig.
-2 is intended to explain things to the read at an intuitive level.
+2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which involve a little bit of jargon.
+
+Free energy and classical entropy are related by the informal description "free energy = internal energy - temperature x entropy". Formally, the relation is
+    F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
+where F(\beta) is the log-partition function (free energy) defined as -\beta^{-1} \log Z(\beta), U(\beta) = \partial_\beta (\beta F(\beta)) is the internal energy and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.
+
+Note that the above equation is about the entirety of the Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn.
+3.
+We can add this discussion to the paper.
 

### 2016-12-22 doc/rebuttal.txt

@@
 4.
-Simply speaking, this is the *free energy of the energy landscape* (defined via a Gibbs distribution).
+This is the *free energy of the energy landscape* (defined via a Gibbs distribution).
 This is thus unrelated to marginal likelihood or variational ELBO where one can impose entropic constraints on the posterior distribution of the *hidden variables* via a prior.
@@
 "I suspect it'd become clear that you are maximizing the marginal likelihood"
-Please note that we are not maximizing the marginal likelihood.
+Please note that we are not maximizing the marginal likelihood, as discussed above.
 
@@
 
-We do not know of any results in literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD.
-In our experience, vanilla SGLD performs poorly due to non-energy preserving dynamics while more complex algorithms like [CFG14 or CCF15] have too many sensitive hyper-parameters.
+We do not know of any results in the literature that train large deep networks such as the one used for CIFAR to get competitive error rates using SGLD.
 We would like to emphasize that Entropy-SGD simply uses MCMC sampling to estimate Eqn.
-7, it is unrelated to SGLD otherwise.
+7 but it is unrelated to SGLD otherwise.
 
@@
 1 (Eqn.
-4) but it is introduced as early as pg.
+4) but it is already introduced on pg.
 2 in the Introduction.
@@
 3 (pg.
-5 first line) defines the classical entropy of a Gibbs distribution.
+5, first line) defines the classical entropy of a Gibbs distribution.
 The beginning of Sec.
 3 based on Fig.
-2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which involve a little bit of jargon.
+2 is intended to explain things to the reader at an intuitive level before proceeding to the formal definitions which are nevertheless present.
 
@@
     F(\beta) = U(\beta) - \frac{1}{\beta} S(\beta)
-where F(\beta) is the log-partition function (free energy) defined as -\beta^{-1} \log Z(\beta), U(\beta) = \partial_\beta (\beta F(\beta)) is the internal energy and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.
+where the log-partition function (free energy) is defined as F(\beta) = -\beta^{-1} \log Z(\beta), the internal energy is U(\beta) = \partial_\beta (\beta F(\beta)) and S(\beta) = \beta^2 \partial_\beta (F(\beta)) is the classical entropy.
 
-Note that the above equation is about the entirety of the Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn.
+Note that the above equation is about the entire Gibbs distribution, in our work we define a "local free energy" F(x, \beta) via the modified Gibbs distribution in Eqn.
 3.

### 2016-12-22 doc/rebuttal.txt

@@
+AC rebuttal
+==========
+
 We are happy to provide clarifications to the questions posed. We enclose them below.
@@
 [CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).
+
+----------------------------------------------------------------------
+
+Thanks for your comments, allow us to elaborate further.
+
+In the notation of the our paper let x denote the parameters and \xi_k denote individual samples. The ELBO is then given by
+    L = E_{x \sim q(x)} [\log p(\Xi | x)] - KL(q(x) || p(x))
+where p(x) is a fixed prior on the parameters and q(x) is the posterior distribution q(x | \Xi).
+In mean-field variational inference, we make a mean-field approximation of the distribution q(x) and have to come up with a prior p(x) (see [GG15,BKM16]).
+Using a Monte-Carlo approximation, the gradient of ELBO can be given by
+    dL = \sum_{k=1}^N \log p(\xi | x_k) - KL(q(x) || p(x))
+for samples x_k \sim q(x).
+The first term is simply how well x fits the data while the second term is the distance of the posterior from the prior.
+If p(x) = N(0, I), one can evaluate the KL term in closed form if q(x) is also a product of Gaussians etc.
+and compute dL.
+Note that this approach requires a number of critical assumptions such as a factorizing of q(x) for ease of sampling in the Monte-Carlo term and evaluating the KL term, Gaussian / some other easy to compute prior on weights.
+These assumptions are not typically valid in CNNs.
+
+The gradient in Eqn. 7 in our paper can be written as:
+    dF = -\gamma (x - E_{x' \sim r(x'; x, \xi_k)} [x'])
+with the corresponding Monte-Carlo estimate of the expectation. The crucial difference is that the distribution r(x'; x) in our case is the one described in Eqn. 3, i.e.,
+    \log r(x'; x) = \log q(x')  + \gamma |x-x'|^2/2 - \log Z_{x,\gamma}.
+q(x') can be the same distribution as variational inference above.
+However, note that r(x'; x) in our work is a function of the current iterate x itself.
+This is akin to a "prior" that moves with the iterates x which is very different from a Bayesian neural network with a fixed prior.
+
+It is thus easy to see the differences between a variational approach on the weights and our method.
+We agree that the motivation for our work and approximate inference methods is similar but our method is much different from variational approximation.
+We'd be very interested if the reviewer has other ways of connecting our paper with it.
+
+"SGLD is a simpler method"
+We respectfully disagree.
+In practice, it is very hard to get SGLD to work as well as SGD or Adam.
+This has motivated numerous other MCMC algorithms such as SGHMC, SGNHT [DFB14], Santa [CCF15] etc.
+None of these have yet been shown to beat SGD's error rates for large deep networks.
+
+While it is not hard to add experiments comparing Entropy-SGD to SGLD, this would be a digression.
+The error rates for baselines in our paper are close to the best error rates reported in the literature for these networks, e.g., on CIFAR, we get 8.3% with SGD and 8.65% with Entropy-SGD both of which are better than those of the original authors (9.08%) [SDB15].
+The authors in [CCF15] get a comparable error rate on MNIST (0.47% vs 0.48% for us) with a larger network (Table 1) and a much more complex algorithm.
+[CCF15] also provides the results for SGLD on LeNet which are much worse than ours.
+Therefore, we feel our comparisons against baselines are entirely fair.
+
+[GG15] Gal, Yarin, and Zoubin Ghahramani. "Dropout as a Bayesian approximation: Representing model uncertainty in deep learning." arXiv:1506.02142 (2015).
+[BKM16] Blei, David M., Alp Kucukelbir, and Jon D. McAuliffe. "Variational inference: A review for statisticians." arXiv:1601.00670 (2016).
+[SDB15] Springenberg, Jost Tobias, et al. "Striving for simplicity: The all convolutional net." arXiv:1412.6806 (2014).
+[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
+[DFB14] Ding, Nan, et al. "Bayesian sampling using stochastic gradient thermostats." NIPS. 2014.

### 2016-12-29 doc/rebuttal.txt
wip on AC rebuttal

--- 2016-12-22 doc/rebuttal.txt
@@
 [CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).
+
+----------------------------------------------------------------------
+
+Thanks for your comments, allow us to elaborate.
+
+Let \Xi denote the dataset, z denote the weights and x be the parameters of the variational distribution on weights, i.e., the weights are sampled from the distribution q_x(z) := p_x(z | \Xi).
+Define the loss as the log-likehood of data, viz.
+-f(z) := \log p(\Xi | z).
+The ELBO is then given by
+(1) \log p(\Xi)
+        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
+        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z))
+and we typically maximize the right hand side wrt x.
+Let us emphasize that in variational inference, x are the parameters and z are random variables.
+Also p(z) is a fixed (parameter-free) prior on the weights z.
+
+As noted in our previous comment, while our approach is not variational in nature, we shall now aim to connect the ELBO with local entropy.
+Focus on the local entropy term defined in Eqn.
+4 in the paper: we have
+(2) \log F(x,\gamma)
+        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
+        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
+Eqns. (1) and (2) are thus very different objects. In general, there does not exist a direct mapping between local entropy and variational inference.
+
+We can however provide the following connected for a special case.
+Consider the family of flat variational distributions q_x(z) such that the distribution is uniform on some compact set around x, e.g., q_x(z) \propto c if |x-z| < C and zero otherwise.
+We will pick a special non-fixed prior p(z) \propto \exp(-\gamma/2 |x-z|^2).
+The expectation in Eqn.
+(1) can now be simplified as:
+(3) \log p(\Xi)
+        \geq \int_{z \in Z} [-f(z)] - KL(q_x(z) || p(z))
+        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
+The differential entropy entropy H(q_x(z)) is a constant wrt x. The other two terms are the same as Eqn. (2) (multiplied by a constant c).
+
+In other words "local entropy is similar to performing variational inference within a special class of flat distributions around the paramter x".
+Such an explanation also validates our claim to finding flat regions in the energy landscape.
+The above mapping is however only heuristic because the prior p(z) is not fixed and depends on x itself which is unusual.
+The gradient of local entropy clarifies this:
+    dF  = -\gamma (x - E_{x' \sim r(z; x)} [z]);
+where the distribution r(z; x) is
+    \log r(z; x) \propto \log p(\Xi | z)  - \gamma/2 |x-z|^2
+and is thus a likelihood and prior that "moves" along with current iteratre x itself.
+
+It is thus easy to see the differences between a variational approach on the weights and our method.
+At present we do not see a more rigorous way to reduce our method to SVI, but we would welcome specific suggestions if that is indeed the case.
+
+We have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the results to the paper.
+In summary, after a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
+Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz.
+0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD).
+For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa).
+To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.
+
+-------------------------------------------------------------------------
+
+You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".
+But in general the HVP can be computed for only twice the time cost as a gradient evaluation.
+Can you elaborate on why this is expensive in your case?
+
+The above comment is for HVP of a general deep network, not our algorithm.
+And indeed, you are right that HVP can be computed using forward differencing with two back-props [P94, CSP93].
+However, as [P94] mentions, it is a
+
+[CSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
+[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
+
+- Bregman vs. KL
+- On the use of stochastic hessian information in optimization methods for machine learning
+- https://goo.gl/qxWzgr

### 2016-12-29 doc/rebuttal.txt
first draft for SVI-local-entropy argument

--- 2016-12-22 doc/rebuttal.txt
@@
 [CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient mcmc and stochastic optimization." arXiv:1512.07962 (2015).
+
+----------------------------------------------------------------------
+
+Allow us to elaborate further.
+
+Let \Xi denote the dataset, z denote the weights and x be the parameters of a variational distribution on weights denoted as q_x(z).
+Define the loss to be the negative log-likelihood of data, viz.
+-f(z) := \log p(\Xi | z).
+The ELBO is then given by
+(1) \log p(\Xi)
+        \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
+        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
+we maximize the right hand side with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.
+
+We shall now try to connect the ELBO with local entropy. In our paper, we maximize the local entropy (Eqn. 4 in the paper) or equivalently, its logarithm:
+(2) \log F(x,\gamma)
+        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
+        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
+The inequality follows by an application of Jensen's inequality.
+As noted in our previous comment, our approach is not variational in nature and it is easy to see that Eqns.
+(1) and (2) are quite different.
+In general, there does not seem to be direct reduction from local entropy to variational inference.
+
+We can however provide the following connection in a special case.
+Consider a family of flat variational distributions q_x(z) such that the distribution has a mean x and it is uniform on a compact set, for instance, let q_x(z) \propto c if |x-z| < C and zero otherwise.
+We will also pick a special prior p(z) \propto \exp(-\gamma/2 |x-z|^2).
+The expectation in Eqn.
+(1) can now be simplified as:
+(3) \log p(\Xi)
+        \geq \int_{z \in Z} [-f(z)] - KL(q_x(z) || p(z))
+        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
+The differential entropy H(q_x(z)) is a constant with respect to x. For the purposes of maximization, ELBO as given in Eqn. (3) is thus equivalent to local entropy (upto an additive constant).
+
+In other words, "local entropy is similar to performing variational inference within a class of uniform distributions with mean x".
+Such an explanation also validates our motivation to find flat regions in the energy landscape.
+The above discussion is however not entirely rigorous because the prior p(z) is not fixed and depends on the parameters x themselves.
+The gradient of local entropy (Eqn.
+7 in the paper) clarifies this point:
+    dF  = -\gamma (x - E_{z \sim r(z; x)} [z]);
+where the distribution r(z; x) is
+    \log r(z; x) \propto \log p(\Xi | z)  - \gamma/2 |x-z|^2
+and is thus a data likelihood term with a prior that "moves" along with current iterate x.
+
+At present we do not see a more rigorous way to reduce our method to SVI but we would welcome specific suggestions if that is indeed the case. We will update the paper with the above discussion.
+
+In response to your previous comments, we have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the following results to the paper.
+After a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
+Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz.
+0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD).
+For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa).
+To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.
+
+[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
+-------------------------------------------------------------------------
+
+You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".
+But in general the HVP can be computed for only twice the time cost as a gradient evaluation.
+Can you elaborate on why this is expensive in your case?
+
+The above comment is for HVP of the loss function of a general deep network, not local entropy.
+Indeed, HVPs can be computed using forward differencing with two back-props.
+However, as [P94] mentions, such an approximation is highly susceptible to numerical errors --- especially in high dimensions [P94, M10].
+Moreover, estimating HVP on a single mini-batch is not enough, as [CSP93] mention one typically has to average the HVP over many mini-batches; this is expensive.
+For instance, the authors in [CSP93] average over 10-20 batches for estimating the largest eigenvalue of the Hessian in a network with ~60,000 weights.
+[P94] offers an accurate, although more complex, algorithm to compute HVPs using the R-operator.
+
+[CSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
+[P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
+[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML. 2010.
+
+- On the use of stochastic hessian information in optimization methods for machine learning
+- https://goo.gl/qxWzgr
+- http://icml2010.haifa.il.ibm.com/papers/458.pdf

### 2017-01-02 doc/rebuttal.txt
wip on rebuttal

@@
         = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
-we maximize the right hand side with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.
+where the right hand side would be maximized with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.
 
@@
 The inequality follows by an application of Jensen's inequality.
-As noted in our previous comment, our approach is not variational in nature and it is easy to see that Eqns.
+As noted in our previous comment, our approach is not variational in nature and indeed, Eqns.
 (1) and (2) are quite different.
@@
 The expectation in Eqn.
-(1) can now be simplified as:
+(2) can now be simplified as:
 (3) \log p(\Xi)
-        \geq \int_{z \in Z} [-f(z)] - KL(q_x(z) || p(z))
+        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z))
         = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
@@
 The above comment is for HVP of the loss function of a general deep network, not local entropy.
-Indeed, HVPs can be computed using forward differencing with two back-props.
-However, as [P94] mentions, such an approximation is highly susceptible to numerical errors --- especially in high dimensions [P94, M10].
-Moreover, estimating HVP on a single mini-batch is not enough, as [CSP93] mention one typically has to average the HVP over many mini-batches; this is expensive.
-For instance, the authors in [CSP93] average over 10-20 batches for estimating the largest eigenvalue of the Hessian in a network with ~60,000 weights.
-[P94] offers an accurate, although more complex, algorithm to compute HVPs using the R-operator.
+Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10].
+Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10].
+Moreover, one typically has to average the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x SGD _per_ HVP).
+More accurate algorithms like that of [P94] also require this averaging over samples.
 
-[CSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
+[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
 [P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
 [M10] Martens, James. "Deep learning via Hessian-free optimization." ICML. 2010.
-
-- On the use of stochastic hessian information in optimization methods for machine learning
-- https://goo.gl/qxWzgr
-- http://icml2010.haifa.il.ibm.com/papers/458.pdf

### 2017-01-02 doc/rebuttal.txt
rebuttal v1.0

@@
 
-Allow us to elaborate further.
+Title: You cannot do SVI on the network weights unless you choose a “moving prior”, a feature of our scheme.
 
-Let \Xi denote the dataset, z denote the weights and x be the parameters of a variational distribution on weights denoted as q_x(z).
-Define the loss to be the negative log-likelihood of data, viz.
--f(z) := \log p(\Xi | z).
-The ELBO is then given by
-(1) \log p(\Xi)
+To address the objection 1) let \Xi denote the dataset, z  the weights and x the parameters of a variational distribution q_x(z). The ELBO can then be written as
+(i) \log p(\Xi)
         \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
         = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
-where the right hand side would be maximized with respect to x. The distribution p(z) is a fixed (parameter-free) prior on the weights.
+and maximized wrt x. The distribution p(z) is a fixed (parameter-free) prior, which one has to postulate.
 
-We shall now try to connect the ELBO with local entropy. In our paper, we maximize the local entropy (Eqn. 4 in the paper) or equivalently, its logarithm:
-(2) \log F(x,\gamma)
-        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
-        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz.
-The inequality follows by an application of Jensen's inequality.
-As noted in our previous comment, our approach is not variational in nature and indeed, Eqns.
-(1) and (2) are quite different.
-In general, there does not seem to be direct reduction from local entropy to variational inference.
+On the other hand, Eqn. 4 in the paper can be used to write the log of local entropy, where f = -\logp(\Xi | z):
+(ii) \log F(x,\gamma)
+        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz.
+We are unaware of any way to choose a prior p(z) and a variational family q_x(z) that that makes (i) resemble (ii), and therefore interpret our method as “integrating out […] the posterior over neural network parameters.” The only way we can do so is to pick a specific “prior" that depends on the parameter x (hence, not really a prior).
+For instance, one could choose a uniform variational family and a Gaussian “prior" with mean x to make (ii) resemble ELBO.
+In this case p(z) would not be fixed, but it would “move” along with the current iterate x.
+This "moving prior" is a crucial feature of our proposed algorithm.
 
-We can however provide the following connection in a special case.
-Consider a family of flat variational distributions q_x(z) such that the distribution has a mean x and it is uniform on a compact set, for instance, let q_x(z) \propto c if |x-z| < C and zero otherwise.
-We will also pick a special prior p(z) \propto \exp(-\gamma/2 |x-z|^2).
-The expectation in Eqn.
-(2) can now be simplified as:
-(3) \log p(\Xi)
-        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z))
-        = -\int_{z \in Z} c*f(z) dz + H(q_x(z)) - \int_{z \in Z} c*\gamma/2 |x-z|^2 dz;
-The differential entropy H(q_x(z)) is a constant with respect to x. For the purposes of maximization, ELBO as given in Eqn. (3) is thus equivalent to local entropy (upto an additive constant).
+Concerning the relation to SGLD, consider Belief Propagation (BP).
+Our proposed algorithm relates to the "focusing-Belief Propagation" variant (fBP), rather than the standard one [BBC16].
+The difference between BP and fBP is analogous to that between SLGD and entropy-SGD: the latter operates on a transformation of the landscape of the former, exploiting local entropic effects.
+This difference is crucial and indeed related to the "moving prior" of the previous discussion; plain SGLD (or BP) can only trade accuracy for entropy via the temperature parameter, which does not allow to directly use any geometric information on the landscape and does not help with narrow minima.
 
-In other words, "local entropy is similar to performing variational inference within a class of uniform distributions with mean x".
-Such an explanation also validates our motivation to find flat regions in the energy landscape.
-The above discussion is however not entirely rigorous because the prior p(z) is not fixed and depends on the parameters x themselves.
-The gradient of local entropy (Eqn.
-7 in the paper) clarifies this point:
-    dF  = -\gamma (x - E_{z \sim r(z; x)} [z]);
-where the distribution r(z; x) is
-    \log r(z; x) \propto \log p(\Xi | z)  - \gamma/2 |x-z|^2
-and is thus a data likelihood term with a prior that "moves" along with current iterate x.
+In view of your comment 2), we also implemented SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and will add the following results to our paper: After a hyper-parameter search, the best rwe obtained were a test error of LeNet on 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
+Disregarding the slow convergence of SGLD, its generalization error is slightly worse than the results in our paper, viz.
+0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD).
+For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (0.47% with Santa), there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.
 
-At present we do not see a more rigorous way to reduce our method to SVI but we would welcome specific suggestions if that is indeed the case. We will update the paper with the above discussion.
+[BBC16] Baldassi, Carlo, et al. "Unreasonable effectiveness of learning neural networks: From accessible states and robust ensembles to basic algorithmic schemes." PNAS (2016).
+[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
 
-In response to your previous comments, we have tested SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and can add the following results to the paper.
-After a hyper-parameter search, the best results we obtained yielded a test error of LeNet of 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
-Disregarding the slow convergence of SGLD, its generalization error is worse than the results in our paper, viz.
-0.48% with Entropy-SGD on LeNet (0.51% with SGD) and 8.65% with Entropy-SGD on All-CNN-BN (8.3% with SGD).
-For comparison, the authors in [CCF15] report an error of 0.71% with SGLD on MNIST with a slightly larger network (and get 0.47% error with Santa).
-To our knowledge, there are no results in the literature where MCMC methods perform comparably to SGD on larger networks.
-
-[CCF15] Chen, Changyou, et al. "Bridging the gap between stochastic gradient MCMC and stochastic optimization." arXiv:1512.07962 (2015).
 -------------------------------------------------------------------------
 
-You claim that "For modern deep networks with a few million weights, computing the Hessian-vector product as done in HS97 is prohibitive".
-But in general the HVP can be computed for only twice the time cost as a gradient evaluation.
-Can you elaborate on why this is expensive in your case?
+Title: Re: Is the Hessian-vector product actually expensive?
 
-The above comment is for HVP of the loss function of a general deep network, not local entropy.
+Comment 3) above is for HVP of the loss function of a general deep network, not local entropy.
 Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10].
 Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10].
-Moreover, one typically has to average the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x SGD _per_ HVP).
+One typically averages the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x one iteration of vanilla SGD).
 More accurate algorithms like that of [P94] also require this averaging over samples.
 
-[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS 1993.
+Practically speaking, one could argue that accuracy in the approximation of the HVP does not matter in SGD; what matters is only the local curvature at the scale of the perturbation, which needs to be chosen carefully.
+In this case, indeed, approximate computation of HVP can be considered cheap.
+
+[LSP93] LeCun, Yann, Patrice Y. Simard, and Barak Pearlmutter. "Automatic learning rate maximization by on-line estimation of the Hessian’s eigenvectors." NIPS (1993).
 [P94] Pearlmutter, Barak A. "Fast exact multiplication by the Hessian." Neural computation 6.1 (1994): 147-160.
-[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML. 2010.
+[M10] Martens, James. "Deep learning via Hessian-free optimization." ICML (2010).

### 2017-01-03 doc/rebuttal.txt
added overleaf as a submodule

@@
 
-Title: You cannot do SVI on the network weights unless you choose a “moving prior”, a feature of our scheme.
+Title: SVI does not resemble Entropy-SGD without a "moving prior", a feature of our scheme
 
-To address the objection 1) let \Xi denote the dataset, z  the weights and x the parameters of a variational distribution q_x(z). The ELBO can then be written as
+To address the objections 1) and 3) let \Xi denote the dataset, z denote the weights and x be the parameters of a variational distribution q_x(z). The ELBO can then be written as
 (i) \log p(\Xi)
         \geq E_{z \sim q_x(z)} [\log p(\Xi | z)] - KL(q_x(z) || p(z))
-        = E_{z \sim q_x(z)} [-f(z)] - KL(q_x(z) || p(z));
-and maximized wrt x. The distribution p(z) is a fixed (parameter-free) prior, which one has to postulate.
+and maximized with respect to x. The distribution p(z) is a fixed (parameter-free) prior, which one has to postulate.
 
-On the other hand, Eqn. 4 in the paper can be used to write the log of local entropy, where f = -\logp(\Xi | z):
+On the other hand, Eqn. 4 in the paper can be used to write the log of local entropy as:
 (ii) \log F(x,\gamma)
-        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz.
-We are unaware of any way to choose a prior p(z) and a variational family q_x(z) that that makes (i) resemble (ii), and therefore interpret our method as “integrating out […] the posterior over neural network parameters.” The only way we can do so is to pick a specific “prior" that depends on the parameter x (hence, not really a prior).
-For instance, one could choose a uniform variational family and a Gaussian “prior" with mean x to make (ii) resemble ELBO.
+        = \log \int_{z \in Z} e^{-f(z; \Xi) - \gamma/2 |x-z|^2} dz
+        \geq \int_{z \in Z} [-f(z; \Xi) - \gamma/2 |x-z|^2] dz;
+where f(z) = -\log p(\Xi | z).
+We are unaware of a general way to choose a prior p(z) and a variational family q_x(z) that that makes (i) resemble (ii), and therefore interpret our method as “integrating out […] the posterior over neural network parameters.” The only way we can do so is to pick a specific “prior" that depends on the parameter x (hence, not really a prior).
+For instance, one could choose a uniform variational family (say, q_x(z) \propto constant for |x-z| \leq C and zero otherwise) and a Gaussian “prior" with mean x (\log p(z) = -\gamma/2 |x-z|^2) to make (ii) resemble ELBO.
 In this case p(z) would not be fixed, but it would “move” along with the current iterate x.
-This "moving prior" is a crucial feature of our proposed algorithm.
+
+This "moving prior" is a crucial feature of our proposed algorithm. The gradient of local entropy (Eqn. 7 in the paper) clarifies this further:
+    dF  = -\gamma (x - E_{z \sim r(z; x)} [z]);
+where the distribution r(z; x) is
+    r(z; x) \propto p(\Xi | z) \exp(-\gamma/2 |x-z|^2);
+i.e. it contains a data likelihood term with a prior that "moves" along with the current iterate x.
 
@@
 Our proposed algorithm relates to the "focusing-Belief Propagation" variant (fBP), rather than the standard one [BBC16].
-The difference between BP and fBP is analogous to that between SLGD and entropy-SGD: the latter operates on a transformation of the landscape of the former, exploiting local entropic effects.
-This difference is crucial and indeed related to the "moving prior" of the previous discussion; plain SGLD (or BP) can only trade accuracy for entropy via the temperature parameter, which does not allow to directly use any geometric information on the landscape and does not help with narrow minima.
+The difference between BP and fBP is analogous to that between SGLD and Entropy-SGD: the latter operates on a transformation of the energy landscape of the former, exploiting local entropic effects.
+This difference is crucial and indeed related to the "moving prior" of the previous discussion; plain SGLD (or BP) can only trade energy for entropy via the temperature parameter which does not allow for direct use of any geometric information of the landscape and does not help with narrow minima.
 
-In view of your comment 2), we also implemented SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and will add the following results to our paper: After a hyper-parameter search, the best rwe obtained were a test error of LeNet on 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
+In view of your comment 2), we also implemented SGLD for LeNet on MNIST and All-CNN-BN on CIFAR and will add the following results to our paper: After a hyper-parameter search, the best we obtained were a test error of LeNet on 0.63 \pm 0.1% on MNIST after 300 epochs and 9.89 \pm 0.11% on All-CNN-BN after 500 epochs.
 Disregarding the slow convergence of SGLD, its generalization error is slightly worse than the results in our paper, viz.
@@
 Comment 3) above is for HVP of the loss function of a general deep network, not local entropy.
-Indeed, HVPs can be computed using forward differencing with two back-props and Hessian-free methods do exploit this, although to a limited degree of success [M10].
+Indeed, HVPs can be computed using forward differencing with two back-props.
 Such an approximation is susceptible to numerical errors --- especially in high dimensions [P94, M10].
-One typically averages the HVP over many samples in the dataset which is expensive (for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x one iteration of vanilla SGD).
+One typically averages the HVP over many samples in the dataset which is expensive, for instance, the authors in [LSP93] average over a few hundred samples, which roughly translates to 5-10x the time required for one iteration of vanilla SGD.
 More accurate algorithms like that of [P94] also require this averaging over samples.
 
-Practically speaking, one could argue that accuracy in the approximation of the HVP does not matter in SGD; what matters is only the local curvature at the scale of the perturbation, which needs to be chosen carefully.
-In this case, indeed, approximate computation of HVP can be considered cheap.
+One could argue that accuracy in the approximation of HVP does not matter in practice for purposes of training; what matters is only the local curvature at a scale commensurate with the typical weight updates.
+The perturbation vector for computing HVPs using back-props thus needs to be chosen carefully.
+If so, indeed, approximate computation of HVP can be considered cheap.
 

### 2017-01-04 dnn/hyperoptim.py
trying SGLD step size = 0.1, L = 10

@@ -1,7 +1,7 @@
 import os, sys, subprocess, json, argparse
 from itertools import product
 
-parser = argparse.ArgumentParser(description='Quick dirty jobber')
+parser = argparse.ArgumentParser(description='Quick dirty hyperoptim')
 parser.add_argument('-c','--command',   help='Main command', type=str, required=True)
 parser.add_argument('-p','--params',    help='JSON dict of the hyper-parameters', type=str)
 parser.add_argument('-r', '--run',      help='run',  action='store_true')

### 2017-01-06 dnn/ploterrors.py
added new allcnn plots

@@ -33,97 +33,95 @@ plt.rc('legend', fontsize=fsz)
 plt.rc('figure', titlesize=fsz)
 
 smooth = lambda x,c=com: pd.ewma(x, c)
-loss = lambda df, m: df[(df['epoch'] <= m) & (df['tv'] == 1)]['loss'].iloc[::loss_subsample]
+loss = lambda df, m: df[(df['epoch'] <= m) & (df['tv'] == 1) & (df['batch'] == 2)].loss
 train = lambda df, m: df[(df['epoch'] <= m) & (df['tv'] == 1) & (df['batch'] == 0)].miss
 valid = lambda df, m: df[(df['epoch'] <= m) & (df['tv'] == 0) & (df['batch'] == 0)].miss
 
-def load(fs1, fs2, m=50):
-    dfs = []
-    r = {'loss': [], 'train': [], 'valid': []}
-    for fs in [fs1, fs2]:
-        Ds = []
-        for f in fs:
-            Ds.append(pd.read_csv(f, sep=None, engine='python'))
-        df = pd.concat(Ds, keys=[i for i in xrange(len(Ds))])
-        dfs.append(df)
+def load(fs, m=50):
+    r = {}
+    Ds = []
+    for f in fs:
+        Ds.append(pd.read_csv(f, sep=None, engine='python'))
+    df = pd.concat(Ds, keys=[i for i in xrange(len(Ds))])
 
-        r['loss'].append(loss(df, m))
-        r['train'].append(train(df, m))
-        r['valid'].append(valid(df, m))
+    r['loss'] = loss(df, m)
+    r['train'] = train(df, m)
+    r['valid'] = valid(df, m)
+
     return r
 
-def plot_lenet():
-    r = load(sorted(glob.glob('../../results/dnn/lenet/*\"langevin\":0*.log')), \
-        sorted(glob.glob('../../results/dnn/lenet/*\"langevin\":20*.log')), 100)
+# def plot_lenet():
+#     r = load(sorted(glob.glob('../results/nov_expts/dnn/lenet/*\"langevin\":0*.log')), \
+#         sorted(glob.glob('../results/dnn/nov_expts/lenet/*\"langevin\":20*.log')), 100)
 
-    fig = plt.figure(1, figsize=(8,7))
-    plt.clf()
-    ax = fig.add_subplot(111)
-    plt.title(r'LeNet: Validation error')
[…]

### 2017-01-06 doc/rebuttal.txt
putting current overleaf version in git

@@
 [M10] Martens, James. "Deep learning via Hessian-free optimization." ICML (2010).
+
+-----------------------------------------------------------------------
+
+REBUTTAL
+
+We thank the Reviewers and others for their valuable feedback.
+We incorporated all specific comments into our current draft.
+We respond below to all reviewers with a general comment (all reviews and comments are also individually addressed by us):
+a) Smoothing the original loss function vs. local entropy
+- smoothing does not bias towards wide minima (e.g., many sharp minima that are close by will be unaffected), it might reduce the ruggedness of the landscape
+- Refer to Fig. 1, the local-entropy landscape is indeed smoother, but more importantly, it has a global minimum in a wide valley.
+- smoothing = local entropy only for a large \gamma
+
+b) [comment about the speed]
+
+c) eigenvalues
+[fix] Our generalization bound in Thm.
+3 relies upon an assumption that requires the Hessian to not have eigenvalues in the set [-2 \gamma-c, c], where the constant c depends on both the dataset and the architecture; this is unfortunately quite unnatural.
+Our argument is indeed only a sketch and we will clarify this.
+It also uses Laplace's method to show that the modified loss function is smoother; this is valid only for large values of \gamma.
+We would however like to point out that one requires the bound to hold only during the optimization and not everywhere in the landscape (as Fig.
+1 shows, this assumption is not true at the ~local minima themselves).
+A rigorous proof of such a statement however requires an analysis of the dynamics of SGD and is currently out of reach.
+
+------------------
+AnonReviewer4:
+
+(1) The confusion about the two claims is probably caused by the wording of our remark, which is a statement about the energy landscape based on our experimental results.
+We always get an equivalent or better generalization error using Entropy-SGD and a lower cross-entropy loss on the _training set_ (Fig.
+5a).
+This suggests that there are wide local minima at lower energy levels in the energy landscape which Entropy-SGD manages to find.
+
+(2) Please see the updates above for additional experiments that demonstrate acceleration in wall-clock time with respect to SGD.
+
+(3) Please see the latest updates to the paper. We can now train all our networks with only the local entropy loss, i.e., rho = 0, using scoping on the parameter \gamma.
+
+(4) [fix] We have expanded the discussion on [HS97] in Sec.
+2 to dwell upon the a few similarities.
+We agree that the Gibbs formalism to prove improved generalization of flat minima is a promising direction for further analysis.
+The Gibbs variant to averaging in weight space (Eqn.
+33, pg.
+22 of [HS97]) is similar to the averaging in Eqn.
+7 of our paper.
+
+------------------
+Csaba Szepesvari:
+
+__Related work__: We have expanded the discussion regarding [Baldassi et. al. '16] in Sec. 2 and Sec. 3.
+
+------------------
+AnonReviewer1:
+
+Thanks, it should indeed be f(x') in Eqn. 8.
+
+We have significantly expanded the experimental section and also added an experiment that trains an LSTM on the Penn Tree Bank corpus. Please see updates to the paper for more details.
+
+------------------
+AnonReviewer2:
+
+4. We set c from Thm. 3.12 of Hardt's paper to be 1 for clarity (c is the initial learning rate). Including it in the exponent does not change our analysis.
+
+------------------
+Areachair1 [Stefano should maybe edit and post that comment]:
+
+Re: You should make these connections clearer and more explicit
+We have discussed this in depth with sufficient technical details above.
+Using terms like "similar to motivations of Bayesian methods" and "using Bayesian methods in the inner loop" is generalizing to the point of absurdity.
+
+It is not accurate to term local entropy as "SVI with a moving prior" for a number of reasons.
+"SVI with a moving prior" has no theoretical basis.
+It is not connected to the log-likelihood of data, it is connected to local entropy instead.
+As we have discussed above, local entropy *resembles* SVI with a moving prior but only in an extremely non-rigorous and merely conceptual fashion.
+
+We have updated the experiments in our paper.
+We now plot the number of gradient evaluations on the x-axis and using scoping, we can in fact show that local entropy performs better in a smaller amount of time.
+As requested, we have also added a section comparing local-entropy with SGLD.
+
+Re: Exact Hessian-vector product is easy to compute
+We have discussed this above (see the comment regarding [P94]). Computing the "exact" HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch, which is expensive.

### 2017-01-06 doc/chaudhari.choromanska.ea.iclr17.tex
putting current overleaf version in git

@@
 
-\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$\\[0.05in]
-$^{1}$ Computer Science, University of California, Los Angeles\\
-$^{2}$ Courant Institute of Mathematical Sciences, New York University\\
-$^{3}$ Facebook AI Research\\[0.05in]
+\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, Carlo Baldassi$^{4}$,\\[0.03in]
+\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, Riccardo Zecchina$^{4}$}\\[0.05in]
+$^{1}$ Computer Science, University of California, Los Angeles.\\
+$^{2}$ Courant Institute of Mathematical Sciences, New York University.\\
+$^{3}$ Facebook AI Research, New York.\\
+$^{4}$ Politecnico di Torino.\\
+$^{5}$ Microsoft Research New England, Cambridge.\\ [0.05in]
 {\footnotesize
@@
 \href{mailto:soatto@cs.ucla.edu}{soatto@cs.ucla.edu},
-\href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu}}
-}
+\href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu},
+\href{mailto:carlobaldassi@gmail.com}{carlobaldassi@gmail.com}}\\[0.03in]
+{\footnotesize
+\hspace{0.33in} \href{mailto:borgs@microsoft.com}{borgs@microsoft.com},
+\href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
+\href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
+}}
 
@@
 \setlength{\marginparwidth}{1in}
-\newcommand{\pc}[2]{{\color{ForestGreen} #1}\marginpar{\tiny\noindent{\raggedright{\color{Sienna}[PC]}\color{Sienna}{ #2} \par}}}
-\renewcommand{\ss}[2]{{\color{cyan} #1}\marginpar{\tiny\noindent{\raggedright{\color{BurntOrange}[SS]}\color{BurntOrange}{ #2} \par}}}
-\newcommand{\ac}[2]{{\color{magenta} #1}\marginpar{\tiny\noindent{\raggedright{\color{magenta}[AC]}\color{magenta}{ #2} \par}}}
+\newcommand{\pc}[2]{{\color{ForestGreen}#1}\marginpar{\tiny\noindent{\raggedright{\color{Sienna}[PC]}\color{Sienna}{#2} \par}}}
+\renewcommand{\ss}[2]{{\color{cyan}#1}\marginpar{\tiny\noindent{\raggedright{\color{BurntOrange}[SS]}\color{BurntOrange}{#2} \par}}}
+\newcommand{\ac}[2]{{\color{magenta}#1}\marginpar{\tiny\noindent{\raggedright{\color{magenta}[AC]}\color{magenta}{#2} \par}}}
 \newcommand{\todo}[1]{{\color{gray}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[TODO]}}}}
@@
 \begin{abstract}
-This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape at solutions found by gradient descent.
-Local extrema with low generalization error have a large proportion of almost-zero eigenvalues in the Hessian with very few positive or negative eigenvalues.
-We leverage upon this observation to construct a local entropy based objective that favors well-generalizable solutions lying in the flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
-%
-Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of local entropy at each update of the weights.
-%
-We prove that incorporating local entropy into the objective function results in a smoother energy landscape and use uniform stability to show improved generalization bounds over SGD.
-%
-Our experiments on competitive baselines demonstrate that $\entropysgd$ leads to improved generalization and has the potential to accelerate training.
+This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape.
+We construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
+Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
+We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD.
+Our experiments on competitive baselines demonstrate that $\entropysgd$ alone leads to generalization performances that compare with state of the art techniques and has the potential to accelerate training.
+\pc{}{We also perform a direct comparison between the information provided by the local entropy measure and the analysis of the Hessian in qualitatively different minima.}
 \end{abstract}
@@
 \label{s:intro}
+This paper presents a new optimization tool for deep learning, designed to exploit the local geometrical properties of the objective function.
+Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
+These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
+The original analysis was done in the case of binary weights, and was later generalized by~\citet{baldassi2016multilevel} to the case of weights with an arbitrary number of precision bits.
+Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.
 
-This paper focuses on developing new optimization tools for deep learning that are tailored to exploit the local geometric properties of the objective function.
-Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the eigenspectrum of the Hessian at a local minimum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian.
+Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the spectrum of the Hessian at the endpoint of the optimization performed by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 Interestingly, this trend is not unique to this particular network.
-Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (refer to Sec.~\ref{s:expt} for more experiments).
-Local minima that generalize well and are discovered by gradient descent lie in "wide valleys" of the energy landscape, that look more like the bottom of a bucket than sharp isolated minima like the holes on a golf-course.
-%For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
-Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
+Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (see references in Sec.~\ref{s:prior_work}, and Sec.~\ref{s:expt} for more experiments).
+\pc{Local minima}{We are using the term ``local minima'' loosely here, since a few of the eigenvalues of the Hessian may be slightly negative} that generalize well and are discovered by gradient descent lie in "wide valleys" of the energy landscape, rather than in sharp isolated minima.
+For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
+Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.
 
-\begin{figure}[!tbh]
-\vspace*{-0.1in}
+\begin{figure}[tbh]
+% \vspace*{-0.1in}
 \centering
@@
 Based on this understanding of how the local geometry looks at the end of optimization, can we modify SGD to actively seek such regions?
-We show that flat regions in the energy landscape correspond to regions where the log-partition function is small, if computed locally on the Gibbs distribution corresponding to the optimization problem.
-Roughly, for parameters denoted by $x \in \reals^n$, we modify a loss function $f(x)$ to be
+Motivated by the work of~\citet{baldassi2015subdominant,baldassi2016unreasonable} on shallow networks, we propose to replace the original loss function $f(x)$ by:
 $$	
-	f(x) - \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx';
+% \rho
+f(x) 
+- \log \int_{x' \in \reals^n}\ \exp \rbrac{-\b f(x') - \f{\b \g}{2}\ \norm{x-x'}^2_2}\ dx'.
 $$
-where the second term is a log-partition function which is seen as a ``local free energy'' and measures both the depth of a valley at a location $x$ and its flatness through the entropy of $f(x')$.
+The second term measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$;  we call it the {\it local free entropy} in analogy to the free entropy in statistical physics.
 The $\entropysgd$ algorithm presented in this paper employs a few iterations of stochastic gradient Langevin dynamics (SGLD) to approximate the second term before each update of the parameters using the back-propagated gradient.
@@
 Actively biasing towards wide valleys aids generalization, in fact, we can optimize solely the free energy term to obtain similar generalization error as SGD on the original loss function.
-Our experiments in Sec.~\ref{s:expt} on MNIST and CIFAR-10~\citep{krizhevsky2009learning} show that $\entropysgd$ scales to deep networks used in practice and in terms of the number of updates to the parameters, trains much quicker than SGD.
+Our experiments in Sec.~\ref{s:expt} on MNIST and CIFAR-10~\citep{krizhevsky2009learning} show that $\entropysgd$ scales to deep networks used in practice and in terms of the number of updates to the parameters, trains much more quickly than SGD.
 
@@
 
-Our above observation about the eigenspectrum of Hessian (further discussed in Sec.~\ref{s:expt}) is similar to results on a perceptron model in~\citet{dauphin2014identifying} where the authors connect the loss function of a deep network to a high-dimensional Gaussian random field.
+Our above observation about the spectrum of Hessian (further discussed in Sec.~\ref{s:expt}) is similar to results on a perceptron model in~\citet{dauphin2014identifying} where the authors connect the loss function of a deep network to a high-dimensional Gaussian random field.
 They also relate to earlier studies such as~\citet{Baldi:1989:NNP:70359.70362,Fyodorov2007,Bray2007} which show that critical points with high training error are exponentially likely to be saddle points with many negative directions and all local minima are likely to have error that is very close to that of the global minimum.
@@
 
-The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
+The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing,sagun2014explorations} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
 Although assumptions in these works are somewhat unrealistic in the context of deep networks used in practice, similar results are also true for linear networks which afford a more thorough analytical treatment~\citep{DBLP:journals/corr/SaxeMG13}.
@@
 
-Such a paradox between a seemingly benign energy landscape and delicate training algorithms is also seen in problems such as random $k$-SAT~\citep{achlioptas2006solution} and compressed sensing~\citep{krzakala2012statistical}.
-Introducing non-local terms in classical inference techniques for these problems has led to state-of-the-art algorithms~\citep{braunstein2005survey,maneva2007new}.
-Closer to deep learning, although motivated from similar ideas, the authors in~\citet{baldassi2015subdominant,baldassi2016unreasonable} show that, both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average.
+Closer to the subject of this work,~\citet{baldassi2015subdominant,baldassi2016unreasonable} show that, both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average.
 Using a novel out-of-equilibrium analysis, the authors show that such an ensemble recovers dense clusters in the solution space.
-Such dense clusters generalize better than isolated local minima that may lie close to the global minimum.
-A related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and encourages individual models to explore by modulating their distance to the ensemble average; this is a special case of robust ensembles.
+Such dense clusters generalize better than isolated local minima.
+A related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and encourages individual models to explore by modulating their distance to the ensemble average; this is a special case of robust ensembles, as introduced in~\citet{baldassi2016unreasonable}.
 
@@
 The enabling technique in our case is that we approximate the local entropy using Langevin dynamics.
-\citet{baldassi2016unreasonable} instead employ belief propagation to do so; this can only be done for simpler perceptron models considered there.
+\citet{baldassi2016local} instead employ belief propagation to do so, but it's not clear whether this scales well beyond the shallow networks considered there.
 
 The authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
-However, as they mention, the effect of hyper-parameters for these constraints is intricately tied together and they are difficult to choose even for small problems.
+\pc{Their analysis of the generalization performance (see their Sec.
+A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
 Their formulation depends on Hessian-vector products.
@@
 
-\section{Local entropy}
+\section{Local free entropy}
 \label{s:local_entropy}
 
-We first provide a simple intuition for the concept of local entropy of an energy landscape.
+We first provide a simple intuition for the concept of local free entropy of an energy landscape.
 Consider a cartoon energy landscape in Fig.~\ref{fig:scoping} where the x-axis denotes the configuration space of the parameters.
-We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\textrm{non-robust}}$.
-Under a Bayesian prior on the parameters, say a Gaussian of a fixed variance at locations $x_{\textrm{robust}}$ and $x_{\textrm{non-robust}}$ respectively, the wider local minimum has a higher log-likelihood than the sharp valley on the right.
-In discrete optimization problems, such wide local minima are dense clusters of solutions in the parameter space where nearby parameters have similar energy, say, the number of clauses violated in $k$-SAT~\citep{sipser2006introduction,achlioptas2006solution}.
-On the other hand, sharp local minima are isolated solutions that are far away from other solutions.
+We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\trm{non-robust}}$.
+Under a Bayesian prior on the parameters, say a Gaussian of a fixed variance at locations $x_{\trm{robust}}$ and $x_{\trm{non-robust}}$ respectively, the wider local minimum has a higher log-likelihood than the sharp valley on the right.
+%[This is incorrect -- k-SAT has different behavior.] In discrete optimization problems, such wide local minima are dense clusters of solutions in the parameter space where nearby parameters have similar energy, say, the number of clauses violated in $k$-SAT~\citep{sipser2006introduction,achlioptas2006solution}.
+\pc{On the other hand, sharp local minima are isolated solutions that are far away from other solutions.}{cite Baldassi in this section}
 
@@
 The above discussion suggests that parameters that lie in wider local minima like $x_{\trm{robust}}$, which may possibly have a higher loss than the global minimum, should generalize better than the ones that are simply at the global minimum.
-This motivates us to use a very well-known quantity from thermodynamics known as ``free energy'' (also sometimes called ``free entropy''):
-$$
-    \textrm{free energy} = \textrm{internal energy} - \textrm{temperature} \times \textrm{entropy}.
-$$
-Roughly, internal energy is the average energy of all configurations while entropy counts the number of distinct configurations in a physical system at a given temperature.
-In a neighborhood of $x_{\textrm{robost}}$, the entropy term term is small while internal energy is large.
-Conversely, near $x_{\textrm{non-robust}}$, internal energy is low while entropy is high.
-The ``negative free entropy'' computed locally near a particular configuration thus provides a way of picking large flat regions of landscape over sharp, narrow valleys in spite of the latter possibly having a lower loss.
-Quite conveniently, free energy is also the log-partition function of a system.
+This motivates us to maximize the ``local free entropy'' term presented in the introduction: In a neighborhood of $x_{\trm{robust}}$, that term is large because it includes the contributions from a whole region of good configurations; conversely, in the region near $x_{\trm{non-robust}}$ there are no such contributions except for the minimum itself and the resulting local free entropy is low.
+The local free entropy thus provides a way of picking large,
+approximately flat, regions of the landscape over sharp, narrow valleys in spite of the latter possibly having a lower loss.
+Quite conveniently, the local free entropy is also computed from the partition function with a local re-weighting term.
 
 Formally, for a parameter vector $x \in \reals^n$, consider a Gibbs distribution corresponding to a given energy landscape $f(x)$:
-\beq{
+\begin{equation}
     \P(x;\ \b) = Z_\b^{-1}\ \exp\ \rbrac{-\b\ f(x)};
     \label{eq:gibbs_original}
-}
-where $\b$ is known as the inverse temperature and $Z_\b$ is a normalizing constant. As $\b \to \infty$, the probability distribution above concentrates on the global minimum of $f(x)$ given as
-\beq{
+\end{equation}
+where $\b$ is known as the inverse temperature and $Z_\b$ is a normalizing constant, also known 
+as the partition function in physics. As $\b \to \infty$, the probability distribution above concentrates on the global minimum of $f(x)$ (assuming it's unique) given as:
+\begin{equation}
     x^* = \argmin_x\ f(x),
     \label{eq:argmin_generic}
-}
+\end{equation}
 which establishes the link between the Gibbs distribution and a generic optimization problem~\eqref{eq:argmin_generic}.
-We would instead like the probability distribution --- and therefore the underlying optimization problem --- to focus on flat regions such as $x_{\textrm{robust}}$.
+We would instead like the probability distribution --- and therefore the underlying optimization problem --- to focus on flat regions such as $x_{\textrm{robust}}$, as in the figure.
 With this in mind, let us construct a modified Gibbs distribution:
-\beq{
+\begin{equation}
     \P(x';\ x, \b, \g) = Z_{x,\b,\ \g}^{-1}\ \exp\ \rbrac{-\b\ f(x') - \b\ \f{\g}{2}\ \norm{x-x'}^2_2}.
     \label{eq:gibbs_modified}
-}
+\end{equation}
 The distribution in~\eqref{eq:gibbs_modified} is a function of a dummy variable $x'$ and is parameterized by the original location $x$.
@@
 For small values of $\g$, the term $f(x')$ in the exponent dominates and the modified distribution is similar to the original Gibbs distribution in~\eqref{eq:gibbs_original}.
-We will set the inverse temperature $\b$ to $1$ because $\g$ affords us similar control on the Gibbs distribution.
+%We will set the inverse temperature $\b$ to $1$ because $\g$ affords us similar control on the Gibbs distribution.
 
-\begin{definition}[Local entropy]
+\begin{definition}[Local free entropy]
 \label{def:local_entropy}
-The local free energy of the Gibbs distribution in~\eqref{eq:gibbs_original}, colloquially called ``local entropy'' in the sequel and denoted by $F(x, \g)$, is defined as the log-partition function of modified Gibbs distribution~\eqref{eq:gibbs_modified}, i.e.,
-\aeq{
-    F(x, \g) &= \log Z_{x, 1, \g} \notag\\
-    &= \log \int_{x'}\ \exp \Big(-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2 \Big)\ dx'.
+The local free entropy of the Gibbs distribution in~\eqref{eq:gibbs_original}, colloquially called ``local entropy'' in the sequel and denoted by $F(x, \g)$, is defined as the log-partition function of modified Gibbs distribution~\eqref{eq:gibbs_modified}, i.e.,
+\begin{align}
+    F(x, \beta, \g) &= \log Z_{x, \beta, \g} \notag\\
+    &= \log \int_{x'}\ \exp \Big(-\beta f(x') - \f{\beta\g}{2}\ \norm{x-x'}^2_2 \Big)\ dx'.
     \label{eq:local_entropy}
-}
+\end{align}
 \end{definition}
 The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
-Fig.~\ref{fig:scoping} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
-Note that $-F(x,\g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.
+Fig.~\ref{fig:scoping} shows the negative local entropy $-F(x, \b,  \g)$ for the original energy landscape.
+Note that $-F(x, \b, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.
 
@@
 $
-    - \int_{x'}\ \log \P(x';\ x, \g)\ d \P(x';\ x, \g)
+    - \int_{x'}\ \log \P(x';\ x, \b, \g)\ d \P(x';\ x, \b, \g)
 $
@@
 The original optimization problem is:
-\beq{
+\begin{equation}
     x^* = \argmin_x\ \f{1}{N}\ \sum_{k=1}^N\ f(x;\ \xi_k);
     \label{eq:dnn_objective}
-}
+\end{equation}
 where the objective $f(x, \xi_k)$ is typically, a non-convex function in both the weights $x$ and the sample $\xi_k$. The $\entropysgd$ algorithm instead solves the problem
-\beq{
+\begin{equation}
     x^*_{\entropysgd} = \argmin_x\ \r\ \rbrac{\f{1}{N}\ \sum_{k=1}^N\ f(x;\ \xi_k)} - F(x, \gamma;\ \Xi);
     \label{eq:entropysgd_objective}
-}
+\end{equation}
 where we have made the dependence of local entropy $F(x,\g)$ on the dataset $\Xi$ explicit.
@@
 The gradient of local entropy over a randomly sampled mini-batch of $m$ samples denoted by $\xi_{\ell_i} \in \minibatch{\ell}$ for $i \leq m$ is easy to derive and is given by
-\beq{
+\begin{equation}
     \nabla_x F\rbrac{x, \g;\ \minibatch{\ell}} = -\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}};
     \label{eq:local_entropy_grad}
-}
+\end{equation}
 where the notation $\ag{\cdot}$ denotes an expectation of its arguments (we have again made the dependence on the data explicit) over a Gibbs distribution of the original optimization problem modified to focus on the neighborhood of $x$; this is given by
-\beq{
+\begin{equation}
     \P(x';\ x, \g) \propto\ \exp \sqbrac{-\rbrac{\f{1}{m}\ \sum_{i=1}^m\ f\rbrac{x;\ \xi_{\ell_i}}} - \f{\g}{2}\ \norm{x-x'}_2^2}.
     \label{eq:entropysgd_modified_gibbs}
-}
+\end{equation}
 Computationally, the gradient in~\eqref{eq:local_entropy_grad} involves estimating $\ag{x';\ \minibatch{\ell}}$ with the current weights fixed to $x$.
@@
 The gradient of the first term in~\eqref{eq:entropysgd_objective} is computed using back-propagation and the final gradient is
-\beq{
+\begin{equation}
     \r\ \rbrac{\f{1}{m}\ \sum_{i=1}^m\ \nabla f(x;\ \xi_{k_i})} + \g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}.
     \label{eq:entropysgd_grad}
-}
+\end{equation}
 Note that we use different mini-batches (denoted by $k$ and $\ell$ respectively) to compute the back-propagated gradient and gradient of the local entropy.
@@
 
-Our main result in this section is to show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
-Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).
+\pc{Our main result in this section is to show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
+Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).}{Clarify that the analysis here is only a sketch, it is not rigorous}
 
@@
 For an $\a$-Lipschitz and $\b$-smooth loss function, if SGD converges in $T$ iterations on $N$ samples with decreasing learning rate $\eta_t \leq 1/t$ the stability is bounded by
-\aeqs{
+\begin{align*}
     \e
@@
     &\lessapprox \f{1}{N}\ \a^{1/(1 + \b)}\ T^{1 - 1/(1 + \b)}.
-}
+\end{align*}
 \end{theorem}
@@
 Using Lem.~\ref{lem:smoothness_reduction} and Thm.~\ref{thm:generalization_bound_nonconvex} we have
-\beq{
+\begin{equation}
     \e_{\ \entropysgd} \lessapprox \rbrac{\a\ T^{-1}}^{\rbrac{1 - \f{1}{1 + \g^{-1}c}}\ \b}\ \e_{\ \textrm{SGD}},
     \label{eq:stability_bound}
-}
+\end{equation}
 which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$.
@@
 
-In our experiments, $\entropysgd$ results in a comparable generalization error as SGD, but always has a lower cross-entropy loss.
+In our experiments, $\entropysgd$ results in a comparable generalization error as SGD, but always has a lower cross-entropy loss on the training set.
 Roughly speaking, wide valleys favored by $\entropysgd$ seem to have lower empirical loss than local minima discovered by SGD.
@@
 For a parameter vector $x \in \reals^n$ with a prior distribution $p(x)$ and if the probability of generating a data item $\xi_k$ given a model parameterized by $x$ is $p(\xi_k \given x)$, the posterior distribution of the parameters based on $N$ data items can be written as
-\beq{
+\begin{equation}
     p\rbrac{x \given \xi_{\ k \leq N}}\ \propto\ p(x)\ \prod_{k=1}^N\ p\rbrac{\xi_k \given x}.
     \label{eq:bayesian_posterior}
-}
+\end{equation}
 Langevin dynamics~\citep{neal2011mcmc} injects Gaussian noise into maximum-a-posteriori (MAP) updates to prevent over-fitting the solution $x^*$ of the above equation. The updates can be written as
-\beq{
+\begin{equation}
     \Delta x_t = \f{\eta}{2}\ \rbrac{ \nabla \log p(x_t) + \sum_{k=1}^N\ \nabla p(\xi_k \given x_t)} + \sqrt{\eta}\ \e_t;
     \label{eq:langevin_bayesian}
-}
+\end{equation}
 where $\e_t \sim \trm{N}(0, 1)$ is Gaussian noise and $\eta$ is the learning rate.
@@
 However, as~\citet{welling2011bayesian} show, one can instead simply use the average gradient over $m$ data samples (mini-batch) as follows:
-\beq{
+\begin{equation}
     \Delta x_t = \f{\eta_t}{2}\ \rbrac{ \nabla \log p(x_t) + \f{N}{m}\ \sum_{k=1}^m\ \nabla p(\xi_k \given x_t)} + \sqrt{\eta_t}\ \e_t.
     \label{eq:sgld_bayesian}
-}
+\end{equation}
 Secondly, Langevin dynamics in~\eqref{eq:langevin_bayesian} is the discrete-time approximation of a continuous-time stochastic differential equation~\citep{mandt2016variational} thereby necessitating a Metropolis-Hastings (MH) rejection step~\citep{roberts2002langevin} which again requires computing $p(\xi_k \given x)$ over the entire dataset.
@@
 The gradient $-\nabla F(x)$ is computed in Sec.~\ref{ss:grad_local_entropy} to be $\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}$. Consider the term
-\aeqs{
+\begin{align*}
     x - \ag{x';\ x}
@@
     &= -Z_{x,\g}^{-1}\ e^{-f(x)}\ \int_{s}\ s\ e^{-\nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}\ ds.
-}
+\end{align*}
 The above expression is the mean of a distribution $\propto e^{-\nabla f(x)^\top s - \f{1}{2}\ s^\top \rbrac{\g + \nabla^2 f(x)} s}$.
@@
 Let us denote $A(x) := \rbrac{I + \g^{-1}\ \nabla^2 f(x)}^{-1}$. Plugging this into the condition for smoothness, we have
-\aeqs{
+\begin{align*}
     \norm{\nabla F(x, \g) - \nabla F(y, \g)}
@@
     &\leq \rbrac{\sup_x\ \norm{A(x)}}\ \b\ \norm{x-y}.
-}
+\end{align*}
 Unfortunately, we can only get a uniform bound if we assume that for a small constant $c > 0$, no eigenvalue of $\nabla^2 f(x)$ lies in the set $[-2 \g -c, c]$. This gives
@@
 For $J^2_{ijk},\ J^3_{ijk} \sim \trm{N}(0,1)$, the energy of a spin glass is given by
-\beq{
+\begin{equation}
     H(x) = \f{1}{10\ \sqrt{n}}\ \sum_{i,j=1}^n\ J^2_{ij}\ x_i\ x_j + \f{1}{n} \sum_{i,j,k=1}^n\ J^3_{ijk}\ x_i\ x_j\ x_k
     \label{eq:sg_hamiltonian}
-}
+\end{equation}
 where the parameter vector $x \in \reals^n$ is called the ``spin''.
@@
 We are interested in computing
-\beq{
+\begin{equation}
     x^*_H = \argmin_{x \in S^{n-1}(\sqrt{n})}\ H(x).
     \label{eq:3sg}
-}
+\end{equation}
 This is a non-convex problem and in general, hard; the number of local minima and saddle points of $H(x)$ scale exponentially in $n$~\citep{auffinger2013random}.
@@
 Our modified optimization problem with a negative local entropy term is given by
-\beq{
+\begin{equation}
     x^*_f = \argmin_{x \in S^{n-1}(\sqrt{n})}\ \rho H(x) - F(x, \g);
     \label{eq:3sg_modified}
-}
+\end{equation}
 with $\rho = 0.01$ and $\g = 10^{-4}$ and the local entropy $F(x, \g)$ is the log-partition function of

### 2017-01-06 doc/chaudhari.choromanska.ea.iclr17.tex

@@
 % notation goes here
+\newcommand\red[1]{{\color{red}#1}}
 \newcommand{\entropysgd}{\mathrm{Entropy}\textrm{-}\mathrm{SGD}}
@@
 
-\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, Carlo Baldassi$^{4}$,\\[0.03in]
-\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, Riccardo Zecchina$^{4}$}\\[0.05in]
+\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, \red{Carlo Baldassi$^{4}$},\\[0.03in]
+\red{\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
+Riccardo Zecchina$^{4}$}}\\[0.05in]
 $^{1}$ Computer Science, University of California, Los Angeles.\\
@@
 \href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
+\red{\href{mailto:leventsagun@gmail.com}{leventsagun@gmail.com},}
 \href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
@@
 This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape.
-We construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
-Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
+We construct a local-{\color{red}free-}entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
+Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local-{\color{red}free-}entropy at each update of the weights.
 We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD.
@@
 This paper presents a new optimization tool for deep learning, designed to exploit the local geometrical properties of the objective function.
-Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
+\red{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
 These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
 The original analysis was done in the case of binary weights, and was later generalized by~\citet{baldassi2016multilevel} to the case of weights with an arbitrary number of precision bits.
-Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.
+Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.}
 
-The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian.
-Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the spectrum of the Hessian at the endpoint of the optimization performed by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+\red{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.} Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the spectrum of the Hessian at the endpoint of the optimization performed by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (see references in Sec.~\ref{s:prior_work}, and Sec.~\ref{s:expt} for more experiments).
-\pc{Local minima}{We are using the term ``local minima'' loosely here, since a few of the eigenvalues of the Hessian may be slightly negative} that generalize well and are discovered by gradient descent lie in "wide valleys" of the energy landscape, rather than in sharp isolated minima.
+\pc{Local minima}{We are using the term ``local minima'' loosely here, since a few of the eigenvalues of the Hessian may be slightly negative} that generalize well and are discovered by gradient descent lie in ``wide valleys'' of the energy landscape, rather than in sharp isolated minima.
+\red{We are using the term ``local minima'' loosely here, since a few of the eigenvalues of the Hessian may be slightly negative and the norm of the gradient is slightly above zero.}
+
 For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
@@
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.
+\red{For a more detailed recent
+analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.}
 
@@
 Based on this understanding of how the local geometry looks at the end of optimization, can we modify SGD to actively seek such regions?
-Motivated by the work of~\citet{baldassi2015subdominant,baldassi2016unreasonable} on shallow networks, we propose to replace the original loss function $f(x)$ by:
+\red{Motivated by the work of~\citet{baldassi2015subdominant,baldassi2016unreasonable} on shallow networks,} we propose to replace the original loss function $f(x)$ by:
 $$	
@@
 $$
-The second term measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$;  we call it the {\it local free entropy} in analogy to the free entropy in statistical physics.
-The $\entropysgd$ algorithm presented in this paper employs a few iterations of stochastic gradient Langevin dynamics (SGLD) to approximate the second term before each update of the parameters using the back-propagated gradient.
+The second term measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$;  \red{we call it the {\it local free entropy} in analogy to the free entropy in statistical physics.} The $\entropysgd$ algorithm presented in this paper employs a few iterations of stochastic gradient Langevin dynamics (SGLD) to approximate the second term before each update of the parameters using the back-propagated gradient.
 At a conceptual level, this algorithm resembles running two nested loops of SGD.
@@
 
-\section{Local free entropy}
+\section{Local \red{free} entropy}
 \label{s:local_entropy}

### 2017-01-09 doc/chaudhari.choromanska.ea.iclr17.tex
commit before surgery

@@
 
-\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, \red{Carlo Baldassi$^{4}$},\\[0.03in]
-\red{\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
-Riccardo Zecchina$^{4}$}}\\[0.05in]
+\author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, Carlo Baldassi$^{4}$,\\[0.03in]
+\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
+Riccardo Zecchina$^{4}$}\\[0.05in]
 $^{1}$ Computer Science, University of California, Los Angeles.\\
@@
 \href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
-\red{\href{mailto:leventsagun@gmail.com}{leventsagun@gmail.com},}
 \href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
@@
 This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape.
-We construct a local-{\color{red}free-}entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
-Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local-{\color{red}free-}entropy at each update of the weights.
+We construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
+Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local-entropy at each update of the weights.
 We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD.

### 2017-01-09 dnn/plotlog.py

@@ -9,7 +9,7 @@ sns.set()
 
 whitelist = set(['seed','model','LR','noise', \
             'gamma','L','scoping','LRstep','rho','LRD', \
-            'langevin', 'langevin_noise'])
+            'langevin', 'langevin_noise','lr','input'])
 
 #colors = ["blue", "darkseagreen", "indianred", "purple", "sepia", "black"]
 #colors = sns.color_palette("muted")
@@ -87,8 +87,15 @@ def plot_idx(fs, df,
     #     'Training error')
     # plt.xlim([0,max_epochs])
 
+    # plt.figure(3)
+    # plt.clf()
+    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].miss,
+    #     'Test error')
+    # plt.xlim([0,max_epochs])
+
+    # lstm
     plt.figure(3)
     plt.clf()
-    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].miss,
-        'Test error')
+    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['iter'] == 0)].loss,
+        'Test loss')
     plt.xlim([0,max_epochs])

### 2017-01-09 doc/chaudhari.choromanska.ea.iclr17.tex
wip on the writeup

--- 2017-01-06 doc/chaudhari.choromanska.ea.iclr17.tex
@@
 
-\pagestyle{empty}
+%\pagestyle{empty}
 
@@
 \author{Pratik Chaudhari$^{1}$, Anna Choromanska$^{2}$, Stefano Soatto$^{1}$, Yann LeCun$^{2,3}$, Carlo Baldassi$^{4}$,\\[0.03in]
-\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, Riccardo Zecchina$^{4}$}\\[0.05in]
+\textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
+Riccardo Zecchina$^{4}$}\\[0.05in]
 $^{1}$ Computer Science, University of California, Los Angeles.\\
@@
 This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape.
-We construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
+Local extrema with low generalization error have a large proportion of almost-zero eigenvalues in the Hessian with very few positive or negative eigenvalues.
+We leverage upon this observation to construct a local entropy based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
 Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
-We prove that incorporating the local entropy into the objective function results in a smoother energy landscape, and we use uniform stability to show improved generalization bounds over SGD.
-Our experiments on competitive baselines demonstrate that $\entropysgd$ alone leads to generalization performances that compare with state of the art techniques and has the potential to accelerate training.
-\pc{}{We also perform a direct comparison between the information provided by the local entropy measure and the analysis of the Hessian in qualitatively different minima.}
+We prove that the local entropy objective function is smoother than the original landscape and use uniform stability to show improved generalization bounds over SGD.
+Our experiments on competitive baselines demonstrate that $\entropysgd$ leads to generalization performances that compare with state of the art techniques and also has the potential to accelerate training.
 \end{abstract}
@@
 \label{s:intro}
-This paper presents a new optimization tool for deep learning, designed to exploit the local geometrical properties of the objective function.
-Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
-These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
-The original analysis was done in the case of binary weights, and was later generalized by~\citet{baldassi2016multilevel} to the case of weights with an arbitrary number of precision bits.
-Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.
-
-The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian.
-Consider the histogram in Fig.~\ref{fig:lenet_hessian}, showing the spectrum of the Hessian at the endpoint of the optimization performed by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+This paper presents a new optimization tool for deep learning designed to exploit the local geometric properties of the objective function.
+Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 Interestingly, this trend is not unique to this particular network.
-Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (see references in Sec.~\ref{s:prior_work}, and Sec.~\ref{s:expt} for more experiments).
-\pc{Local minima}{We are using the term ``local minima'' loosely here, since a few of the eigenvalues of the Hessian may be slightly negative} that generalize well and are discovered by gradient descent lie in "wide valleys" of the energy landscape, rather than in sharp isolated minima.
-For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
-Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
+Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (refer to Sec.~\ref{s:expt} for more experiments).
+Local minima that generalize well and are discovered by gradient descent lie in ``wide valleys'' of the energy landscape, rather than in sharp isolated minima.
+\todo{For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the marginal likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.} Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.
@@
 $$
-% \rho
-f(x) 
-- \log \int_{x' \in \reals^n}\ \exp \rbrac{-\b f(x') - \f{\b \g}{2}\ \norm{x-x'}^2_2}\ dx'.
+- \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx'.
 $$
-The second term measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$;  we call it the {\it local free entropy} in analogy to the free entropy in statistical physics.
-The $\entropysgd$ algorithm presented in this paper employs a few iterations of stochastic gradient Langevin dynamics (SGLD) to approximate the second term before each update of the parameters using the back-propagated gradient.
-At a conceptual level, this algorithm resembles running two nested loops of SGD.
+The above is a log-partition function measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$; we call it ``local entropy'' in analogy to the free entropy used in statistical physics.
+The $\entropysgd$ algorithm presented in this paper employs stochastic gradient Langevin dynamics (SGLD) to approximate the gradient of local entropy.
+Our algorithm resembles two nested loops of SGD: the inner loop consists of SGLD iterations while the outer loop updates the parameters.
 We show that the above modified loss function results in a smoother energy landscape defined by the hyper-parameter $\g$ which we can think of as a ``scope'' that seeks out valleys of specific widths.
-Actively biasing towards wide valleys aids generalization, in fact, we can optimize solely the free energy term to obtain similar generalization error as SGD on the original loss function.
-Our experiments in Sec.~\ref{s:expt} on MNIST and CIFAR-10~\citep{krizhevsky2009learning} show that $\entropysgd$ scales to deep networks used in practice and in terms of the number of updates to the parameters, trains much more quickly than SGD.
+Actively biasing the optimization towards wide valleys in the energy landscape results in better generalization error.
+\pc{We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB) and character-level text prediction.
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD.}{}
 
@@
 
-The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing,sagun2014explorations} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
+The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
 Although assumptions in these works are somewhat unrealistic in the context of deep networks used in practice, similar results are also true for linear networks which afford a more thorough analytical treatment~\citep{DBLP:journals/corr/SaxeMG13}.
@@
 
-Closer to the subject of this work,~\citet{baldassi2015subdominant,baldassi2016unreasonable} show that, both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average.
+Closer to the subject of this work are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} which show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
+These "dense local minima" can be shown to generalize well for random binary input data; more importantly, they are also algorithmically accessible.
+
+A belief propagation based approximation of the local entropy is the key to efficiently discovering such dense minima.
+
+both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average.
 Using a novel out-of-equilibrium analysis, the authors show that such an ensemble recovers dense clusters in the solution space.
-Such dense clusters generalize better than isolated local minima.
+
 A related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and encourages individual models to explore by modulating their distance to the ensemble average; this is a special case of robust ensembles, as introduced in~\citet{baldassi2016unreasonable}.
+
+Our work generalizes this approach to modern, general deep architectures and continuous weights.
 
@@
 The enabling technique in our case is that we approximate the local entropy using Langevin dynamics.
-\citet{baldassi2016local} instead employ belief propagation to do so, but it's not clear whether this scales well beyond the shallow networks considered there.
 
@@
 
-\section{Local free entropy}
+\section{Local entropy}
 \label{s:local_entropy}
 
-We first provide a simple intuition for the concept of local free entropy of an energy landscape.
-Consider a cartoon energy landscape in Fig.~\ref{fig:scoping} where the x-axis denotes the configuration space of the parameters.
+We first provide a simple intuition for the concept of local entropy of an energy landscape.
+Consider a cartoon energy landscape in Fig.~\ref{fig:entropyfig} where the x-axis denotes the configuration space of the parameters.
 We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\trm{non-robust}}$.
-Under a Bayesian prior on the parameters, say a Gaussian of a fixed variance at locations $x_{\trm{robust}}$ and $x_{\trm{non-robust}}$ respectively, the wider local minimum has a higher log-likelihood than the sharp valley on the right.
-%[This is incorrect -- k-SAT has different behavior.] In discrete optimization problems, such wide local minima are dense clusters of solutions in the parameter space where nearby parameters have similar energy, say, the number of clauses violated in $k$-SAT~\citep{sipser2006introduction,achlioptas2006solution}.
-\pc{On the other hand, sharp local minima are isolated solutions that are far away from other solutions.}{cite Baldassi in this section}
+Under a Bayesian prior on the parameters, say a Gaussian of a fixed variance at locations $x_{\trm{robust}}$ and $x_{\trm{non-robust}}$ respectively, the wider local minimum has a higher marginalized likelihood than the sharp valley on the right.
 
@@
 \centering
-\includegraphics[width=0.4 \textwidth]{scoping.pdf}
-\caption{\small Local entropy concentrates on dense clusters in the energy landscape.}
-\label{fig:scoping}
-\vspace{0.2in}
+\includegraphics[width=0.4 \textwidth]{entropyfig.pdf}
+\caption{\small Local entropy concentrates on wide valleys in the energy landscape.}
+\label{fig:entropyfig}
+\vspace*{-0.1in}
 \end{wrapfigure}
@@
 The above discussion suggests that parameters that lie in wider local minima like $x_{\trm{robust}}$, which may possibly have a higher loss than the global minimum, should generalize better than the ones that are simply at the global minimum.
-This motivates us to maximize the ``local free entropy'' term presented in the introduction: In a neighborhood of $x_{\trm{robust}}$, that term is large because it includes the contributions from a whole region of good configurations; conversely, in the region near $x_{\trm{non-robust}}$ there are no such contributions except for the minimum itself and the resulting local free entropy is low.
-The local free entropy thus provides a way of picking large,
-approximately flat, regions of the landscape over sharp, narrow valleys in spite of the latter possibly having a lower loss.
-Quite conveniently, the local free entropy is also computed from the partition function with a local re-weighting term.
+In a neighborhood of $x_{\trm{robust}}$, ``local entropy'' as introduced in Sec.~\ref{s:intro} is large because it includes the contributions from a large region of good parameters; conversely, near $x_{\trm{non-robust}}$, there are fewer such contributions and the resulting local entropy is low.
+The local entropy thus provides a way of picking large, approximately flat, regions of the landscape over sharp, narrow valleys in spite of the latter possibly having a lower loss.
+Quite conveniently, the local entropy is also computed from the partition function with a local re-weighting term.
 
@@
 where $\b$ is known as the inverse temperature and $Z_\b$ is a normalizing constant, also known 
-as the partition function in physics. As $\b \to \infty$, the probability distribution above concentrates on the global minimum of $f(x)$ (assuming it's unique) given as:
+as the partition function. As $\b \to \infty$, the probability distribution above concentrates on the global minimum of $f(x)$ (assuming it is unique) given as:
 \begin{equation}
@@
 which establishes the link between the Gibbs distribution and a generic optimization problem~\eqref{eq:argmin_generic}.
-We would instead like the probability distribution --- and therefore the underlying optimization problem --- to focus on flat regions such as $x_{\textrm{robust}}$, as in the figure.
+We would instead like the probability distribution --- and therefore the underlying optimization problem --- to focus on flat regions such as $x_{\textrm{robust}}$ in Fig.~\ref{fig:entropyfig}.
 With this in mind, let us construct a modified Gibbs distribution:
@@
 The distribution in~\eqref{eq:gibbs_modified} is a function of a dummy variable $x'$ and is parameterized by the original location $x$.
-The parameter $\g$ biases the modified distribution~\eqref{eq:gibbs_modified} towards $x$; a large $\g$ results in a $P(x';\ x,\b,\g)$ with all its mass near $x$ irrespective of the energy landscape of $f(x')$.
+The parameter $\g$ biases the modified distribution~\eqref{eq:gibbs_modified} towards $x$; a large $\g$ results in a $P(x';\ x,\b,\g)$ with all its mass near $x$ irrespective of the energy term $f(x')$.
 For small values of $\g$, the term $f(x')$ in the exponent dominates and the modified distribution is similar to the original Gibbs distribution in~\eqref{eq:gibbs_original}.
-%We will set the inverse temperature $\b$ to $1$ because $\g$ affords us similar control on the Gibbs distribution.
+We will set the inverse temperature $\b$ to $1$ because $\g$ affords us similar control on the Gibbs distribution.
 
-\begin{definition}[Local free entropy]
+\begin{definition}[Local entropy]
 \label{def:local_entropy}
@@
 \begin{align}
-    F(x, \beta, \g) &= \log Z_{x, \beta, \g} \notag\\
-    &= \log \int_{x'}\ \exp \Big(-\beta f(x') - \f{\beta\g}{2}\ \norm{x-x'}^2_2 \Big)\ dx'.
+    F(x, \g) &= \log Z_{x, 1, \g} \notag\\
+    &= \log \int_{x'}\ \exp \Big(-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2 \Big)\ dx'.
     \label{eq:local_entropy}
@@
 The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
-Fig.~\ref{fig:scoping} shows the negative local entropy $-F(x, \b,  \g)$ for the original energy landscape.
-Note that $-F(x, \b, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.
+Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
+Note that $-F(x, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.
 
-Let us also note that the quantity we have defined as local entropy in Def.~\ref{def:local_entropy} is different from classical entropy which counts the number of likely configurations under a given distribution.
-This is given by
-$
-    - \int_{x'}\ \log \P(x';\ x, \b, \g)\ d \P(x';\ x, \b, \g)
-$
+\paragraph{Connection to classical entropy:}
+The quantity we have defined as local entropy in Def.~\ref{def:local_entropy} is different from classical entropy which counts the number of likely configurations under a given distribution.
+For a continuous parameter space, this is given by
+$$
+    S(x, \b, \g) = - \int_{x'}\ \log \P(x';\ x, \b, \g)\ d \P(x';\ x, \b, \g)
+$$
 for the Gibbs distribution in~\eqref{eq:gibbs_modified}.
 Minimizing classical entropy however does not differentiate between flat regions that have very high loss versus dense regions that lie deeper in the energy landscape.
-For instance, in Fig.~\ref{fig:scoping}, classical entropy is smallest in the neighborhood of $x_{\textrm{candidate}}$ which is a large region with very high loss on the training dataset and is unlikely to generalize well.
+For instance, in Fig.~\ref{fig:entropyfig}, classical entropy is smallest in the neighborhood of $x_{\textrm{candidate}}$ which is a large region with very high loss on the training dataset and is unlikely to generalize well.
 \ignore{We could minimize a modified loss function of the form $f(x) + \lambda S(x, \g)$ whose gradient can be computed to be
@@
 
-We now present the $\entropysgd$ algorithm, a variant of SGD that is motivated from local entropy.
-Simply speaking, we modify the loss function of a deep network to include the negative local entropy from Sec.~\ref{s:local_entropy}.
-The gradient of the modified loss can then be computed using Langevin dynamics.
-The $\entropysgd$ algorithm (cf.
-Alg.~\ref{alg:entropysgd}) thus has a strong flavor of ``SGD-inside-SGD'': the outer SGD updates the parameters, while an inner SGD estimates the gradient of local entropy at each iteration.
+We now present the $\entropysgd$ algorithm that is a variation of SGD motivated from local entropy.
+Simply speaking, it minimizes the negative local entropy from Sec.~\ref{s:local_entropy}.
+This section discusses how the gradient of local entropy can be computed via Langevin dynamics; the resulting algorithm has a strong flavor of ``SGD-inside-SGD'': the outer SGD updates the parameters, while an inner SGD estimates the gradient of local entropy.
 
@@
 \begin{equation}
-    x^*_{\entropysgd} = \argmin_x\ \r\ \rbrac{\f{1}{N}\ \sum_{k=1}^N\ f(x;\ \xi_k)} - F(x, \gamma;\ \Xi);
+    x^*_{\entropysgd} = \argmin_x\ -F(x, \gamma;\ \Xi);
     \label{eq:entropysgd_objective}
@@
 where we have made the dependence of local entropy $F(x,\g)$ on the dataset $\Xi$ explicit.
-The objective in~\eqref{eq:entropysgd_objective} thus scales the original loss by a hyper-parameter $\rho$ and adds the negative local entropy term $F(x, \g, \Xi)$ introduced in Sec.~\ref{s:local_entropy}.
-The parameter $\rho$ allows us to control the contribution of the original back-propagated gradient in $\entropysgd$.
-Sec.~\ref{ss:intuition_local_entropy} and Sec.~\ref{ss:theoretical_properties} below discuss the effect of $\rho$ and the scoping parameter $\g$.
 
-\subsection{Computing the gradient of local entropy}
+\subsection{Gradient of local entropy}
 \label{ss:grad_local_entropy}
@@
 \begin{equation}
-    \nabla_x F\rbrac{x, \g;\ \minibatch{\ell}} = -\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}};
-    \label{eq:local_entropy_grad}
+    -\nabla_x F\rbrac{x, \g;\ \minibatch{\ell}} = \g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}};
+    \label{eq:entropysgd_grad}
 \end{equation}
@@
 \end{equation}
-Computationally, the gradient in~\eqref{eq:local_entropy_grad} involves estimating $\ag{x';\ \minibatch{\ell}}$ with the current weights fixed to $x$.
+Computationally, the gradient in~\eqref{eq:entropysgd_grad} involves estimating $\ag{x';\ \minibatch{\ell}}$ with the current weights fixed to $x$.
 This is an expectation over a Gibbs distribution and is hard to compute.
@@
 Please see Sec.~\ref{s:app:langevin} in the Supplement for a brief overview of SGLD.
-For our application, as lines 3-6 of Alg.~\ref{alg:entropysgd} show SGLD resembles a few iterations of SGD with additive gradient noise.
+For our application, as lines 3-6 of Alg.~\ref{alg:entropysgd} show, SGLD resembles a few iterations of SGD with additive gradient noise.
 
-The gradient of the first term in~\eqref{eq:entropysgd_objective} is computed using back-propagation and the final gradient is
-\begin{equation}
-    \r\ \rbrac{\f{1}{m}\ \sum_{i=1}^m\ \nabla f(x;\ \xi_{k_i})} + \g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}.
-    \label{eq:entropysgd_grad}
-\end{equation}
-Note that we use different mini-batches (denoted by $k$ and $\ell$ respectively) to compute the back-propagated gradient and gradient of the local entropy.
-From our experiments, we observe that using the same mini-batches for the gradients performs marginally worse in terms of convergence speed.
-This is explained by the fact that local entropy is really a log-partition function and depends upon the entire dataset.
-
-\subsection{Intuition behind local entropy}
-\label{ss:intuition_local_entropy}
-
-Let us give some intuition on how local entropy works using the expression for the gradient: the term $\ag{x';\ \cdot}$ is the average over a locally focused Gibbs distribution and for two local minima in the neighborhood of $x$ roughly equivalent in loss, this term points towards the wider one because $\ag{x';\ \cdot}$ is closer to it.
+We can obtain some intuition of how $\entropysgd$ works using the expression for the gradient: the term $\ag{x';\ \cdot}$ is the average over a locally focused Gibbs distribution and for two local minima in the neighborhood of $x$ roughly equivalent in loss, this term points towards the wider one because $\ag{x';\ \cdot}$ is closer to it.
 This results in a net gradient that takes SGD towards wider valleys.
@@
 These two phenomena intuitively explain the improved generalization performance of $\entropysgd$.
-
-The local entropy term in the objective results in a smoother energy landscape than the original loss function.
-From our experience (cf.\@ experiments in Sec.~\ref{ss:expt:mnist}), we can also set $\rho$ to zero, i.e., consider the modified loss function for $\entropysgd$ to be simply $-F(x, \g)$ and completely eliminate $\nabla f(x)$ from the outer loop of $\entropysgd$.
-This gives similar generalization error as that of SGD for simpler architectures such as a convolutional neural network on MNIST.
-However, for more complex datasets and networks, the smooth energy landscape generated by $-F(x, \g)$ might have a different local minimum as compared to the original loss $f(x)$.
-We found that in order to fine-tune larger networks using the objective in~\eqref{eq:entropysgd_objective} with a small $\rho$ performs slightly better than setting $\rho$ to zero.
 
@@
 The weights $x$ are updated with the modified gradient on line $9$.
-
-We set the number of SGLD iterations $L$ to $5$-$20$ depending upon the complexity of the dataset, e.g., for experiments on convolutional networks on MNIST and CIFAR-10, we use $L = 20$ while for experiments on a fully-connected network on MNIST $L=5$ works well (cf.
-Sec.~\ref{s:expt}).
-We perform exponential averaging of the Langevin iterates $x'$ to estimate $\mu$ with a parameter $\a = 0.9$ so as to put more weight on later samples.
-In order to quickly transition to the Langevin dynamics phase of SGLD and ensure the we are sampling from the posterior even for small values of $L$, we set the learning rate $\eta'$ on line $5$ to be small factor of the outer SGD's learning rate: $\eta' = 0.1\ \eta$~\citep{welling2011bayesian}.
 
@@
     \Input{\quad $\textrm{current weights}\ x$, $\textrm{Langevin iterations}\ L$}
-    \HyperParameters{\quad $\textrm{Scope}\ \g$, $\textrm{Data weight}\ \rho$, $\textrm{Learning rate}\ \eta$}
+    \HyperParameters{\quad $\textrm{Scope}\ \g$, $\textrm{Learning rate}\ \eta$}
     \vspace{0.1in}
@@
         \vspace{0.03in}
-        $x' \la x' - \eta'\ dx' + 2\sqrt{\eta'}\ \trm{N}(0, \trm{I})$\;
+        $x' \la x' - \eta'\ dx' + 2\sqrt{\eta'}\ \trm{N}(0, \e\ \trm{I})$\;
         \vspace{0.03in}
@@
     \nonl \textrm{// Update weights}\;
-    \vspace{0.025in}
-    $\minibatch{k} \la \textrm{sample mini-batch}$\;
     \vspace{0.03in}
-    $dx \la \f{1}{m}\ \sum_{i=1}^m\ \nabla_{x} f\rbrac{x;\ \xi_{k_i}}$\;
-    \vspace{0.03in}
-    $x \la x - \eta\ \big(\rho\ dx + \g\ (x-\mu) \big)$
+    $x \la x - \eta\ \g\ (x-\mu)$
     \caption{$\entropysgd$ algorithm}
@@
 
+\todo{
+We set the number of SGLD iterations $L$ to $5$-$20$ depending upon the complexity of the dataset, e.g., for experiments on convolutional networks on MNIST and CIFAR-10, we use $L = 20$ while for experiments on a fully-connected network on MNIST $L=5$ works well (cf.
+Sec.~\ref{s:expt}).
+We perform exponential averaging of the Langevin iterates $x'$ to estimate $\mu$ with a parameter $\a = 0.9$ so as to put more weight on later samples.
+In order to quickly transition to the Langevin dynamics phase of SGLD and ensure the we are sampling from the posterior even for small values of $L$, we set the learning rate $\eta'$ on line $5$ to be small factor of the outer SGD's learning rate: $\eta' = 0.1\ \eta$~\citep{welling2011bayesian}.}
+
 Let us note that although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
 In our experiments, we have used both SGD with Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
+
+\subsubsection{Scoping of $\gamma$}
+
+\subsection{Connection to variational inference}
+\label{ss:connection_variational}
 
@@
 
-\pc{Our main result in this section is to show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
+\pc{We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
 Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).}{Clarify that the analysis here is only a sketch, it is not rigorous}
@@
 which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$.
-For the objective~\eqref{eq:entropysgd_objective}, we have the constant $1 - \r - \f{1}{1 + \g^{-1}\ c}$ in the exponent instead.
 

### 2017-01-09 hessian/entropyfig.py

@@ -60,13 +60,13 @@ f, s,_ = smooth(0, 1e-3)
 f = f + 1
 s = s - s.min()
 
-fig = plt.figure(1, figsize=(7,8))
+fig = plt.figure(1, figsize=(9,8))
 plt.clf()
 ax = fig.add_subplot(111)
 plt.plot(f, 'k-', label=r'$\mathrm{Original\ landscape}$', lw=1)
 plt.plot(s, 'indianred', label=r'$\mathrm{Negative\ local\ entropy}:\ \gamma = 0.001$', lw=2)
 plt.legend(loc=2)
-plt.title(r'$\mathrm{Scoping\ out\ dense\ clusters}$')
+#plt.title(r'$\mathrm{Local\ entropy\ energy\ landscape}$')
 plt.xticks([], [])
 plt.plot([576.73,340.52,811.8], [0.938, 0.005, 0.228], 'o', c='indianred', markersize=10)
 
@@ -78,4 +78,4 @@ ax.text(835.8, 0.2, r'$x_{\mathrm{non}\textrm{-}\mathrm{robust}}$', fontsize=fon
         verticalalignment='center')
 
 if opt['save']:
-    plt.savefig('../../doc/fig/scoping.pdf', bbox_inches='tight')
+    plt.savefig('../doc/fig/entropyfig.pdf', bbox_inches='tight')

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 We leverage upon this observation to construct a local entropy based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
-Our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
+Conceptually, our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
 We prove that the local entropy objective function is smoother than the original landscape and use uniform stability to show improved generalization bounds over SGD.
@@
 Based on this understanding of how the local geometry looks at the end of optimization, can we modify SGD to actively seek such regions?
-Motivated by the work of~\citet{baldassi2015subdominant,baldassi2016unreasonable} on shallow networks, we propose to replace the original loss function $f(x)$ by:
+Motivated by the work of~\citet{baldassi2015subdominant} on shallow networks, we propose to replace the original loss function $f(x)$ by:
 $$
-- \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx'.
+F(x, \g) = - \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx'.
 $$
@@
 Actively biasing the optimization towards wide valleys in the energy landscape results in better generalization error.
-\pc{We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB) and character-level text prediction.
-Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD.}{}
+We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB)~\citep{marcus1993building} and character-level text prediction.
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD.
 
@@
 
-Closer to the subject of this work are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} which show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
-These "dense local minima" can be shown to generalize well for random binary input data; more importantly, they are also algorithmically accessible.
+\todo{Closer to the subject of this paper are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} which show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
+These dense local minima can be shown to generalize well for random input data; more importantly, they are also accessible by efficient algorithms using a novel measure called ``robust ensemble'' that amplifies the weight of such dense regions.
+Algorithms
 
-A belief propagation based approximation of the local entropy is the key to efficiently discovering such dense minima.
-
-both theoretically and practically, ensembles of perceptrons can be trained in parallel by penalizing each individual model to be close to the ensemble average.
-Using a novel out-of-equilibrium analysis, the authors show that such an ensemble recovers dense clusters in the solution space.
-
-A related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and encourages individual models to explore by modulating their distance to the ensemble average; this is a special case of robust ensembles, as introduced in~\citet{baldassi2016unreasonable}.
-
-Our work generalizes this approach to modern, general deep architectures and continuous weights.
-
-An ensemble training procedure thus provides improved generalization but it incurs a large complexity in terms of hardware and communication across workers.
-In this paper, we construct an optimization algorithm for a general deep network that can exploit the same phenomenon of out-of-equilibrium dense clusters but does not require distributed training.
-The enabling technique in our case is that we approximate the local entropy using Langevin dynamics.
+Let us note that a related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker to the ensemble average.
+An ensemble training procedure provides improved generalization by ensuring that different works land in the same wide valley and indeed, this turns out to be closely related to the robust ensemble approach.
+%
+Our work generalizes the local entropy approach to modern deep networks with continuous weights.
+It exploits the same phenomenon of wide valleys in the energy landscape but does so without incurring the hardware and communication complexity of a distributed training algorithm.}
 
@@
         \vspace{0.03in}
-        $x' \la x' - \eta'\ dx' + 2\sqrt{\eta'}\ \trm{N}(0, \e\ \trm{I})$\;
+        $x' \la x' - \eta'\ dx' + \sqrt{\eta'}\ \trm{N}(0, \e\ \trm{I})$\;
         \vspace{0.03in}
@@
 \footnotesize
+\linespread{1.0}
 \bibliographystyle{iclr2017_conference}
@@
 $$
-This shows that a smaller value of $\g$ results in a smoother energy landscape, except at places with very flat directions.
-The Lipschitz constant also decreases by the same factor.
-The objective $\r f(x) - F(x,\g)$ results in a modified factor $\r + \f{1}{1+\g^{-1}\ c}$.
+This shows that a smaller value of $\g$ results in a smoother energy landscape, except at places with very flat directions. The Lipschitz constant also decreases by the same factor.
 \end{proof}

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
wip on expts. section

@@
 \newcommand{\smallmnistfc}{\textrm{small-mnistfc}}
-\newcommand{\charlstm}{\textrm{char-lstm}}
+\newcommand{\charlstm}{\textrm{char-LSTM}}
+\newcommand{\ptblstm}{\textrm{PTB-LSTM}}
 \newcommand{\lenet}{\textrm{LeNet}}
@@
 \begin{abstract}
-This paper proposes a new optimization algorithm called $\entropysgd$ for training deep neural networks that is motivated by the local geometry of the energy landscape.
+This paper proposes a new optimization algorithm called Entropy-SGD for training deep neural networks that is motivated by the local geometry of the energy landscape.
 Local extrema with low generalization error have a large proportion of almost-zero eigenvalues in the Hessian with very few positive or negative eigenvalues.
 We leverage upon this observation to construct a local entropy based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
-Conceptually, our algorithm resembles two nested loops of SGD, where we use Langevin dynamics to compute the gradient of the local entropy at each update of the weights.
-We prove that the local entropy objective function is smoother than the original landscape and use uniform stability to show improved generalization bounds over SGD.
-Our experiments on competitive baselines demonstrate that $\entropysgd$ leads to generalization performances that compare with state of the art techniques and also has the potential to accelerate training.
+Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
+We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.
+Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state of the art techniques in terms of generalization performance and training time.
 \end{abstract}
@@
 Based on this understanding of how the local geometry looks at the end of optimization, can we modify SGD to actively seek such regions?
-Motivated by the work of~\citet{baldassi2015subdominant} on shallow networks, we propose to replace the original loss function $f(x)$ by:
+Motivated by the work of~\citet{baldassi2015subdominant} on shallow networks, instead of minimizing the original loss $f(x)$, we propose to maximize
 $$
-F(x, \g) = - \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx'.
+F(x, \g) = \log \int_{x' \in \reals^n}\ \exp \rbrac{-f(x') - \f{\g}{2}\ \norm{x-x'}^2_2}\ dx'.
 $$
@@
 These dense local minima can be shown to generalize well for random input data; more importantly, they are also accessible by efficient algorithms using a novel measure called ``robust ensemble'' that amplifies the weight of such dense regions.
-Algorithms
+The authors use belief propagation to estimate local entropy for simpler models considered there.
+A related work in this context is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker to the ensemble average.
+Such an ensemble training procedure enables improved generalization by ensuring that different workers land in the same wide valley and indeed, it turns out to be closely related to the replica theoretic analysis of~\citep{baldassi2016unreasonable}.
 
-Let us note that a related work in this direction is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker to the ensemble average.
-An ensemble training procedure provides improved generalization by ensuring that different works land in the same wide valley and indeed, this turns out to be closely related to the robust ensemble approach.
-%
-Our work generalizes the local entropy approach to modern deep networks with continuous weights.
-It exploits the same phenomenon of wide valleys in the energy landscape but does so without incurring the hardware and communication complexity of a distributed training algorithm.}
+Our work generalizes the local entropy approach above to modern deep networks with continuous weights.
+It exploits the same phenomenon of wide valleys in the energy landscape but does so without incurring the hardware and communication complexity of replicated training or being limited to models where one can estimate local entropy using belief propagation.
+The enabling technique in our case is using Markov chain Monte Carlo for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.}
 
-The authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
+The authors in~\citet{hochreiter1997flat} introduce a similar technique as our work, they introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
 \pc{Their analysis of the generalization performance (see their Sec.
-A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
+A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{one sentence about the similarities} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
 Their formulation depends on Hessian-vector products.
@@
 At each iteration, lines $3$-$6$ perform $L$ iterations of Langevin dynamics to estimate $\mu = \ag{x'; \minibatch{\ell}}$.
-The weights $x$ are updated with the modified gradient on line $9$.
+The weights $x$ are updated with the modified gradient on line $7$.
 
@@
     \Input{\quad $\textrm{current weights}\ x$, $\textrm{Langevin iterations}\ L$}
-    \HyperParameters{\quad $\textrm{Scope}\ \g$, $\textrm{Learning rate}\ \eta$}
+    \HyperParameters{\quad $\textrm{scope}\ \g$, $\textrm{learning rate}\ \eta$, $\textrm{SGLD step size}\ \eta'$}
     \vspace{0.1in}
@@
         \vspace{0.03in}
-        $x' \la x' - \eta'\ dx' + \sqrt{\eta'}\ \trm{N}(0, \e\ \trm{I})$\;
+        $x' \la x' - \eta'\ dx' + \sqrt{\eta'}\ \e\ \trm{N}(0, \trm{I})$\;
         \vspace{0.03in}
@@
 
+Let us note that although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
+In our experiments, we use Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
+
 \todo{
@@
 
-Let us note that although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
-In our experiments, we have used both SGD with Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
+\subsubsection{Scoping of $\gamma$}
+\label{ss:scoping}
 
-\subsubsection{Scoping of $\gamma$}
+Our scoping mechanism introduced in Sec.~\ref{ss:scoping} increases the value of $\g$ as training progresses in order to focus on more local neighborhoods.
 
-\subsection{Connection to variational inference}
-\label{ss:connection_variational}
+The ``reverse-annealing'' of the scope $\g$ typically interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
+In practice, we therefore scale the local entropy gradient by $\g$ before the update and modify the line to read
+$$
+    x \leftarrow x - \eta (x - \mu).
+$$
+We found this to be very effective in simplifying hyper-parameter tuning.
 
@@
 \pc{We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
-Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).}{Clarify that the analysis here is only a sketch, it is not rigorous}
+Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).}{clarify that the analysis here is only a sketch, it is not rigorous}
 
@@
 \centering
-    \begin{subfigure}[t]{0.47 \textwidth}
+    \begin{subfigure}[t]{0.44 \textwidth}
         \includegraphics[width=1.08\textwidth]{mnistfc_hessian.pdf}
@@
     \hspace{0.2in}
-    \begin{subfigure}[t]{0.47 \textwidth}
+    \begin{subfigure}[t]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{charlstm_hessian.pdf}
@@
         \centering
-        \begin{subfigure}[b]{0.47 \textwidth}
+        \begin{subfigure}[b]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dm.pdf}
@@
         \hspace{0.2in}
-        \begin{subfigure}[b]{0.47 \textwidth}
+        \begin{subfigure}[b]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dp.pdf}
@@
 
-In this section, we show that local entropy is a good measure of generalization, in fact, setting $\rho = 0$ in $\entropysgd$, i.e., completely disabling the original loss function is enough to train on MNIST.
-We consider two networks: the first is a fully-connected network with two hidden layers of $1024$ units each which we denote as ``$\mnistfc$'' and the second is a convolutional neural network similar to $\lenet$ but which batch-normalization~\citep{ioffe2015batch} and a dropout layer of probability $0.5$ added after each convolutional layer.
-We train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs and compare the average error over $5$ independent runs in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test}.
+We consider two prototypical networks: the first is a fully-connected network with two hidden layers of $1024$ units each which we denote as ``$\mnistfc$'' and the second is a convolutional neural network with the same size as $\lenet$ but with batch-normalization~\citep{ioffe2015batch}; both with a dropout of probability $0.5$ after each layer.
+As a baseline, we train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs to obtain an average error of \pc{$1.39\%$ and $0.51\%$}{update std.
+dev numbers} for $\mnistfc$ and $\lenet$ respectively over $5$ independent runs.
 
-$\entropysgd$ converges to an error of $1.39\%$ in $22$ epochs on $\mnistfc$ while Adam obtains the same (best) error after $67$ epochs.
-For a small network such as this, with a presumably easy energy landscape, we perform $5$ SGLD iterations and set $\g = 1$.
-For $\lenet$, we perform $20$ SGLD iterations and set $\g = 0.1$.
-In this case, $\entropysgd$ converges around epoch $40$ to an average error of $0.48\%$; Adam obtains a comparable error of $0.51\%$ after $100$ epochs.
-For this experiment, we use a small $\rho = 10^{-4}$ to fine-tune the network; the error without this is $0.52\%$.
+For both these networks, we train $\entropysgd$ for $5$ epochs with $L = 20$ and reduce the dropout probability ($0.15$ for $\mnistfc$ and $0.25$ for $\lenet$).
+The learning rate of the SGLD updates is fixed to $\eta' = 0.1$ while the outer loop's learning rate is set to $\eta=1$ and drops by a factor of $10$ after the second epoch; we use Nesterov's momentum for both the loops.
+The thermal noise in SGLD updates (Line 5 of Alg.~\ref{alg:entropysgd}) is set to $10^{-3}$.
+We use an exponentially increasing value of $\g$ for scoping, the initial value of the scope is set to $\g = 10^{-4}$ and this increases by a factor of $1.001$ after each parameter update.
+The results in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test} show that $\entropysgd$ obtains a comparable generalization error \pc{in fewer effective epochs.}{massage}
 
-\textbf{Remark:} A single ``epoch'' for all algorithms considered here corresponds to the number of parameter updates necessary to run through the entire dataset once.
-Note that both SGD and $\entropysgd$ update parameters after each mini-batch, but $\entropysgd$ also runs $L$ iterations of SGLD between two updates of the parameters.
+\textbf{Computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
+We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
+by multiplying the abscissae by a factor of $L$.
+Note that $L = 1$ for SGD or Adam.
 
-\begin{figure}[h!]
+\begin{figure}[htp!]
 \centering
-    \begin{subfigure}[t]{0.45\textwidth}
+    \begin{subfigure}[t]{0.4\textwidth}
         \centering
-        \includegraphics[width=0.98\textwidth]{mnistfc_valid.pdf}
-        \caption{\small $\mnistfc$: $\entropysgd$ with $\rho = 0$, $L = 5$.}
+        \includegraphics[width=\textwidth]{mnistfc_valid.pdf}
+        \caption{\small $\mnistfc$: Validation error}
         \label{fig:mnistfc_test}
     \end{subfigure}
-    \hspace{0.1in}
-    \begin{subfigure}[t]{0.45\textwidth}
+    \hspace{0.2in}
+    \begin{subfigure}[t]{0.4\textwidth}
         \centering
         \includegraphics[width=\textwidth]{lenet_valid.pdf}
-        \caption{\small $\lenet$: $\entropysgd$ $\rho = 10^{-4}$, $L = 20$.}
+        \caption{\small $\lenet$: Validation error}
         \label{fig:lenet_test}
@@
 
+\subsection{CIFAR-10}
+\label{ss:expt:cifar}
+
 \begin{figure}[htp!]
 \centering
-    \begin{subfigure}[b]{0.45\textwidth}
+    \begin{subfigure}[b]{0.4\textwidth}
         \centering
-        \includegraphics[width=1.03\textwidth]{allcnn_loss.pdf}
+        \includegraphics[width=\textwidth]{allcnn_loss.pdf}
         \caption{\small $\allcnn$: Training loss}
@@
     \end{subfigure}
-    \hspace{0.1in}
-    \begin{subfigure}[b]{0.45\textwidth}
+    \hspace{0.2in}
+    \begin{subfigure}[b]{0.4\textwidth}
         \centering
@@
     \end{subfigure}
-\caption{\small Comparison of $\entropysgd$ vs. Adam on CIFAR-10}
+\caption{\small Comparison of $\entropysgd$ vs. SGD on CIFAR-10}
 \label{fig:allcnn}
@@
 
-\subsection{CIFAR-10}
-\label{ss:expt:cifar}
+We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}.
+We consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer and all other architectural parameters such as filter-sizes and weight decay are kept the same.
+We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
+We obtain an average error of \pc{$8.30\%$}{std-dev?} over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
+\pc{Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.}{maybe remove?}
 
-We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}.
-As a baseline, we consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer; this is absent in the original paper.
-We recreate the training pipeline of the original authors with the same hyper-parameters and use SGD with Nesterov's momentum.
-We however decrease the initial learning rate of $0.1$ by a factor of $5$ after every $60$ epochs for SGD and after every $5$ epochs for $\entropysgd$.
-Probably due to batch-normalization, we obtain an average error of $8.30\%$ over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report.
-This gives us a very competitive baseline to compare $\entropysgd$ against.
-Let us note that the best result in literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error while a wide ResNet of~\citet{zagoruyko2016wide} obtains an error of $4.17\%$ with data augmentation.
+We train $\entropysgd$ with $L = 20$ for $10$ epochs with the original dropout of $0.5$.
+The initial learning rate of the outer loop is set to $\eta = 1$ and drops by a factor of $5$ every $4$ epochs, while the learning rate of the SGLD updates is fixed to $\eta' = 0.1$ with thermal noise $\e = 10^{-4}$.
+As the scoping scheme, we set the initial value of the scope to $\g = 0.03$ which increases by a factor of $1.001$ after each parameter update.
+%
+Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
+It shows that local entropy performs as well as SGD on a large CNN, both in terms of generalization error and training time.
 
-%We list a few other comparable networks in the literature in Table~\ref{tab:cifar10}.
-% \begin{wraptable}{r}{0.6 \textwidth}
-% \centering
-% \renewcommand{\arraystretch}{1.2}
-% \resizebox{0.5 \columnwidth}{!}
-% {
-% \begin{tabular}{p{6.5cm} r}
-%     Model & Test error ($\%$)\\  \hline
-%     %Maxout~\citep{goodfellow2013maxout} & $11.68$\\
-%     NiN~\citep{lincy13} & $10.41$\\
-%     All-CNN-C~\citep{springenberg2014striving} & $9.04$\\
-%     All-CNN-BN (baseline: SGD, 200 epochs) & $8.30\ \pm\ 0.08$\\
-%     All-CNN-BN ($\entropysgd$, 15 epochs) & $8.65\ \pm\ 0.06$\\
-%     ELU~\citep{clevert2015fast} & $6.55$\\ \hline
-% \end{tabular}
-% }
-% \caption{\small Error rates on CIFAR-10 without data augmentation.}
-% \label{tab:cifar10}
-% \end{wraptable}
+We see almost no plateauing of training loss or validation error for $\entropysgd$ in Fig.~\ref{fig:allcnn_loss}; this trait is shared across different networks and datasets in our experiments.
+Moreover, while the magnitude of the gradients is similar for both SGD and $\entropysgd$ --- in fact, we use this as a way to tune SGLD's hyper-parameters $\eta'$ and $\g$ --- we can use significantly larger learning rates for $\entropysgd$.
+SGD with large learning rates, say $\eta = 1$ for the above network, would not train at all.
+These two observations are a direct consequence of the additional smoothness of the local entropy landscape.
 
-Fig.~\ref{fig:allcnn} compares $\entropysgd$ against SGD with $20$ SGLD iterations for $15$ epochs.
-Since CIFAR-10 is a more challenging dataset than MNIST, we could not obtain better test errors than $9.5\%$ while training with only the local entropy term ($\rho = 0$).
-We therefore set $\rho = 0.01$ to obtain an error of $8.65\%$ over $5$ independent runs.
-Refer to Fig.~\ref{fig:allcnn_loss}, $\entropysgd$ is designed to be an algorithm that seeks out wide local minima as opposed to sharp valleys; a side-effect of this is to smooth the energy landscape as we show in Sec.~\ref{ss:theoretical_properties}.
-In practice, we see this as a steady decrease in training loss for $\entropysgd$ without any plateaus that appear in the training loss for SGD.
-Similarly, the decrease in validation error in Fig.~\ref{fig:allcnn_valid} is also steady without any plateaus.
-Note that the original network $\textrm{All-CNN-C}$ was trained using SGD for $350$ epochs~\citep{springenberg2014striving}.
+\subsection{Recurrent neural networks}
+\label{ss:expt:rnn}
+
+In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction.
+This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$.
+As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}.
+Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each with a dropout of probability $0.65$; this is a large network with about $66$ million weights.
+It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
+\pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which matches that of the original authors.
+
+We run $\entropysgd$ for this network for $5$ epochs with $L = 5$, which results in only $25$ effective epochs.
+We do not use scoping for this network and fix the scope to $\g = 10^{-3}$.
+The initial learning rate of the outer loop is $\eta = 1$ which reduces by a factor of $10$ at each epoch after the third epoch; the SGLD learning rate is fixed to $\eta' = 1$ with $\e = 10^{-4}$.
+We obtain a word perplexity of $80.116 \pm 0.069$ on the validation set and $77.656 \pm 0.171$ on the test set.
+As Fig.~\ref{fig:ptb_valid} shows, $\entropysgd$ trains significantly faster than SGD to a slightly better generalization perplexity.
+
+Next,
+
+\todo{Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs.
+While the SGD baseline on $\ptblstm$ does not use momentum and in fact does not train well with momentum, we set it to $0.5$ for $\entropysgd$.
+On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we reduced this to $\b_1 = 0.5$ for $\entropyadam$.}
+
+\begin{figure}[htp!]
+\centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{allcnn_valid.pdf}
+        \caption{\small $\ptblstm$: Validation, test perplexity}
+        \label{fig:ptb_valid}
+    \end{subfigure}
+    \hspace{0.2in}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{allcnn_loss.pdf}
+        \caption{\small $\charlstm$: Validation perplexity}
+        \label{fig:char_valid}
+    \end{subfigure}
+\caption{\small Comparison of $\entropysgd$ vs. SGD on PTB}
+\label{fig:rnn}
+\end{figure}
 
@@
 
-As discussed in Sec.~\ref{ss:intuition_local_entropy}, SGLD updates in the inner loop of our algorithm compute the average gradient of the original loss in the neighborhood of the current weights.
+\todo{
+As discussed in Sec.~\ref{ss:grad_local_entropy}, SGLD updates in the inner loop of our algorithm compute the average gradient of the original loss in the neighborhood of the current weights.
 An inexpensive and approximate way of computing these updates is to average the gradient over random perturbations of weights during the forward pass itself and, on the same mini-batch as the outer loop.
 In our experiments, using the same mini-batch performs slightly worse than $\entropysgd$ although it is still better than SGD.
-Such an implementation is close to averaged SGD of~\citep{polyak1992acceleration,bottou2012stochastic} and worth further study.
+Such an implementation is close to averaged SGD of~\citep{polyak1992acceleration,bottou2012stochastic} and worth further study.}
 
 In our experiments, $\entropysgd$ results in a comparable generalization error as SGD, but always has a lower cross-entropy loss on the training set.
-Roughly speaking, wide valleys favored by $\entropysgd$ seem to have lower empirical loss than local minima discovered by SGD.
-Theoretical models of deep networks (cf.\@ Sec.~\ref{s:prior_work}) suggest multiple local minima with the same loss but there is an absence of results in literature about the local geometry at these minima.
+This suggests the following in the context of energy landscapes of deep networks.
+Roughly, wide valleys favored by $\entropysgd$ are located deeper in the landscape with a lower empirical loss than local minima discovered by SGD where it presumably gets stuck.
+Such an interpretation is contrast to theoretical models of deep networks (cf.\@ Sec.~\ref{s:prior_work}) which predicts multiple equivalent local minima with the same loss.
+Our work suggests that geometric properties of the energy landscape are crucial to generalize well and provides algorithmic approaches to exploit them.
+However, the literature lacks general results about the geometry of the loss functions of deep networks --- convolutional neural networks in particular --- and this is a promising direction for future work.
 
@@
 
-\section{Acknowledgements}
+\section{Acknowledgments}
 \label{s:acknowledgements}
@@
 
+\section{Connection to variational inference}
+\label{s:app:connection_variational}
+The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
+
+\paragraph{Comparison with SGLD:}
+A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
+Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
+We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
+
 \end{appendices}

### 2017-01-10 dnn/ploterrors.py

@@ -57,7 +57,7 @@ def plot_lenet():
     fig = plt.figure(1, figsize=(8,7))
     plt.clf()
     ax = fig.add_subplot(111)
-    plt.title(r'LeNet: Validation error')
+    #plt.title(r'LeNet: Validation error')
 
     v1 = np.array([smooth(r1['valid'].ix[i]) for i in range(5)])
     v2 = np.array([smooth(r2['valid'].ix[i]) for i in range(5)])
@@ -72,7 +72,7 @@ def plot_lenet():
     plt.yticks(yticks, [str(y) for y in yticks])
     xticks = [20, 40, 60, 80, 100]
     plt.xticks(xticks, [str(x) for x in xticks])
-    plt.xlabel(r'\#backprops/\#parameter updates')
+    plt.xlabel(r'Epochs $\times$ L')
     plt.ylabel(r'\% Error')
 
     plt.plot([98.89], [0.512], 'o', c='k', markersize=10)
@@ -96,7 +96,7 @@ def plot_mnistfc():
     fig = plt.figure(2, figsize=(8,7))
     plt.clf()
     ax = fig.add_subplot(111)
-    plt.title(r'mnistfc: Validation error')
+    #plt.title(r'mnistfc: Validation error')
 
     v1 = np.array([smooth(r1['valid'].ix[i]) for i in range(5)])
     v2 = np.array([smooth(r2['valid'].ix[i]) for i in range(5)])
@@ -110,7 +110,7 @@ def plot_mnistfc():
     plt.xlim([20,120])
     yticks = [1.2, 1.4, 1.6, 1.8, 2]
     plt.yticks(yticks, [str(y) for y in yticks])
-    plt.xlabel(r'\#backprops/\#parameter updates')
+    plt.xlabel(r'Epochs $\times$ L')
     plt.ylabel(r'\% Error')
 
     plt.plot([69], [1.39], 'o', c='k', markersize=10)
@@ -126,78 +126,78 @@ def plot_mnistfc():
         plt.savefig('../doc/fig/mnistfc_valid.pdf', bbox_inches='tight')
 
 
-#def plot_allcnn():
-r1, r2 = load(sorted(glob.glob('../results/nov_expts/dnn/allcnn/original/*.log')), 200), \
-        load(sorted(glob.glob('../results/jan_expts/allcnn10/toplot/*.log')), 10)
+def plot_allcnn():
[…]

### 2017-01-10 dnn/plotlog.py

@@ -97,5 +97,11 @@ def plot_idx(fs, df,
     plt.figure(3)
     plt.clf()
     helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].loss,
+        'Validation loss')
+    plt.xlim([0,max_epochs])
+
+    plt.figure(4)
+    plt.clf()
+    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
         'Test loss')
     plt.xlim([0,max_epochs])

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
minor, adding testing code for char-rnn

@@
 As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}.
-Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each with a dropout of probability $0.65$; this is a large network with about $66$ million weights.
+Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each unrolled for $35$ time steps and a dropout of probability $0.65$; this is a large network with about $66$ million weights.
 It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
@@
 
-Next,
+Next, we train an LSTM to perform character-level text-prediction.
+As a dataset, following the experiments of~\citet{karpathy2015visualizing}, we use the text of War and Peace by Leo Tolstoy which contains about $3$ million divided into training ($80\%$), validation ($10\%$) and test ($10\%$) sets.
+We use an LSTM consisting of two layers of $128$ hidden units unrolled for $50$ time steps and a vocabulary of size $87$.
+We train the baseline with Adam with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of $1.227 \pm 0.006$.
 
-\todo{Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs.
+Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs.
 While the SGD baseline on $\ptblstm$ does not use momentum and in fact does not train well with momentum, we set it to $0.5$ for $\entropysgd$.
-On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we reduced this to $\b_1 = 0.5$ for $\entropyadam$.}
+On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we set $\b_1 = 0.5$ for $\entropyadam$.
+In contrast to this observation on RNNs, our experiments on CNNs above always used a momentum of $0.9$.
 

### 2017-01-10 dnn/plotlog.py

@@ -100,8 +100,8 @@ def plot_idx(fs, df,
         'Validation loss')
     plt.xlim([0,max_epochs])
 
-    plt.figure(4)
-    plt.clf()
-    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
-        'Test loss')
-    plt.xlim([0,max_epochs])
+    # plt.figure(4)
+    # plt.clf()
+    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
+    #     'Test loss')
+    # plt.xlim([0,max_epochs])

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 \newcommand{\entropysgd}{\mathrm{Entropy}\textrm{-}\mathrm{SGD}}
+\newcommand{\entropyadam}{\mathrm{Entropy}\textrm{-}\mathrm{Adam}}
 \newcommand{\minibatch}[1]{\Xi^{#1}}
@@
         \centering
-        \includegraphics[width=\textwidth]{mnistfc_valid.pdf}
+        \includegraphics[width=1.02\textwidth]{mnistfc_valid.pdf}
         \caption{\small $\mnistfc$: Validation error}
@@
         \centering
-        \includegraphics[width=\textwidth]{allcnn_loss.pdf}
+        \includegraphics[width=1.02\textwidth]{allcnn_loss.pdf}
         \caption{\small $\allcnn$: Training loss}
@@
 We use an LSTM consisting of two layers of $128$ hidden units unrolled for $50$ time steps and a vocabulary of size $87$.
-We train the baseline with Adam with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of $1.227 \pm 0.006$.
+We train the baseline with Adam for $50$ epochs with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of \pc{$1.227 \pm 0.006$}{this is a place-holder, redo}.
+
+As noted in Sec.~\ref{ss:alg}, we can easily wrap Alg.~\ref{alg:entropysgd} inside other variants of SGD such as Adam; indeed this involves simply supplying the local entropy gradient in place of the usual back-propagated gradient.
+For this example, we construct $\entropyadam$ which is simply Adam that uses the local entropy gradient (computed via SGLD).
+We run $\entropyadam$ for $5$ epochs with $L = 5$ with an initial learning rate of $0.01$ that decreases by a factor of $2$ at each epoch.
+Note that this again results in only $25$ effective epochs.
+We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{place-holder} over $4$ independent runs which is slightly better than the baseline in half as much wall-clock time.
+We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
 
@@
 
+\todo{the two plots below are place holders}
 \begin{figure}[htp!]
@@
         \centering
-        \includegraphics[width=\textwidth]{allcnn_loss.pdf}
+        \includegraphics[width=1.02\textwidth]{allcnn_loss.pdf}
         \caption{\small $\charlstm$: Validation perplexity}
@@
     \end{subfigure}
-\caption{\small Comparison of $\entropysgd$ vs. SGD on PTB}
+\caption{\small Comparison of $\entropysgd$ vs. SGD / Adam on RNNs}
 \label{fig:rnn}

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
wip on sec. 5 brush up

@@
 
-\todo{Closer to the subject of this paper are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} which show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
+Closer to the subject of this paper are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} who show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
 These dense local minima can be shown to generalize well for random input data; more importantly, they are also accessible by efficient algorithms using a novel measure called ``robust ensemble'' that amplifies the weight of such dense regions.
@@
 It exploits the same phenomenon of wide valleys in the energy landscape but does so without incurring the hardware and communication complexity of replicated training or being limited to models where one can estimate local entropy using belief propagation.
-The enabling technique in our case is using Markov chain Monte Carlo for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.}
+The enabling technique in our case is using Markov chain Monte Carlo for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.
 
@@
 
+\section{Connection to variational inference}
+\label{s:app:connection_variational}
+The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
+
 \section{Experiments}
@@
 We consider two prototypical networks: the first is a fully-connected network with two hidden layers of $1024$ units each which we denote as ``$\mnistfc$'' and the second is a convolutional neural network with the same size as $\lenet$ but with batch-normalization~\citep{ioffe2015batch}; both with a dropout of probability $0.5$ after each layer.
-As a baseline, we train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs to obtain an average error of \pc{$1.39\%$ and $0.51\%$}{update std.
-dev numbers} for $\mnistfc$ and $\lenet$ respectively over $5$ independent runs.
+As a baseline, we train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs to obtain an average error of $1.39 \pm 0.03\%$ and $0.51 \pm 0.01 \%$ for $\mnistfc$ and $\lenet$ respectively over $5$ independent runs.
 
@@
 We use an exponentially increasing value of $\g$ for scoping, the initial value of the scope is set to $\g = 10^{-4}$ and this increases by a factor of $1.001$ after each parameter update.
-The results in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test} show that $\entropysgd$ obtains a comparable generalization error \pc{in fewer effective epochs.}{massage}
+The results in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test} show that $\entropysgd$ obtains a comparable generalization error: $1.37 \pm 0.03 \%$ and $0.50 \pm 0.01 \%$, for $\mnistfc$ and $\lenet$ respectively.
 
@@
 We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
-We obtain an average error of \pc{$8.30\%$}{std-dev?} over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
-\pc{Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.}{maybe remove?}
+We obtain an average error of $8.30 \pm 0.1\%$ over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
+Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.
 
@@
 Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
-It shows that local entropy performs as well as SGD on a large CNN, both in terms of generalization error and training time.
+It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $8.28 \pm 0.11 \%$ in the same number of effective epochs.
 
@@
 These two observations are a direct consequence of the additional smoothness of the local entropy landscape.
+
+\paragraph{Comparison with SGLD:}
+\todo{A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
+Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
+We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.}
 
@@
 As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}.
-Our network called $\ptblstm$ consists of two layers with $1500$ hidden units each unrolled for $35$ time steps and a dropout of probability $0.65$; this is a large network with about $66$ million weights.
+Our network called $\ptblstm$ consists of two layers with $1500$ hidden units, each unrolled for $35$ time steps and a dropout of probability $0.65$; note that this is a large network with about $66$ million weights.
 It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
-\pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which matches that of the original authors.
+\pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which closely matches the results of the original authors.
 
-We run $\entropysgd$ for this network for $5$ epochs with $L = 5$, which results in only $25$ effective epochs.
-We do not use scoping for this network and fix the scope to $\g = 10^{-3}$.
-The initial learning rate of the outer loop is $\eta = 1$ which reduces by a factor of $10$ at each epoch after the third epoch; the SGLD learning rate is fixed to $\eta' = 1$ with $\e = 10^{-4}$.
+We run $\entropysgd$ on $\ptblstm$ for $5$ epochs with $L = 5$; note that this results in only $25$ effective epochs.
+We do not use scoping for this network and instead fix $\g = 10^{-3}$.
+The initial learning rate of the outer loop is $\eta = 1$ which reduces by a factor of $10$ at each epoch after the third epoch.
+The SGLD learning rate is fixed to $\eta' = 1$ with $\e = 10^{-4}$.
 We obtain a word perplexity of $80.116 \pm 0.069$ on the validation set and $77.656 \pm 0.171$ on the test set.
-As Fig.~\ref{fig:ptb_valid} shows, $\entropysgd$ trains significantly faster than SGD to a slightly better generalization perplexity.
+As Fig.~\ref{fig:ptb_valid} shows, $\entropysgd$ trains significantly faster than SGD and also achieves a slightly better generalization perplexity.
 
 Next, we train an LSTM to perform character-level text-prediction.
-As a dataset, following the experiments of~\citet{karpathy2015visualizing}, we use the text of War and Peace by Leo Tolstoy which contains about $3$ million divided into training ($80\%$), validation ($10\%$) and test ($10\%$) sets.
+As a dataset, following the experiments of~\citet{karpathy2015visualizing}, we use the text of War and Peace by Leo Tolstoy which contains about $3$ million characters divided into training ($80\%$), validation ($10\%$) and test ($10\%$) sets.
 We use an LSTM consisting of two layers of $128$ hidden units unrolled for $50$ time steps and a vocabulary of size $87$.
-We train the baseline with Adam for $50$ epochs with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of \pc{$1.227 \pm 0.006$}{this is a place-holder, redo}.
+We train the baseline with Adam for $50$ epochs with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of \pc{$1.227 \pm 0.006$}{this is a place-holder}.
 
@@
 Note that this again results in only $25$ effective epochs.
-We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{place-holder} over $4$ independent runs which is slightly better than the baseline in half as much wall-clock time.
+We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{this is a place-holder} over $4$ independent runs which is slightly better than the baseline in half as much wall-clock time.
 We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
@@
 
-\todo{
-As discussed in Sec.~\ref{ss:grad_local_entropy}, SGLD updates in the inner loop of our algorithm compute the average gradient of the original loss in the neighborhood of the current weights.
-An inexpensive and approximate way of computing these updates is to average the gradient over random perturbations of weights during the forward pass itself and, on the same mini-batch as the outer loop.
-In our experiments, using the same mini-batch performs slightly worse than $\entropysgd$ although it is still better than SGD.
-Such an implementation is close to averaged SGD of~\citep{polyak1992acceleration,bottou2012stochastic} and worth further study.}
-
 In our experiments, $\entropysgd$ results in a comparable generalization error as SGD, but always has a lower cross-entropy loss on the training set.
@@
 
+\pc{If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters.
+Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study.
+Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in wall-clock time for CNNs.
+Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}
+
 \section{Conclusions}
@@
 We connected this observation to the concept of local entropy which we used to bias the optimization towards flat regions that have low generalization error.
-Our experiments showed that this algorithm is applicable to large deep networks used in practice.
+Our experiments showed that this algorithm is applicable to large convolutional and recurrent deep networks used in practice.
 
@@
 \footnotesize
-\linespread{1.0}
+\linespread{0.8}
 \bibliographystyle{iclr2017_conference}
@@
 
-\begin{proof}[Proof of Lemma~\ref{lem:smoothness_reduction}]
+\begin{proof}[Proof sketch of Lemma~\ref{lem:smoothness_reduction}]
 The gradient $-\nabla F(x)$ is computed in Sec.~\ref{ss:grad_local_entropy} to be $\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}$. Consider the term
@@
 
-\section{Connection to variational inference}
-\label{s:app:connection_variational}
-The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
-
-\paragraph{Comparison with SGLD:}
-A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
-Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
-We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
-
 \end{appendices}

### 2017-01-10 dnn/ploterrors.py

@@ -125,7 +125,6 @@ def plot_mnistfc():
     if opt['save']:
         plt.savefig('../doc/fig/mnistfc_valid.pdf', bbox_inches='tight')
 
-
 def plot_allcnn():
     r1, r2 = load(sorted(glob.glob('../results/nov_expts/dnn/allcnn/original/*.log')), 200), \
             load(sorted(glob.glob('../results/jan_expts/allcnn10/toplot/*.log')), 10)
@@ -198,6 +197,6 @@ def plot_allcnn():
     if opt['save']:
         plt.savefig('../doc/fig/allcnn_loss.pdf', bbox_inches='tight')
 
-plot_lenet()
-plot_mnistfc()
-plot_allcnn()
+# plot_lenet()
+# plot_mnistfc()
+# plot_allcnn()

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 \newcommand{\todo}[1]{{\color{gray}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[TODO]}}}}
+\newcommand{\fix}[2]{{\color{blue}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[FIX]}\color{blue}{#2} \par}}}
 \newcommand{\ignore}[1]{}
@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.
-Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state of the art techniques in terms of generalization performance and training time.
+\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state-of-the-art techniques in terms of generalization performance and training time.
 \end{abstract}
@@
 Local minima that generalize well and are discovered by gradient descent lie in ``wide valleys'' of the energy landscape, rather than in sharp isolated minima.
-\todo{For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the marginal likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.} Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
+For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the marginal likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
+Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.
@@
 $$
-The above is a log-partition function measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$; we call it ``local entropy'' in analogy to the free entropy used in statistical physics.
+The above is a log-partition function that measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$; we call it ``local entropy'' in analogy to the free entropy used in statistical physics.
 The $\entropysgd$ algorithm presented in this paper employs stochastic gradient Langevin dynamics (SGLD) to approximate the gradient of local entropy.
@@
 We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB)~\citep{marcus1993building} and character-level text prediction.
-Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD.
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up for RNNs).
 
@@
 The authors in~\citet{hochreiter1997flat} introduce a similar technique as our work, they introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
-\pc{Their analysis of the generalization performance (see their Sec.
+\fix{Their analysis of the generalization performance (see their Sec.
 A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{one sentence about the similarities} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
@@
 
-In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it resembles continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems.
-These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{,DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
-Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.
-%\pc{Let us also note that motivated from continuation,~\citet{gulcehre2016mollifying} insert annealed additive noise in the activation function to approximately convolve the original objective function.}{This paper is not related to our work, it is related to continuation.
+\todo{In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it resembles continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems.
+These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
+Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.} %\pc{Let us also note that motivated from continuation,~\citet{gulcehre2016mollifying} insert annealed additive noise in the activation function to approximately convolve the original objective function.}{This paper is not related to our work, it is related to continuation.
 I don't see any reason to add this sentence}
@@
 
+\begin{remark}
+discuss the eigenvalue assumption here
+\end{remark}
+
 \section{Experiments}
@@
 
-\textbf{Computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
+\textbf{Computational complexity:} \pc{Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
 We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
@@
 Note that $L = 1$ for SGD or Adam.
+The number of effective epochs
+
+}{add wall-clock times}
 
@@
 
-As noted in Sec.~\ref{ss:alg}, we can easily wrap Alg.~\ref{alg:entropysgd} inside other variants of SGD such as Adam; indeed this involves simply supplying the local entropy gradient in place of the usual back-propagated gradient.
-For this example, we construct $\entropyadam$ which is simply Adam that uses the local entropy gradient (computed via SGLD).
-We run $\entropyadam$ for $5$ epochs with $L = 5$ with an initial learning rate of $0.01$ that decreases by a factor of $2$ at each epoch.
+As noted in Sec.~\ref{ss:alg}, we can easily wrap Alg.~\ref{alg:entropysgd} inside other variants of SGD such as Adam; indeed this involves simply substituting the local entropy gradient in place of the usual back-propagated gradient.
+For this experiment, we construct $\entropyadam$ which is equivalent to Adam that uses the local entropy gradient (computed via SGLD).
+We run $\entropyadam$ for $5$ epochs with $L = 5$ and a fixed $\g = 0.01$ with an initial learning rate of $0.01$ that decreases by a factor of $2$ at each epoch.
 Note that this again results in only $25$ effective epochs.
-We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{this is a place-holder} over $4$ independent runs which is slightly better than the baseline in half as much wall-clock time.
+We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{this is a place-holder} over $4$ independent runs which is slightly better than the baseline, in half as much wall-clock time.
 We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
 
-Let us note that tuning the momentum hyper-parameter was crucial to getting good results on RNNs.
-While the SGD baseline on $\ptblstm$ does not use momentum and in fact does not train well with momentum, we set it to $0.5$ for $\entropysgd$.
-On the other hand, the baseline for $\charlstm$ was trained with Adam with a $\b_1 = 0.9$ and we set $\b_1 = 0.5$ for $\entropyadam$.
-In contrast to this observation on RNNs, our experiments on CNNs above always used a momentum of $0.9$.
+Tuning the momentum in $\entropysgd$ was crucial to getting good results on RNNs.
+While the SGD baseline on $\ptblstm$ does not use momentum (and in fact, does not train well with momentum) we used a momentum of $0.5$ for $\entropysgd$.
+On the other hand, the baseline for $\charlstm$ was trained with Adam with $\b_1 = 0.9$ ($\b_1$ in Adam controls the moving average of the gradient) while we set $\b_1 = 0.5$ for $\entropyadam$.
+In contrast to this observation about RNNs, all our experiments on CNNs used a momentum of $0.9$.
+\pc{This is also corroborated by monitoring the angle between the local entropy gradient and the vanilla SGD gradient during training.
+This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former.}{}
 
@@
 Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study.
-Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in wall-clock time for CNNs.
+Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in \fix{wall-clock time for CNNs}{}.
 Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}

### 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex

@@
 \newcommand{\todo}[1]{{\color{gray}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[TODO]}}}}
-\newcommand{\fix}[2]{{\color{blue}#1}\marginpar{\tiny\noindent{\raggedright{\color{blue}[FIX]}\color{blue}{#2} \par}}}
 \newcommand{\ignore}[1]{}
@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state-of-the-art techniques in terms of generalization performance and training time.
+We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.
+Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state of the art techniques in terms of generalization performance and training time.
 \end{abstract}
@@
 $$
-The above is a log-partition function that measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$; we call it ``local entropy'' in analogy to the free entropy used in statistical physics.
+The above is a log-partition function measures both the depth of a valley at a location  $x \in \reals^n$, and its flatness through the entropy of $f(x')$; we call it ``local entropy'' in analogy to the free entropy used in statistical physics.
 The $\entropysgd$ algorithm presented in this paper employs stochastic gradient Langevin dynamics (SGLD) to approximate the gradient of local entropy.
@@
 We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB)~\citep{marcus1993building} and character-level text prediction.
-Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up for RNNs).
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD.
 
@@
 The authors in~\citet{hochreiter1997flat} introduce a similar technique as our work, they introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
-\fix{Their analysis of the generalization performance (see their Sec.
+\pc{Their analysis of the generalization performance (see their Sec.
 A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{one sentence about the similarities} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
@@
 \todo{In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it resembles continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems.
-These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
+These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{,DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
 Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.} %\pc{Let us also note that motivated from continuation,~\citet{gulcehre2016mollifying} insert annealed additive noise in the activation function to approximately convolve the original objective function.}{This paper is not related to our work, it is related to continuation.
@@
 
-\begin{remark}
-discuss the eigenvalue assumption here
-\end{remark}
-
 \section{Experiments}
@@
 
-\textbf{Computational complexity:} \pc{Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
+\textbf{Computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
 We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
@@
 Note that $L = 1$ for SGD or Adam.
-The number of effective epochs
-
-}{add wall-clock times}
 
@@
 Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study.
-Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in \fix{wall-clock time for CNNs}{}.
+Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in wall-clock time for CNNs.
 Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

--- 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
@@
 It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
-\pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which closely matches the results of the original authors.
+We obtained a word perplexity of $81.43 \pm 0.2$ on the validation set and $78.6 \pm 0.26$ on the test set with this setup, which matches the results of the original authors very closely.
 
@@
 In contrast to this observation about RNNs, all our experiments on CNNs used a momentum of $0.9$.
-\pc{This is also corroborated by monitoring the angle between the local entropy gradient and the vanilla SGD gradient during training.
-This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former.}{}
+\pc{In order to investigate this difficulty, we monitored the angle between the local entropy gradient and the vanilla SGD gradient during training.
+This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy's gradient is thus highly uncorrelated with the SGD gradient in this case}{}
 
@@
 Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}
+
+- one paragraph on the efficacy of SGD and why it matters
 

### 2017-01-11 dnn/plotlog.py

@@ -100,8 +100,8 @@ def plot_idx(fs, df,
         'Validation loss')
     plt.xlim([0,max_epochs])
 
-    # plt.figure(4)
-    # plt.clf()
-    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
-    #     'Test loss')
-    # plt.xlim([0,max_epochs])
+    plt.figure(4)
+    plt.clf()
+    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
+        'Test loss')
+    plt.xlim([0,max_epochs])

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip

--- 2017-01-10 doc/chaudhari.choromanska.ea.iclr17.tex
@@
 
-The authors in~\citet{hochreiter1997flat} introduce a similar technique as our work, they introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
-\fix{Their analysis of the generalization performance (see their Sec.
-A1) necessitates a few unrealistic assumptions whereas in our analysis, as will be shown, we use uniform stability~\citet{bousquet2002stability} to obtain a precise characterization of generalization.}{one sentence about the similarities} Furthermore, as they mention, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
-Their formulation depends on Hessian-vector products.
-It is unclear how this method could be scaled to modern deep networks.
-Motivated by the same final goal, viz., flat local minima, our approach eliminates these extraneous and difficult to control hyper-parameters and instead results in a scalable algorithm.
-Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
+Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} \fix{(their analysis necessitates a few unrealistic assumptions, see their Sec.
+A1)}{we might consider deleting this} that this leads to improved generalization.
+However, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems, \textcolor{red}{as discussed by the authors.} Their formulation finally depends on Hessian-vector products.
+Driven by the same goal of encouraging the optimizer to favor flat regions in the opimization landscape, we however formulate a different objective function based on local entropy and obtain a more scalable algorithm which does not rely on extraneous and difficult to control hyper-parameters.
+\fix{Despite fundamental differences, both approaches...
+}{one sentence about the similarities} Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
 The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.
@@
 
-\pc{If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters.
+If we focus on the inner loop of the algorithm, SGLD updates compute the average gradient (with Langevin noise) in a neighborhood of the parameters while maintaining the Polyak average of the new parameters.
 Such an interpretation is very close to averaged SGD of~\citet{polyak1992acceleration} and~\citet{bottou2012stochastic} and worth further study.
-Moreover, our experiments show that $\entropysgd$ trains significantly faster than SGD for recurrent networks with relatively minor gains in \fix{wall-clock time for CNNs}{}.
-Estimating the gradient of local entropy cheaply with few SGLD iterations, or more interestingly, by using a smaller network in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.}{don't want to give away the farm}
+Our experiments show that while $\entropysgd$ trains significantly faster than SGD for recurrent networks, it gets relatively minor gains in terms of wall-clock time for CNNs.
+Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.
 

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
HS97 para

@@
 
-Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the width of local minima and training loss and show using the Gibbs formalism~\citep{haussler1997mutual} \fix{(their analysis necessitates a few unrealistic assumptions, see their Sec.
-A1)}{we might consider deleting this} that this leads to improved generalization.
-However, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems, \textcolor{red}{as discussed by the authors.} Their formulation finally depends on Hessian-vector products.
-Driven by the same goal of encouraging the optimizer to favor flat regions in the opimization landscape, we however formulate a different objective function based on local entropy and obtain a more scalable algorithm which does not rely on extraneous and difficult to control hyper-parameters.
-\fix{Despite fundamental differences, both approaches...
-}{one sentence about the similarities} Also recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
+Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
+As the authors discuss, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
+Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley) and also results in a more scalable algorithm.
+The scope $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{ss:scoping}, this provides significant computational advantages.
+Let us note that the Gibbs formalism introduced in~\citet{hochreiter1997flat} seems to be a very promising direction to understanding generalization in deep networks, it bears supercial similarities to the uniform stability analysis~\citep{haussler1997mutual} we employ.
+Recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
 The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.
@@
 It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
-\pc{We obtained a word perplexity of $81.125$ on the validation set and $78.232$}{std-dev} on the test set with this setup, which closely matches the results of the original authors.
+We obtained a word perplexity of $81.43 \pm 0.2$ on the validation set and $78.6 \pm 0.26$ on the test set with this setup, which closely matches the results of the original authors.
 
-We run $\entropysgd$ on $\ptblstm$ for $5$ epochs with $L = 5$; note that this results in only $25$ effective epochs.
+We run $\entropysgd$ on $\ptblstm$ for $5$ epochs with $L = 5$, note that this results in only $25$ effective epochs.
 We do not use scoping for this network and instead fix $\g = 10^{-3}$.
@@
 Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.
+
+- one paragraph on the efficacy of SGD and why it matters
 

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB)~\citep{marcus1993building} and character-level text prediction.
-Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up for RNNs).
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up on RNNs).
 
@@
 The scope $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{ss:scoping}, this provides significant computational advantages.
-Let us note that the Gibbs formalism introduced in~\citet{hochreiter1997flat} seems to be a very promising direction to understanding generalization in deep networks, it bears supercial similarities to the uniform stability analysis~\citep{haussler1997mutual} we employ.
-Recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
+Let us note that the Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} seems to be a very promising direction to understanding generalization in deep networks.
+%, it bears supercial similarities to the uniform stability analysis~\citep{haussler1997mutual} we employ.
+%Recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
 The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.
 
-\todo{In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it resembles continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems.
-These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions~\citep{DBLP:conf/icml/HazanLS16,mobahi2015link}; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
-Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.} %\pc{Let us also note that motivated from continuation,~\citet{gulcehre2016mollifying} insert annealed additive noise in the activation function to approximately convolve the original objective function.}{This paper is not related to our work, it is related to continuation.
-I don't see any reason to add this sentence}
+In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it is related to continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems~\citep{DBLP:conf/icml/HazanLS16,mobahi2015link} or methods such as~\citet{gulcehre2016mollifying} that \pc{perturb the weights with random noise to average the gradient}{cite some other paper instead}.
+Smoothing the
+
+More principled techniques based on continuation
+
+These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
+
+Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.
 

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex

@@
 \end{lemma}
-The local entropy objective is thus smoother than the original objective while the modified objective in~\eqref{eq:entropysgd_objective} results in a factor of $\r + \f{1}{1 + \g^{-1}\ c}$ in the above lemma.
+The local entropy objective is thus smoother than the original objective.
 
@@
 
-\section{Connection to variational inference}
-\label{s:app:connection_variational}
-The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
-
 \begin{remark}
@@
 \end{remark}
+
+\subsection{Connection to variational inference}
+\label{ss:connection_variational}
+The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
 

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip

@@
 We use an LSTM consisting of two layers of $128$ hidden units unrolled for $50$ time steps and a vocabulary of size $87$.
-We train the baseline with Adam for $50$ epochs with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of \pc{$1.227 \pm 0.006$}{this is a place-holder}.
+We train the baseline with Adam for $50$ epochs with an initial learning rate of $0.002$ that decays by a factor of $2$ after every $5$ epochs to obtain a validation perplexity of $1.224 \pm 0.008$ and a test perplexity of $1.226 \pm 0.01$.
 
@@
 Note that this again results in only $25$ effective epochs.
-We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of \pc{$1.20 \pm 00$}{this is a place-holder} over $4$ independent runs which is slightly better than the baseline, in half as much wall-clock time.
+We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of $1.217 \pm 0.005$ over $4$ independent runs which is slightly better than the baseline, in half as much wall-clock time.
 We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
@@
 \pc{In order to investigate this difficulty, we monitored the angle between the local entropy gradient and the vanilla SGD gradient during training.
-This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy's gradient is thus highly uncorrelated with the SGD gradient in this case}{}
+This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy gradient is highly uncorrelated with the SGD gradient in this case.}{}
 

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip on sec. 4

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably with state-of-the-art techniques in terms of generalization performance and training time.
+\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization performance and training time.
 \end{abstract}
@@
 For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the marginal likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
-Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters --- all of which are widely-used techniques to achieve good generalization.
-This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning --- without being explicitly tailored to do so --- seem to discover exactly that.
+Almost-flat regions of the energy landscape are robust to data perturbations, noise in the activations, as well as perturbations of the parameters, all of which are widely-used techniques to achieve good generalization.
+This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning seem to discover exactly that --- without being explicitly tailored to do so.
 
@@
 The authors use belief propagation to estimate local entropy for simpler models considered there.
-A related work in this context is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker to the ensemble average.
+A related work in this context is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker from the ensemble average.
 Such an ensemble training procedure enables improved generalization by ensuring that different workers land in the same wide valley and indeed, it turns out to be closely related to the replica theoretic analysis of~\citep{baldassi2016unreasonable}.
@@
 Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
-As the authors discuss, the effect of hyper-parameters for the constraints they use is intricately tied together and they are difficult to choose even for small problems.
+As the authors discuss, the effect of hyper-parameters for the constraints is intricately tied together and they are difficult to choose even for small problems.
 Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley) and also results in a more scalable algorithm.
-The scope $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{ss:scoping}, this provides significant computational advantages.
-Let us note that the Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} seems to be a very promising direction to understanding generalization in deep networks.
+The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{ss:scoping}, this also provides significant computational advantages.
+The Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is however a very promising direction to understanding generalization in deep networks in the context of this paper.
 %, it bears supercial similarities to the uniform stability analysis~\citep{haussler1997mutual} we employ.
@@
 
-In the sense that local entropy results in a smoother energy landscape (cf.\@ Sec.~\ref{ss:theoretical_properties}), it is related to continuation methods~\citep{allgower2012numerical} that convolve the loss function to solve sequentially refined optimization problems~\citep{DBLP:conf/icml/HazanLS16,mobahi2015link} or methods such as~\citet{gulcehre2016mollifying} that \pc{perturb the weights with random noise to average the gradient}{cite some other paper instead}.
-Smoothing the
-
-More principled techniques based on continuation
-
-These techniques however require a special class of loss functions where successively smaller convolutions result in minimizers that lie close to each other or closed-form expressions of smoothened loss functions; these may not be possible for deep networks with isolated minima and highly-nonlinear loss functions.
-
-Our modified objective instead weighs each location by the local entropy in its neighborhood and does not require stringent assumptions on the objective, it is also applicable to full-scale deep networks.
+Another related area is homotopy continuation methods that convolve the loss function to solve sequentially refined optimization problems~\citep{allgower2012numerical,mobahi2015link} or methods that perturb the weights or activations to average the gradient~\citep{gulcehre2016mollifying} with an aim to smooth the rugged energy landscape.
+Such a smoothing is however different from local entropy.
+For instance, the latter places more weight on wide local minima even if they are much shallower than the global minimum (cf.
+Fig.~\ref{fig:entropyfig}) which cannot be obtained by smoothing.
+In fact, smoothing can introduce an artifical minimum between two sharp valleys connected by a high energy barrier which is detrimental to generalization.
+Please refer to Sec.~\ref{ss:connection_variational} for a more elaborate discussion.
+In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
+Local entropy exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}; our experiments on the eigenspectrum of the Hessian are a step to showing the existence of such wide minima in modern deep networks.
 
@@
 
-Let us note that although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
+Let us discuss the algorithm in a little more detail.
+Although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
 In our experiments, we use Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
+%
+We use exponential averaging to estimate $\mu$ in the SGLD loop (Line 6) with $\a = 0.75$ so as to put more weight on the later samples, this is akin to a burn-in of SGLD iterations.
 
-\todo{
-We set the number of SGLD iterations $L$ to $5$-$20$ depending upon the complexity of the dataset, e.g., for experiments on convolutional networks on MNIST and CIFAR-10, we use $L = 20$ while for experiments on a fully-connected network on MNIST $L=5$ works well (cf.
-Sec.~\ref{s:expt}).
-We perform exponential averaging of the Langevin iterates $x'$ to estimate $\mu$ with a parameter $\a = 0.9$ so as to put more weight on later samples.
-In order to quickly transition to the Langevin dynamics phase of SGLD and ensure the we are sampling from the posterior even for small values of $L$, we set the learning rate $\eta'$ on line $5$ to be small factor of the outer SGD's learning rate: $\eta' = 0.1\ \eta$~\citep{welling2011bayesian}.}
+We set the number of SGLD iterations to $L = [5, 20]$ depending upon the complexity of the dataset.
+The learning rate $\eta'$ is fixed for all our experiments to values between $\eta' \in\ [0.1, 1]$.
+We found that annealing $\eta'$ (for instance, setting it to be the same as the outer learning rate $\eta$) is detrimental; indeed a small learning rate leads to poor estimates of local entropy towards the end of training where they are most needed.
+The parameter $\e$ (Line 5) is called the ``thermal noise'' in SGLD, we fix this to $\e = 10^{-4}$.
+Having thus fixed $L, \e$ and $\eta'$, an effective heuristic to tune the remaining parameter $\g$ is to match the magnitude of the gradient of the local entropy term, viz.
+$\g\ (x - \mu)$, to the gradient for vanilla SGD, viz.
+$m^{-1}\ \sum_{i=1}^m\ \nabla_x f(x; \xi_{\ell_i})$.
 
@@
 
-Our scoping mechanism introduced in Sec.~\ref{ss:scoping} increases the value of $\g$ as training progresses in order to focus on more local neighborhoods.
+The scoping parameter $\g$ is fixed in Alg.~\ref{alg:entropysgd}.
+For large values of $\g$, the SGLD updates focus on local neighborhoods of the current parameters $x$ while small values of $\g$ allow the inner SGLD to explore further away from $x$.
+A ``reverse-annealing'' of the scope $\g$, i.e.
+increasing $\g$ as training progresses, thus has the effect of exploring the energy landscape on progressively finer scales.
+We call this process ``scoping'' and use a simple exponential schedule given by
+$$
+    \g(t) = \g_0\ \rbrac{1 + \g_1}^t;
+$$
+for the $t^{\trm{th}}$ parameter update.
 
-The ``reverse-annealing'' of the scope $\g$ typically interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
+Scoping of $\g$ typically interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
 In practice, we therefore scale the local entropy gradient by $\g$ before the update and modify the line to read
@@
 We found this to be very effective in simplifying hyper-parameter tuning.
+In general, we set $\g_0$ to be very close to zero, pick a large value of $\eta$ and set $\g_1$ to be such that the magnitude of the local entropy gradient is comparable to that of SGD.
+We use much larger learning rates than SGD in our experiments to enable fast progress in the smooth energy landscape with low values of $\g$, we can do so because the local entropy gradient is less noisier than the original back-propagated gradient.
 
@@
 For the last network, we sacrifice the exact computation and instead approximate the Hessian of a large deep network.
-We note that recovering an approximate Hessian from Hessian-vector products~\citep{pearlmutter1994fast} could be a viable strategy for medium-scale networks.
+We note that recovering an approximate Hessian from Hessian-vector products~\citep{pearlmutter1994fast} could be a viable strategy for large networks.
 
@@
 \centering
-    \begin{subfigure}[t]{0.44 \textwidth}
+    \begin{subfigure}[t]{0.4\textwidth}
         \includegraphics[width=1.08\textwidth]{mnistfc_hessian.pdf}
@@
     \hspace{0.2in}
-    \begin{subfigure}[t]{0.44 \textwidth}
+    \begin{subfigure}[t]{0.4 \textwidth}
         \includegraphics[width=\textwidth]{charlstm_hessian.pdf}
@@
         \centering
-        \begin{subfigure}[b]{0.44 \textwidth}
+        \begin{subfigure}[b]{0.4 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dm.pdf}
@@
         \hspace{0.2in}
-        \begin{subfigure}[b]{0.44 \textwidth}
+        \begin{subfigure}[b]{0.4 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dp.pdf}
@@
 
-\textbf{Computational complexity:} \pc{Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
+\textbf{Remark on the computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
 We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
 by multiplying the abscissae by a factor of $L$.
-Note that $L = 1$ for SGD or Adam.
-The number of effective epochs
-
-}{add wall-clock times}
+(define $L = 1$ for SGD or Adam).
+This is a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
 
@@
 Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
-It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $8.28 \pm 0.11 \%$ in the same number of effective epochs.
+It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $8.22 \pm 0.11 \%$ in the same number of effective epochs.
 
@@
 \label{ss:expt:rnn}
+
+\subsubsection{Penn Tree Bank}
 
@@
 
+\subsubsection{char-LSTM on War and Peace}
+
 Next, we train an LSTM to perform character-level text-prediction.
@@
 
+\begin{figure}[htp!]
+\centering
+    \begin{subfigure}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=1.03\textwidth]{ptblstm_valid.pdf}
+        \caption{\small $\ptblstm$: Validation, test perplexity}
+        \label{fig:ptb_valid}
+    \end{subfigure}
+    \hspace{0.2in}
+    \begin{subfigure}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{charlstm_valid.pdf}
+        \caption{\small $\charlstm$: Validation perplexity}
+        \label{fig:char_valid}
+    \end{subfigure}
+\caption{\small Comparison of $\entropysgd$ vs. SGD / Adam on RNNs}
+\label{fig:rnn}
+\end{figure}
+
 Tuning the momentum in $\entropysgd$ was crucial to getting good results on RNNs.
@@
 This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy gradient is highly uncorrelated with the SGD gradient in this case.}{}
-
-\todo{the two plots below are place holders}
-\begin{figure}[htp!]
-\centering
-    \begin{subfigure}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{allcnn_valid.pdf}
-        \caption{\small $\ptblstm$: Validation, test perplexity}
-        \label{fig:ptb_valid}
-    \end{subfigure}
-    \hspace{0.2in}
-    \begin{subfigure}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=1.02\textwidth]{allcnn_loss.pdf}
-        \caption{\small $\charlstm$: Validation perplexity}
-        \label{fig:char_valid}
-    \end{subfigure}
-\caption{\small Comparison of $\entropysgd$ vs. SGD / Adam on RNNs}
-\label{fig:rnn}
-\end{figure}
 
@@
 Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.
-
-- one paragraph on the efficacy of SGD and why it matters
 
@@
 Concretely, the posterior expectation $\E\sqbrac{g(x)}$ is given by
-$
+$$
     \E\sqbrac{g(x)} \approx \f{\sum_{s=1}^t \eta_t\ g(x_t)}{\sum_{s=1}^t \eta_t};
-$
+$$
 which is the average computed by weighing each sample by the corresponding learning rate in~\eqref{eq:sgld_bayesian}.

### 2017-01-11 dnn/plotlog.py

@@ -87,21 +87,21 @@ def plot_idx(fs, df,
     #     'Training error')
     # plt.xlim([0,max_epochs])
 
-    # plt.figure(3)
-    # plt.clf()
-    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].miss,
-    #     'Test error')
-    # plt.xlim([0,max_epochs])
-
-    # lstm
     plt.figure(3)
     plt.clf()
-    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].loss,
-        'Validation loss')
+    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].miss,
+        'Test error')
     plt.xlim([0,max_epochs])
 
-    plt.figure(4)
-    plt.clf()
-    helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
-        'Test loss')
-    plt.xlim([0,max_epochs])
+    # lstm
+    # plt.figure(3)
+    # plt.clf()
+    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 0) & (df['batch'] == 0)].loss,
+    #     'Validation loss')
+    # plt.xlim([0,max_epochs])
+
+    # plt.figure(4)
+    # plt.clf()
+    # helper(df[(df['epoch'] <= max_epochs) & (df['tv'] == 2) & (df['batch'] == 0)].loss,
+    #     'Test loss')
+    # plt.xlim([0,max_epochs])

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip on sec. 4.3

@@
 Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
-Note that $-F(x, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; indeed, $x_{\textrm{robust}}$ has a higher local entropy than $x_{\textrm{non-robust}}$.
+Note that $-F(x, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
 
@@
 We found that annealing $\eta'$ (for instance, setting it to be the same as the outer learning rate $\eta$) is detrimental; indeed a small learning rate leads to poor estimates of local entropy towards the end of training where they are most needed.
-The parameter $\e$ (Line 5) is called the ``thermal noise'' in SGLD, we fix this to $\e = 10^{-4}$.
+The parameter $\e$ in SGLD on Line 5 is the thermal noise and we fix this to $\e = 10^{-4}$.
 Having thus fixed $L, \e$ and $\eta'$, an effective heuristic to tune the remaining parameter $\g$ is to match the magnitude of the gradient of the local entropy term, viz.
 $\g\ (x - \mu)$, to the gradient for vanilla SGD, viz.
-$m^{-1}\ \sum_{i=1}^m\ \nabla_x f(x; \xi_{\ell_i})$.
+$m^{-1}\ \sum_{i=1}^m\ \nabla_x f(x;\ \xi_{\ell_i})$.
 
@@
 The scoping parameter $\g$ is fixed in Alg.~\ref{alg:entropysgd}.
-For large values of $\g$, the SGLD updates focus on local neighborhoods of the current parameters $x$ while small values of $\g$ allow the inner SGLD to explore further away from $x$.
+For large values of $\g$, the SGLD updates happen in a small neighborhood of the current parameters $x$ while low values of $\g$ allow the inner SGLD to explore further away from $x$.
 A ``reverse-annealing'' of the scope $\g$, i.e.
@@
 $$
-for the $t^{\trm{th}}$ parameter update.
+for the $t^{\trm{th}}$ parameter update. We have experimented with linear, quadratic and bounded exponential ($\g_0 \rbrac{1 - e^{-\tau t}}$) scoping schedules with similar qualitative results.
 
-Scoping of $\g$ typically interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
+Scoping of $\g$ unfortunately interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
 In practice, we therefore scale the local entropy gradient by $\g$ before the update and modify the line to read

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
wip on sec. 4.3

@@
 In fact, smoothing can introduce an artifical minimum between two sharp valleys connected by a high energy barrier which is detrimental to generalization.
-Please refer to Sec.~\ref{ss:connection_variational} for a more elaborate discussion.
+Please refer to Sec.~\ref{ss:app:connection_variational} for a more elaborate discussion.
 In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
@@
 $$
-for the $t^{\trm{th}}$ parameter update. We have experimented with linear, quadratic and bounded exponential ($\g_0 \rbrac{1 - e^{-\tau t}}$) scoping schedules with similar qualitative results.
+for the $t^{\trm{th}}$ parameter update.
+We have experimented with linear, quadratic and bounded exponential ($\g_0 \rbrac{1 - e^{-\tau t}}$) scoping schedules and obtained qualitatively similar results.
 
@@
 
-\pc{We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
-Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).}{clarify that the analysis here is only a sketch, it is not rigorous}
+We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
+Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).
 
@@
 The local entropy objective is thus smoother than the original objective.
-
+%
 Let us now obtain a bound on the improvement in generalization error.
@@
 \end{remark}
-
-\subsection{Connection to variational inference}
-\label{ss:connection_variational}
-The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
 
@@
 These two observations are a direct consequence of the additional smoothness of the local entropy landscape.
-
-\paragraph{Comparison with SGLD:}
-\todo{A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
-Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
-We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.}
 
@@
 
+\todo{
+\subsection{Connection to variational inference}
+\label{ss:app:connection_variational}
+The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
+
+\paragraph{Comparison with SGLD:}
+A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
+Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
+We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
+}
+
 \end{appendices}

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
new plots for allcnn

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization performance and training time.
+\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
 \end{abstract}
@@
 In fact, smoothing can introduce an artifical minimum between two sharp valleys connected by a high energy barrier which is detrimental to generalization.
-Please refer to Sec.~\ref{ss:app:connection_variational} for a more elaborate discussion.
+Please refer to Sec.~\ref{s:app:connection_variational} for a more elaborate discussion.
 In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
@@
         \centering
-        \includegraphics[width=1.02\textwidth]{allcnn_loss.pdf}
+        \includegraphics[width=1.06\textwidth]{allcnn_loss.pdf}
         \caption{\small $\allcnn$: Training loss}
@@
 We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
-We obtain an average error of $8.30 \pm 0.1\%$ over $5$ runs of $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
+We obtain an average error of $7.71 \pm 0.19\%$ in $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
 Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.
@@
 Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
-It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $8.22 \pm 0.11 \%$ in the same number of effective epochs.
+It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $7.81 \pm 0.09 \%$ in the same number of effective epochs.
 
@@
 \todo{
-\subsection{Connection to variational inference}
-\label{ss:app:connection_variational}
+\section{Connection to variational inference}
+\label{s:app:connection_variational}
 The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.

### 2017-01-11 dnn/ploterrors.py

@@ -134,8 +134,8 @@ def plot_mnistfc():
         plt.savefig('../doc/fig/mnistfc_valid.pdf', bbox_inches='tight')
 
 def plot_allcnn():
-    r1, r2 = load(sorted(glob.glob('../results/nov_expts/dnn/allcnn/original/*.log')), 200), \
-            load(sorted(glob.glob('../results/jan_expts/allcnn10/toplot/*.log')), 10)
+    r1, r2 = load(sorted(glob.glob('../results/jan_expts/allcnn10/toplotbn/*\"L\":0*.log')), 200), \
+            load(sorted(glob.glob('../results/jan_expts/allcnn10/toplotbn/*\"L\":20*.log')), 10)
 
     # validation error
     fig = plt.figure(3, figsize=(8,7))
@@ -143,8 +143,8 @@ def plot_allcnn():
     ax = fig.add_subplot(111)
     #plt.title(r'All-CNN-BN: Validation error')
 
-    v1 = np.array([smooth(r1['valid'].ix[i]) for i in range(5)])
-    v2 = np.array([smooth(r2['valid'].ix[i]) for i in range(3)])
+    v1 = np.array([smooth(r1['valid'].ix[i]) for i in range(2)])
+    v2 = np.array([smooth(r2['valid'].ix[i]) for i in range(2)])
     sns.tsplot(v1,
             condition=r'SGD', rasterized=True, color='k')
     sns.tsplot(v2, time=np.arange(20,220,20),
@@ -158,13 +158,13 @@ def plot_allcnn():
     plt.xlabel(r'Epochs $\times$ L')
     plt.ylabel(r'\% Error')
 
-    plt.plot([200], [8.4], 'o', c='k', markersize=10)
-    plt.plot([200], [8.2], 'o', c='r', markersize=10)
-    plt.plot(range(200), 8.28*np.ones(200), 'r--', lw=1)
+    plt.plot([194], [7.71], 'o', c='k', markersize=10)
+    plt.plot([160], [7.81], 'o', c='r', markersize=10)
+    plt.plot(range(200), 7.81*np.ones(200), 'r--', lw=1)
 
-    ax.text(160, 10.0, r'$8.30\%$', fontsize=fsz,
+    ax.text(180, 9.0, r'$7.71\%$', fontsize=fsz,
             verticalalignment='center', color='k')
[…]

### 2017-01-11 dnn/ploterrors.py
new allcnn plots

@@ -151,9 +151,9 @@ def plot_allcnn():
             condition=r'Entropy-SGD', rasterized=True, color='r')
     plt.grid('on')
 
-    plt.ylim([5, 25])
+    plt.ylim([5, 20])
     plt.xlim([0, 200])
-    yticks = [5, 10, 15, 20, 25]
+    yticks = [5, 10, 15, 20]
     plt.yticks(yticks, [str(y) for y in yticks])
     plt.xlabel(r'Epochs $\times$ L')
     plt.ylabel(r'\% Error')

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
first rough draft ready, wip on connection to SVI

@@
 % save space
-\setlength{\floatsep}{0.0in}
-\setlength{\textfloatsep}{0.0in}
+\setlength{\floatsep}{0.05in}
+\setlength{\textfloatsep}{0.05in}
 \setlength{\intextsep}{0.05in}
@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-\fix{We prove that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
+\fix{We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth out} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
 \end{abstract}
@@
 \end{definition}
-The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
-Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
-Note that $-F(x, \g)$ has a global minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
+\pc{The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
+Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.}{talk about small gamma and large gamma valleys} Note that $-F(x, \g)$ has a \emph{global} minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
 
@@
 %
-We use exponential averaging to estimate $\mu$ in the SGLD loop (Line 6) with $\a = 0.75$ so as to put more weight on the later samples, this is akin to a burn-in of SGLD iterations.
+We use exponential averaging to estimate $\mu$ in the SGLD loop (Line 6) with $\a = 0.75$ so as to put more weight on the later samples, this is akin to a burn-in in standard SGLD.
 
@@
 $$
-We found this to be very effective in simplifying hyper-parameter tuning.
-In general, we set $\g_0$ to be very close to zero, pick a large value of $\eta$ and set $\g_1$ to be such that the magnitude of the local entropy gradient is comparable to that of SGD.
-We use much larger learning rates than SGD in our experiments to enable fast progress in the smooth energy landscape with low values of $\g$, we can do so because the local entropy gradient is less noisier than the original back-propagated gradient.
+Our general strategy during hyper-parameter tuning is to set the initial scope $\g_0$ to be very small, pick a large value of $\eta$ and set $\g_1$ to be such that the magnitude of the local entropy gradient is comparable to that of SGD.
+We can use much larger learning rates than SGD in our experiments because the local entropy gradient is less noisier than the original back-propagated gradient.
 
@@
 
-We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective} if trained for the same number of iterations.
-Proofs are deferred to the Appendix (see Sec.~\ref{s:app:proofs}).
-
+We can show that $\entropysgd$ results in a smoother loss function and obtains better generalization error than the original objective~\eqref{eq:dnn_objective}.
 With some overload of notation, we assume that the original loss $f(x)$ is $\b$-smooth, i.e., for all $x,y \in \reals^n$, we have $\norm{\nabla f(x) - \nabla f(y)} \leq \b\ \norm{x-y}$.
@@
 \end{lemma}
-The local entropy objective is thus smoother than the original objective.
+Please see Sec.~\ref{s:app:proofs} for the proof. The local entropy objective is thus smoother than the original objective.
 %
@@
 $$
-where the left hand side is the generalization error: it is the difference between the empirical loss $R_\Xi(x) := \f{1}{N}\ \sum_{k=1}^N\ f(x, \xi_k)$ and the population loss $R(x) := \E_{\xi}\ f(x, \xi)$.
-
+where the left hand side is the generalization error: it is the difference between the empirical loss $R_\Xi(x) := N^{-1} \sum_{k=1}^N\ f(x, \xi_k)$ and the population loss $R(x) := \E_{\xi}\ f(x, \xi)$.
+%
 We now employ the following theorem that bounds the stability of an optimization algorithm through the smoothness of its loss function and the number of iterations on the training set.
@@
 \end{equation}
-which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$.
-
+which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$ if they both converge within $T$ parameter updates.
+\pc{However, from our experiments, $\entropysgd$ converges in roughly $L$ times fewer parameter updates than SGD, using this we can draw out the dependency on the number of Langevin iterations to get
+$$
+    \e_{\ \entropysgd} \lessapprox L^{-\f{c\ \b}{\g}}\ \e_{\ \textrm{SGD}}.
+$$
 As an aside, it is easy to see from the proof of Lem.~\ref{lem:smoothness_reduction} that for a convex loss function $f(x)$, adding the local entropy term does not change the minimizer of the original problem.
@@
 \begin{remark}
-discuss the eigenvalue assumption here
+The above analysis hinges upon an assumption that the Hessian $\nabla^2 f(x)$ does not have eigenvalues in the set $[-2\g-c, c]$ for a constant $c > 0$.
+This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.~\ref{fig:lenet_hessian} has a large fraction of its eigenvalues near zero.
+Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} assumes global conditions on the smoothness of the loss function~\citep{hardt2015train}; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
 \end{remark}
+}{}
 
@@
 \centering
-    \begin{subfigure}[t]{0.4\textwidth}
+    \begin{subfigure}[t]{0.45\textwidth}
         \centering
@@
     \hspace{0.2in}
-    \begin{subfigure}[t]{0.4\textwidth}
+    \begin{subfigure}[t]{0.45\textwidth}
         \centering
@@
 
+We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}.
+We consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer and all other architectural parameters such as filter-sizes and weight decay are kept the same.
+We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
+We obtain an average error of $7.71 \pm 0.19\%$ in $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
+Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.
+
+We train $\entropysgd$ with $L = 20$ for $10$ epochs with the original dropout of $0.5$.
+The initial learning rate of the outer loop is set to $\eta = 1$ and drops by a factor of $5$ every $4$ epochs, while the learning rate of the SGLD updates is fixed to $\eta' = 0.1$ with thermal noise $\e = 10^{-4}$.
+As the scoping scheme, we set the initial value of the scope to $\g_0 = 0.03$ which increases by a factor of $1.001$ after each parameter update.
+%
+Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
+It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $7.81 \pm 0.09 \%$ in about $160$ effective epochs.
+
+We see almost no plateauing of training loss or validation error for $\entropysgd$ in Fig.~\ref{fig:allcnn_loss}; this trait is shared across different networks and datasets in our experiments and is a direct consequence of the additional smoothness of the local entropy landscape.
+
 \begin{figure}[htp!]
 \centering
-    \begin{subfigure}[b]{0.4\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \centering
@@
     \hspace{0.2in}
-    \begin{subfigure}[b]{0.4\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \centering
@@
 
-We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}.
-We consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer and all other architectural parameters such as filter-sizes and weight decay are kept the same.
-We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
-We obtain an average error of $7.71 \pm 0.19\%$ in $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
-Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.
-
-We train $\entropysgd$ with $L = 20$ for $10$ epochs with the original dropout of $0.5$.
-The initial learning rate of the outer loop is set to $\eta = 1$ and drops by a factor of $5$ every $4$ epochs, while the learning rate of the SGLD updates is fixed to $\eta' = 0.1$ with thermal noise $\e = 10^{-4}$.
-As the scoping scheme, we set the initial value of the scope to $\g = 0.03$ which increases by a factor of $1.001$ after each parameter update.
-%
-Fig.~\ref{fig:allcnn} shows the training and validation error curves for $\entropysgd$ compared with SGD.
-It shows that local entropy performs as well as SGD on a large CNN; we obtain a validation error of $7.81 \pm 0.09 \%$ in the same number of effective epochs.
-
-We see almost no plateauing of training loss or validation error for $\entropysgd$ in Fig.~\ref{fig:allcnn_loss}; this trait is shared across different networks and datasets in our experiments.
-Moreover, while the magnitude of the gradients is similar for both SGD and $\entropysgd$ --- in fact, we use this as a way to tune SGLD's hyper-parameters $\eta'$ and $\g$ --- we can use significantly larger learning rates for $\entropysgd$.
-SGD with large learning rates, say $\eta = 1$ for the above network, would not train at all.
-These two observations are a direct consequence of the additional smoothness of the local entropy landscape.
-
 \subsection{Recurrent neural networks}
@@
 
-In order to test the efficacy of $\entropysgd$ on recurrent architectures, we train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction.
+We train an LSTM network on the Penn Tree Bank (PTB) dataset for word-level text prediction.
 This dataset contains about one million words divided into a training set of about $930,000$ words, a validation set of $74,000$ words and $82,000$ words with a vocabulary of size $10,000$.
-As a baseline, we follow the experiments in~\citet{zaremba2014recurrent}.
-Our network called $\ptblstm$ consists of two layers with $1500$ hidden units, each unrolled for $35$ time steps and a dropout of probability $0.65$; note that this is a large network with about $66$ million weights.
-It is trained with SGD (without momentum, like the original authors) for $55$ epochs with a learning rate $\eta = 1$ for the first $14$ epochs that is annealed by a factor of $1.15$ after every epoch.
-We obtained a word perplexity of $81.43 \pm 0.2$ on the validation set and $78.6 \pm 0.26$ on the test set with this setup, which closely matches the results of the original authors.
+Our network called $\ptblstm$ consists of two layers with $1500$ hidden units, each unrolled for $35$ time steps; note that this is a large network with about $66$ million weights.
+We recreated the training pipeline of~\citet{zaremba2014recurrent} for this network (SGD without momentum) and obtained a word perplexity of $81.43 \pm 0.2$ on the validation set and $78.6 \pm 0.26$ on the test set with this setup; these numbers closely match the results of the original authors.
 
@@
 We obtain a word perplexity of $80.116 \pm 0.069$ on the validation set and $77.656 \pm 0.171$ on the test set.
-As Fig.~\ref{fig:ptb_valid} shows, $\entropysgd$ trains significantly faster than SGD and also achieves a slightly better generalization perplexity.
+As Fig.~\ref{fig:ptb_valid} shows, $\entropysgd$ trains significantly faster than SGD ($25$ effective epochs vs.
+$55$ epochs of SGD) and also achieves a slightly better generalization perplexity.
 
@@
 
-As noted in Sec.~\ref{ss:alg}, we can easily wrap Alg.~\ref{alg:entropysgd} inside other variants of SGD such as Adam; indeed this involves simply substituting the local entropy gradient in place of the usual back-propagated gradient.
-For this experiment, we construct $\entropyadam$ which is equivalent to Adam that uses the local entropy gradient (computed via SGLD).
+As noted in Sec.~\ref{ss:alg}, we can easily wrap Alg.~\ref{alg:entropysgd} inside other variants of SGD such as Adam; this involves simply substituting the local entropy gradient in place of the usual back-propagated gradient.
+For this experiment, we constructed $\entropyadam$ which is equivalent to Adam with the local entropy gradient (which is computed via SGLD).
 We run $\entropyadam$ for $5$ epochs with $L = 5$ and a fixed $\g = 0.01$ with an initial learning rate of $0.01$ that decreases by a factor of $2$ at each epoch.
-Note that this again results in only $25$ effective epochs.
-We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of $1.217 \pm 0.005$ over $4$ independent runs which is slightly better than the baseline, in half as much wall-clock time.
+Note that this again results in only $25$ effective epochs, i.e.
+half as much wall-clock time as Adam.
+We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of $1.217 \pm 0.005$ over $4$ independent runs which is better than the baseline.
 We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
@@
 \centering
-    \begin{subfigure}[b]{0.4\textwidth}
+    \begin{subfigure}[b]{0.45\textwidth}
         \centering
@@
 
-\begin{proof}[Proof sketch of Lemma~\ref{lem:smoothness_reduction}]
+\begin{proof}[Proof of Lemma~\ref{lem:smoothness_reduction}]
 The gradient $-\nabla F(x)$ is computed in Sec.~\ref{ss:grad_local_entropy} to be $\g\ \rbrac{x - \ag{x';\ \minibatch{\ell}}}$. Consider the term

### 2017-01-11 hessian/entropyfig.py

@@ -60,11 +60,15 @@ f, s,_ = smooth(0, 1e-3)
 f = f + 1
 s = s - s.min()
 
+f2, s2,_ = smooth(0, 5e-5)
+s2 = s2 - s2.min()
+
 fig = plt.figure(1, figsize=(9,8))
 plt.clf()
 ax = fig.add_subplot(111)
 plt.plot(f, 'k-', label=r'$\mathrm{Original\ landscape}$', lw=1)
 plt.plot(s, 'indianred', label=r'$\mathrm{Negative\ local\ entropy}:\ \gamma = 0.001$', lw=2)
+plt.plot(s2, 'indianred', ls='--', label=r'$\mathrm{Negative\ local\ entropy}:\ \gamma = 0.00005$', lw=2)
 plt.legend(loc=2)
 #plt.title(r'$\mathrm{Local\ entropy\ energy\ landscape}$')
 plt.xticks([], [])

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
gitignore
added svi+sgld section

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-\fix{We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.}{smooth out} Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
+We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.
+Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
 \end{abstract}
@@
 We present experimental results on fully-connected and convolutional neural networks (CNNs) on the MNIST and CIFAR-10~\citep{krizhevsky2009learning} datasets and recurrent neural networks (RNNs) on the Penn Tree Bank dataset (PTB)~\citep{marcus1993building} and character-level text prediction.
-Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up on RNNs).
+Our experiments show that $\entropysgd$ scales to deep networks used in practice, obtains comparable generalization error as competitive baselines and also trains much more quickly than SGD (we get a $2\trm{x}$ speed-up over SGD on RNNs).
 
@@
 increasing $\g$ as training progresses, thus has the effect of exploring the energy landscape on progressively finer scales.
-We call this process ``scoping'' and use a simple exponential schedule given by
+This process is called ``scoping''~\citep{baldassi2016unreasonable} and use a simple exponential schedule given by
 $$
@@
 
-\textbf{Remark on the computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of iterations over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
+\textbf{Remark on the computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of passes over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
 We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
@@
 (define $L = 1$ for SGD or Adam).
-This is a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
+This is thus a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
 
@@
 We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
-We obtain an average error of $7.71 \pm 0.19\%$ in $200$ epochs vs.\ $9.08\%$ error in $350$ epochs as the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
+We obtain an average error of $7.71 \pm 0.19\%$ in $200$ epochs vs.\ $9.08\%$ error in $350$ epochs that the authors in~\citet{springenberg2014striving} report and this is thus a very competitive baseline for this network.
 Let us note that the best result in the literature on non-augmented CIFAR-10 is the ELU-network by~\citet{clevert2015fast} with $6.55\%$ test error.
@@
 
-We see almost no plateauing of training loss or validation error for $\entropysgd$ in Fig.~\ref{fig:allcnn_loss}; this trait is shared across different networks and datasets in our experiments and is a direct consequence of the additional smoothness of the local entropy landscape.
+We see almost no plateauing of training loss or validation error for $\entropysgd$ in Fig.~\ref{fig:allcnn_loss}; this trait is shared across different networks and datasets in our experiments and is an indicator of the additional smoothness of the local entropy landscape coupled with a good scoping schedule for $\g$.
 
@@
         \includegraphics[width=\textwidth]{charlstm_valid.pdf}
-        \caption{\small $\charlstm$: Validation perplexity}
+        \caption{\small $\charlstm$: Validation, test perplexity}
         \label{fig:char_valid}
@@
 \end{figure}
+
+% {
+% \renewcommand{\arraystretch}{1.3}
+% \begin{table}[H]
+% \centering
+% \resizebox{0.8 \columnwidth}{!}
+% {%
+% \LARGE
+% \begin{tabular}{p{8cm} c c r}
+%     Model & Validation perplexity & Test perplexity & Time per effective epoch (s)\\  \hline
+%     $\ptblstm$ (SGD) & $81.43 \pm 0.2$ & $78.6 \pm 0.26$ & $0$\\
+%     $\ptblstm$ ($\entropysgd$) & $80.12 \pm 0.07$ & $77.66 \pm 0.17$ & $0$\\\hline
+%     $\charlstm$ (Adam) & $1.224 \pm 0.008$ & $1.226 \pm 0.01$ & $0$\\
+%     $\charlstm$ ($\entropyadam$) & $1.213 \pm 0.007$ & $1.217 \pm 0.005$ & $0$\\ \hline
+% \end{tabular}
+% }
+% \caption{\small $\entropysgd$ on RNNs}
+% \label{tab:rnns}
+% \end{table}
+% }
 
@@
 We connected this observation to the concept of local entropy which we used to bias the optimization towards flat regions that have low generalization error.
-Our experiments showed that this algorithm is applicable to large convolutional and recurrent deep networks used in practice.
+Our experiments showed that $\entropysgd$ is applicable to large convolutional and recurrent deep networks used in practice.
 
@@
 Concretely, the posterior expectation $\E\sqbrac{g(x)}$ is given by
-$$
+$
     \E\sqbrac{g(x)} \approx \f{\sum_{s=1}^t \eta_t\ g(x_t)}{\sum_{s=1}^t \eta_t};
-$$
+$
 which is the average computed by weighing each sample by the corresponding learning rate in~\eqref{eq:sgld_bayesian}.
@@
 \label{s:app:connection_variational}
-The local entropy based objective in $\entropysgd$ has similar motivations as variational inference.
 
-\paragraph{Comparison with SGLD:}
-A common way to train Bayesian neural networks is through MCMC algorithms to approximate the likelihood of data marginalized over the weights.
-Indeed, SGLD which forms the inner loop of $\entropysgd$ is a popular algorithm for variational inference~\citep{balan2015bayesian}.
-We run vanilla SGLD for two prototypical networks in Sec.~\ref{s:expt}, $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
+\newcommand{\kl}{\trm{KL}}
+The fundamental motivations of (stochastic) variational inference (SVI) and local entropy are similar: they both aim to generalize well by constructing a distribution on the weight space.
+In this section, we explore whether they are related and how one might reconcile the theoretical and algorithmic implications of local entropy objective with that of SVI.
+
+Let $\Xi$ denote the entire dataset, $z$ denote the weights of a deep neural network and $x$ be the parameters of a variational distribution $q_x(z)$.
+The Evidence Lower Bound (ELBO) can be then written as
+\beq{
+    \log p(\Xi) \geq \E_{z \sim q_x(z)}\ \sqbrac{\log p(\Xi\ |\ z)} - \kl \rbrac{q_x(z)\ ||\ p(z)};
+    \label{eq:elbo}
+}
+where $p(z)$ denotes a parameter-free prior on the weights and controls how well the posited posterior $q_x(z)$ fits the data through their $\kl$-divergence.
+Stochastic variational inference involves maximizing the right hand side of the above equation with respect to $x$ after choosing a suitable prior $p(z)$ and a family of distributions $q_x(z)$.
+These choices are typically dictated by the ease of sampling from $z \sim q_x(z)$, e.g.
+a mean-field model where $q_x(z)$ factorizes over $z$, and being able to compute the $\kl$-divergence term, e.g.
+a mixture of Gaussians.
+
+On the other hand, if we define the loss as the log-likelihood of data, viz. $f(z) := -\log p(\Xi | z)$, we can write the logarithm of the local entropy in Eqn.~\eqref{eq:local_entropy} as
+\aeq{
+    \log F(x, \g) &= \log \int_{z\ \in\ \reals^n}\ \exp \sqbrac{-f(z;\ \Xi) - \f{\g}{2}\ \norm{x-z}^2}\ dz \notag\\
+    &\geq \int_{z\ \in\ \reals^n}\ \sqbrac{\log p(\Xi\ |\ z) - \f{\g}{2}\ \norm{x-z}^2}\ dz;
+    \label{eq:logF}
+}
+by an application of Jensen's inequality.
+It is thus clear that Eqn.~\eqref{eq:elbo} and~\eqref{eq:logF} are very different in general and one cannot choose a prior or a variational family that makes them equivalent and interpret local entropy as ELBO.
+
+Eschewing rigor, formally, if we modify Eqn.~\eqref{eq:elbo} to allow the prior $p(z)$ to depend upon $x$, we can see that the two lower bounds above are equivalent iff $q_x(z)$ belongs to a ``flat variational family'', i.e.
+uniform distributions with $x$ as the mean and $p_x(z) \propto \exp \rbrac{-\f{\g}{2}\ \norm{x-z}^2}$.
+We emphasize that the distribution $p_x(z)$ depends on the parameters $x$ themselves and is thus, not really a prior.
+
+This ``moving prior'' is absent in variational inference and indeed a crucial feature of the local entropy objective.
+The gradient of local entropy in Eqn.~\eqref{eq:entropysgd_grad} clarifies this point:
+\aeqs{
+    \nabla F(x, \g) &= -\g\ \rbrac{x - \ag{z;\ \Xi}} = -\g\ \E_{z\ \sim r(z; x)}\ \sqbrac{z};
+}
+where the distribution $r(z; x)$ is given by
+$$
+    r(z;\ x) \propto\ p(\Xi\ |\ z)\ \exp \rbrac{-\f{\g}{2}\ \norm{x-z}^2};
+$$
+it thus contains a data likelihood term along with a prior that ``moves'' along with the current iterate $x$.
+
+Let us remark that methods in the deep learning literature that average the gradient in a local neighborhood of the weights $x$ through perturbations~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term) and are thus different from SVI and certainly, local entropy.
+
+\subsection{Comparison with SGLD}
+\label{ss:app:comparison_sgld}
+
+We use stochastic gradient Langevin dynamics (cf.
+Sec.~\ref{s:app:langevin}) to estimate the gradient of local entropy in Alg.~\ref{alg:entropysgd}.
+It is natural then, to ask the question whether vanilla SGLD performs as well as local entropy.
+%
+To this end, we compare the performance of SGLD on two prototypical networks: $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
+After a hyper-parameter search, we obtained a test error on $\lenet$ of $0.63 \pm 0.1\%$ after $300$ epochs and $9.89 \pm 0.11\%$ on $\allcnn$ after $500$ epochs.
+Even if one were to disregard the slow convergence of SGLD, its generalization error is much worse than our experimental results; we get $0.50 \pm 0.01 \%$ on $\lenet$ and $7.81 \pm 0.09 \%$ on $\allcnn$.
+For comparison, the authors in~\citet{chen2015bridging} report $0.71\%$ error on MNIST on a slightly larger network.
+Our results with local entropy on RNNs are much better than those reported in~\citet{gan2016scalable} for SGLD.
+On the PTB dataset, we obtain a test perplexity of $77.656 \pm 0.171$ vs.
+$94.03$ for the same model whereas we obtain a test perplexity of $1.213 \pm 0.007$ vs.
+$1.3375$ for $\charlstm$ on the War and Peace dataset.
+
+Training Bayesian deep networks with SGLD, or other more sophisticated algorithms such as SGHMC~\citep{chen2014stochastic,neal2011mcmc} is difficult and the lack of such results in the literature corroborates our experimental results.
+Roughly speaking, local entropy is so effective because it operates on a transformation of the energy landscape that exploits local entropic effects.
+Conventional MCMC techniques such as SGLD or Nose'-Hoover thermostats~\citep{ding2014bayesian} can only trade energy for entropy via the temperature parameter which does not allow direct use of the geometric information of the energy landscape and does not help with narrow minima.
 }

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 Our experiments show that while $\entropysgd$ trains significantly faster than SGD for recurrent networks, it gets relatively minor gains in terms of wall-clock time for CNNs.
-Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate the gradient in the teacher-student fashion~\citep{balan2015bayesian} is another avenue for extensions to this work.
+Estimating the gradient of local entropy cheaply with few SGLD iterations, or by using a smaller network to estimate it in a teacher-student framework~\citep{balan2015bayesian} is another avenue for extensions to this work.
 
@@
 
-This ``moving prior'' is absent in variational inference and indeed a crucial feature of the local entropy objective.
+This ``moving prior'' is absent in variational inference and indeed, a crucial feature of the local entropy objective.
 The gradient of local entropy in Eqn.~\eqref{eq:entropysgd_grad} clarifies this point:
@@
 
-Let us remark that methods in the deep learning literature that average the gradient in a local neighborhood of the weights $x$ through perturbations~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term) and are thus different from SVI and certainly, local entropy.
+Let us remark that methods in the deep learning literature that average the gradient through perturbations in the neighborhood of $x$~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term); such an averaging it thus different from local entropy.
 

### 2017-01-11 dnn/ploterrors.py

@@ -199,7 +199,7 @@ def plot_allcnn():
 
     ax.text(190, 0.1, r'$0.0353$', fontsize=fsz,
             verticalalignment='center', color='k')
-    ax.text(100, 0.05, r'$0.036$', fontsize=fsz,
+    ax.text(100, 0.05, r'$0.0336$', fontsize=fsz,
             verticalalignment='center', color='r')
 
     if opt['save']:

### 2017-01-11 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability.
+We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability, \pc{under certain assumptions}{}.
 Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.

### 2017-01-12 doc/chaudhari.choromanska.ea.iclr17.tex
pulling back updates from overleaf

@@
 We first provide a simple intuition for the concept of local entropy of an energy landscape.
-Consider a cartoon energy landscape in Fig.~\ref{fig:entropyfig} where the x-axis denotes the configuration space of the parameters.
+\pc{The discussion in this section builds upon the results of~\citet{baldassi2016unreasonable} and extends it for the case of continuous variables.}{} Consider a cartoon energy landscape in Fig.~\ref{fig:entropyfig} where the x-axis denotes the configuration space of the parameters.
 We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\trm{non-robust}}$.
@@
 \end{equation}
-which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$ if they both converge within $T$ parameter updates.
-\pc{However, from our experiments, $\entropysgd$ converges in roughly $L$ times fewer parameter updates than SGD, using this we can draw out the dependency on the number of Langevin iterations to get
+which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$ if they both converge in $T$ parameter updates.
+However, from our experiments, $\entropysgd$ converges in roughly $L$ times fewer parameter updates than SGD, using this we can draw out the dependency on the number of Langevin iterations to get
 $$
@@
 
+\pc{
 \begin{remark}
@@
 This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.~\ref{fig:lenet_hessian} has a large fraction of its eigenvalues near zero.
-Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} assumes global conditions on the smoothness of the loss function~\citep{hardt2015train}; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
+Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} by~\citet{hardt2015train} assumes global conditions on the smoothness of the loss function; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
+Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach of the current literature.
 \end{remark}
@@
 \textbf{Remark on the computational complexity:} Since $\entropysgd$ runs $L$ steps of SGLD before each parameter update, the effective number of passes over the dataset is $L$ times that of SGD or Adam for the same number of parameter updates.
-We therefore plot the error curves of $\entropysgd$ in Fig.~\ref{fig:mnist_test} and~\ref{fig:allcnn} against the ``effective number of epochs'', i.e.
+We therefore plot the error curves of $\entropysgd$ in Figs.~\ref{fig:mnist_test},~\ref{fig:allcnn}, and~\ref{fig:rnn} against the ``effective number of epochs'', i.e.
 by multiplying the abscissae by a factor of $L$.
@@
 In contrast to this observation about RNNs, all our experiments on CNNs used a momentum of $0.9$.
-\pc{In order to investigate this difficulty, we monitored the angle between the local entropy gradient and the vanilla SGD gradient during training.
-This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy gradient is highly uncorrelated with the SGD gradient in this case.}{}
+In order to investigate this difficulty, we monitored the angle between the local entropy gradient and the vanilla SGD gradient during training.
+This angle changes much more rapidly for RNNs than for CNNs which suggests a more rugged energy landscape for the former; local entropy gradient is highly uncorrelated with the SGD gradient in this case.
 
@@
 
-\todo{
 \section{Connection to variational inference}
@@
 Conventional MCMC techniques such as SGLD or Nose'-Hoover thermostats~\citep{ding2014bayesian} can only trade energy for entropy via the temperature parameter which does not allow direct use of the geometric information of the energy landscape and does not help with narrow minima.
-}
 

### 2017-01-12 doc/chaudhari.choromanska.ea.iclr17.tex
added back old paragraphs

@@
 \label{s:intro}
+
+\pc{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
+These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
+The original analysis was done in the case of binary weights, and was later generalized by~\citet{baldassi2016multilevel} to the case of weights with an arbitrary number of precision bits.
+Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.}{I moved this paragraph to the related work section (grey paragraph).
+The material and citations are preserved.
+Please feel free to change it.}
+
 This paper presents a new optimization tool for deep learning designed to exploit the local geometric properties of the objective function.
-Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+\pc{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.}{we can certainly add this sentence back, but I removed it for the following reasons, (i) their experiments are quite imprecise, (ii) the eigenspectrum is qualitatively different from ours, I suspect they are using forward-differencing to compute it / not averaging over the dataset etc., and (iii) we have done these experiments completely independently and it does not make sense to cite them before even presenting our own results} Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning seem to discover exactly that --- without being explicitly tailored to do so.
+\pc{For a more detailed recent
+analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.}{I removed this sentence}
 
@@
 
+\todo{
 Closer to the subject of this paper are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} who show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
@@
 The enabling technique in our case is using Markov chain Monte Carlo for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.
+}
 

### 2017-01-12 doc/rebuttal.txt
wip on rebuttal

--- 2017-01-03 doc/rebuttal.txt
@@
 [M10] Martens, James. "Deep learning via Hessian-free optimization." ICML (2010).
+
+-----------------------------------------------------------------------
+
+REBUTTAL
+
+We thank the reviewers and the area chair for their insightful feedback.
+We have incorporated all comments into our current draft.
+We first discuss the updates to the paper that address common questions raised by the reviewers.
+We have posted comments to each reviewer to address specific questions.
+
+Updates
+-------
+
+a) We have updated the experimental section of the paper with new experiments on MNIST, CIFAR-10 and two datasets on RNNs (PTB and char-LSTM).
+We also modified the algorithm to increase the value of the scoping parameter \gamma as training progresses instead of fixing it (Sec.
+4.3).
+As a result of this, we can now train all our networks with only the local entropy loss instead of treating it as a regularizer (Eqn.
+6).
+
+We obtain significant speed-ups with respect to earlier results due to scoping of the parameter \gamma and the wall-clock training time for all our networks with Entropy-SGD is now comparable to SGD/Adam.
+In fact, Entropy-SGD is twice as fast as SGD on our experiments on RNNs and also obtains a better generalization error (cf.
+Fig.
+6).
+The acceleration for CNNs on MNIST and CIFAR-10 is about 20%.
+
+b) Improved exposition of the algorithm in Sec. 4.2 and discussion of experiments in Sec. 5.3, 5.4 that includes intuition for hyper-parameter tuning.
+
+c) Appendix C discusses the possible connections to variational inference (this is the same as the discussion with the AC below).
+Sec.
+C.1 presents a comparison of local entropy and SGLD.
+We note here that our results using Entropy-SGD for both CNNs and RNNs are much better than SGLD is significantly smaller (~3x) wall-clock times.
+
+d)
+
+Response to the reviewers:
+-------------------------
+
+a) comment about speed
+We now plot the error curves in Figs.
+4, 5 and 6 against the "effective number of epochs", i.e., the number of epochs of Entropy-SGD is multipled by the number of Langevin iterations L (we set L=1 for SGD/Adam).
+Thus the x-axis is now a direct measure of the wall-clock time (agnostic to the underlying implementation and proportional to the number of back-props as suggested), which makes the comparison fair.
+
+b) smoothing of the original loss vs. local entropy
+We discuss this in detail in the related work in Sec.
+2 and Appendix C.
+While smoothing the original loss function using convolutions or averaging gradient over perturbations of weights reduces the ruggedness of the energy landscape, it does not help with sharp, narrow valleys.
+Local entropy introduces a measure that focuses on wide local minima in the energy landscape which is the primary reason for its efficacy.
+Smoothing the loss function can also, for instance, generate an artificial local minimum between two sharp valleys separated by a tall energ barrier which is detrimental to generalization.
+
+c) unrealistic eigenvalue assumption in Sec. 4.3
+We have clarified this point in Remark 4.
+Our analysis employs an assumption that the Hessian \nabla^2 f(x) does not have eigenvalues in the set [-2\gamma-c, c].
+This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.
+1 has a large fraction of its eigenvalues almost zero.
+Let us however remark that the bound on uniform stability in Thm.
+3 by Hardt et al.
+assumes global conditions on the smoothness of the loss function; one imagines that Eqn.
+9 remains qualitatively the same (with respect to the number of training iterations in particular) even if this assumption is violated to an extent before convergence happens.
+Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach of the current literature.
+
+------------------
+AnonReviewer4:
+
+>> how to reconcile "Entropy-SGD obtains better generalization but results in lower cross-entropy loss"
+The confusion is probably caused by our omission, it should read "lower cross-entropy loss on the training set" and is fixed now.
+This experimental insight suggests that there exist wide valleys that are deeper in the energy landscape that generalize well; (possibly) Entropy-SGD manages to find them while SGD gets stuck at a higher loss.
+
+>> the parameter \rho is used on CIFAR-10
+We have modified the algorithm to only train with local entropy (i.e., \rho = 0 always). Using scoping we can obtain better results both in terms of generalization error and wall-clock time.
+
+>> also discuss similarities to Hochreiter and Schmidhuber '97 [HS97]
+We have modified our discussion of HS97 in Sec.
+2.
+While the motivation is exactly the same, the similarities with their formulation are few and only superficial, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley, this is conceptually similar to their usage of Hessian.
+We agree with the reviewer that the elaborate analysis of generalization of HS97 using the Gibbs formalism is a promising direction.
+In our case, we benefit from similar technical results introduced for uniform stability [BE02].
+
+>> experiments are on a toy example
+Please note that our results on CIFAR-10 are close to the best result on non-augmented CIFAR-10 (6.55% error using ELU units) and certainly the best reported result for All-CNN-BN which is a medium-sized model with about 1.6 million weights.
+The largest model we experimented with is an LSTM on the PTB dataset with 66 million weights, please see the updates above for more details.
+
+[BE02] Olivier Bousquet and Andre Elisseeff. Stability and generalization. JMLR, 2002.
+
+------------------
+Csaba Szepesvari:
+
+>> discussion of Baldassi et. al. '16 in related work
+We have expanded the discussion of local entropy previously introduced by Baldassi et.
+al.
+in Sec.
+2.
+This paper builds upon their work and generalizes it to models with continuous variables, deep networks in particular.
+After posting the first version of this paper, we have collaborated with Baldassi et.
+al.
+on experiments with scoping techniques and they have co-authored this version of the paper.
+
+Please see the "updates to the paper" comment above for the response to the questions regarding (i) eigenvalue assumption, (ii) more thorough experimental results and, (iii) smoothing the loss vs.
+local entropy.
+
+------------------
+AnonReviewer1:
+
+>> Eqn. 8 should have f(x')
+Thanks, it is fixed now.
+
+>> experiments on RNNs
+Please see updates to the paper above (a), we have included results for word-level and character-level text prediction on PTB and char-LSTM with War and Peace.
+We not only obtain a better test perplexity than a competitive baseline with SGD, but also train in half as much wall-clock time.
+
+------------------
+AnonReviewer2:
+
+4. We set c from Thm. 3.12 of Hardt's paper to be 1 for clarity (c is the initial learning rate). Including it in the exponent does not change our analysis.
+
+------------------
+Areachair1 [Stefano should maybe edit and post that comment]:
+
+Re: You should make these connections clearer and more explicit
+We have discussed this in depth with sufficient technical details above.
+Using terms like "similar to motivations of Bayesian methods" and "using Bayesian methods in the inner loop" is generalizing to the point of absurdity.
+
+It is not accurate to term local entropy as "SVI with a moving prior" for a number of reasons.
+"SVI with a moving prior" has no theoretical basis.
+It is not connected to the log-likelihood of data, it is connected to local entropy instead.
+As we have discussed above, local entropy *resembles* SVI with a moving prior but only in an extremely non-rigorous and merely conceptual fashion.
+
+We have updated the experiments in our paper.
+We now plot the number of gradient evaluations on the x-axis and using scoping, we can in fact show that local entropy performs better in a smaller amount of time.
+As requested, we have also added a section comparing local-entropy with SGLD.
+
+Re: Exact Hessian-vector product is easy to compute
+We have discussed this above (see the comment regarding [P94]). Computing the "exact" HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch, which is expensive.

### 2017-01-12 doc/chaudhari.choromanska.ea.iclr17.tex
wip on rebuttal

@@
 
-\pc{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
+\todo{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
 These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
 The original analysis was done in the case of binary weights, and was later generalized by~\citet{baldassi2016multilevel} to the case of weights with an arbitrary number of precision bits.
-Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.}{I moved this paragraph to the related work section (grey paragraph).
-The material and citations are preserved.
-Please feel free to change it.}
+Here, we generalize this approach to the case of deep architectures and continuous weights, and develop a Langevin stochastic gradient descent algorithm based on the local entropy measure.}
 
 This paper presents a new optimization tool for deep learning designed to exploit the local geometric properties of the objective function.
-\pc{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.}{we can certainly add this sentence back, but I removed it for the following reasons, (i) their experiments are quite imprecise, (ii) the eigenspectrum is qualitatively different from ours, I suspect they are using forward-differencing to compute it / not averaging over the dataset etc., and (iii) we have done these experiments completely independently and it does not make sense to cite them before even presenting our own results} Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+\todo{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.} Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning seem to discover exactly that --- without being explicitly tailored to do so.
-\pc{For a more detailed recent
-analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.}{I removed this sentence}
+\todo{For a more detailed recent
+analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.}
 
@@
 It exploits the same phenomenon of wide valleys in the energy landscape but does so without incurring the hardware and communication complexity of replicated training or being limited to models where one can estimate local entropy using belief propagation.
-The enabling technique in our case is using Markov chain Monte Carlo for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.
+The enabling technique in our case is using Langevin dynamics for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.
 }
@@
 Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley) and also results in a more scalable algorithm.
-The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{ss:scoping}, this also provides significant computational advantages.
+The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{s:scoping}, this also provides significant computational advantages.
 The Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is however a very promising direction to understanding generalization in deep networks in the context of this paper.
@@
 \begin{equation}
-    \P(x';\ x, \g) \propto\ \exp \sqbrac{-\rbrac{\f{1}{m}\ \sum_{i=1}^m\ f\rbrac{x;\ \xi_{\ell_i}}} - \f{\g}{2}\ \norm{x-x'}_2^2}.
+    \P(x';\ x, \g) \propto\ \exp \sqbrac{-\rbrac{\f{1}{m}\ \sum_{i=1}^m\ f\rbrac{x';\ \xi_{\ell_i}}} - \f{\g}{2}\ \norm{x-x'}_2^2}.
     \label{eq:entropysgd_modified_gibbs}
@@
 
-\subsubsection{Scoping of $\gamma$}
-\label{ss:scoping}
+\subsection{Scoping of $\gamma$}
+\label{s:scoping}
 
@@
 The above analysis hinges upon an assumption that the Hessian $\nabla^2 f(x)$ does not have eigenvalues in the set $[-2\g-c, c]$ for a constant $c > 0$.
-This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.~\ref{fig:lenet_hessian} has a large fraction of its eigenvalues near zero.
+This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.~\ref{fig:lenet_hessian} has a large fraction of its eigenvalues almost zero.
 Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} by~\citet{hardt2015train} assumes global conditions on the smoothness of the loss function; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
@@
 To this end, we compare the performance of SGLD on two prototypical networks: $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
+We follow the experiments in~\citep{welling2011bayesian} and~\citep{chen2015bridging} and set the learning rate schedule to be $\eta/(1 + t)^b$ where the initial learning rate $\eta$ and $b$ are hyper-parameters.
+We make sure that other architectural aspects (dropout, batch-normalization) and regularization (weight decay) are consistent with the experiments in Sec.~\ref{s:expt}.
+
 After a hyper-parameter search, we obtained a test error on $\lenet$ of $0.63 \pm 0.1\%$ after $300$ epochs and $9.89 \pm 0.11\%$ on $\allcnn$ after $500$ epochs.
@@
 Roughly speaking, local entropy is so effective because it operates on a transformation of the energy landscape that exploits local entropic effects.
-Conventional MCMC techniques such as SGLD or Nose'-Hoover thermostats~\citep{ding2014bayesian} can only trade energy for entropy via the temperature parameter which does not allow direct use of the geometric information of the energy landscape and does not help with narrow minima.
+Conventional MCMC techniques such as SGLD or Nose'-Hoover thermostats~\citep{ding2014bayesian} can only trade energy for entropy via the temperature parameter which does not allow the direct use of the geometric information of the energy landscape and does not help with narrow minima.
 

### 2017-01-12 doc/rebuttal.txt
wip on rebuttal.txt

@@
 
+We now plot the error curves in Figs.
+4, 5 and 6 against the "effective number of epochs", i.e., the number of epochs of Entropy-SGD is multiplied by the number of Langevin iterations L (we set L=1 for SGD/Adam).
+Thus the x-axis is now a direct measure of the wall-clock time agnostic to the underlying implementation and proportional to the number of back-props as suggested, which makes the comparison fair.
+
 We obtain significant speed-ups with respect to earlier results due to scoping of the parameter \gamma and the wall-clock training time for all our networks with Entropy-SGD is now comparable to SGD/Adam.
@@
 
-d)
-
 Response to the reviewers:
@@
 
-a) comment about speed
-We now plot the error curves in Figs.
-4, 5 and 6 against the "effective number of epochs", i.e., the number of epochs of Entropy-SGD is multipled by the number of Langevin iterations L (we set L=1 for SGD/Adam).
-Thus the x-axis is now a direct measure of the wall-clock time (agnostic to the underlying implementation and proportional to the number of back-props as suggested), which makes the comparison fair.
-
-b) smoothing of the original loss vs. local entropy
+a) smoothing of the original loss vs. local entropy
 We discuss this in detail in the related work in Sec.
@@
 Local entropy introduces a measure that focuses on wide local minima in the energy landscape which is the primary reason for its efficacy.
-Smoothing the loss function can also, for instance, generate an artificial local minimum between two sharp valleys separated by a tall energ barrier which is detrimental to generalization.
+Smoothing the loss function can also, for instance, generate an artificial local minimum between two sharp valleys separated by a tall energy barrier which is detrimental to generalization.
 
-c) unrealistic eigenvalue assumption in Sec. 4.3
+b) unrealistic eigenvalue assumption in Sec. 4.3
 We have clarified this point in Remark 4.
@@
 >> how to reconcile "Entropy-SGD obtains better generalization but results in lower cross-entropy loss"
-The confusion is probably caused by our omission, it should read "lower cross-entropy loss on the training set" and is fixed now.
-This experimental insight suggests that there exist wide valleys that are deeper in the energy landscape that generalize well; (possibly) Entropy-SGD manages to find them while SGD gets stuck at a higher loss.
+The confusion is probably caused by our omission, it should read "lower cross-entropy loss on the training set" (fixed now).
+This experimental observation suggests that there exist wide valleys that are deeper in the energy landscape that generalize well; (possibly) Entropy-SGD manages to find them while SGD gets stuck at a higher loss.
 
->> the parameter \rho is used on CIFAR-10
+>> \rho = 0.01 on CIFAR-10 while \rho = 0 on MNIST
 We have modified the algorithm to only train with local entropy (i.e., \rho = 0 always). Using scoping we can obtain better results both in terms of generalization error and wall-clock time.
 
->> also discuss similarities to Hochreiter and Schmidhuber '97 [HS97]
+>> discuss similarities to Hochreiter and Schmidhuber '97 [HS97]
 We have modified our discussion of HS97 in Sec.
-2.
-While the motivation is exactly the same, the similarities with their formulation are few and only superficial, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley, this is conceptually similar to their usage of Hessian.
+2 to include this discussion: While the motivations are exactly the same, the similarities with their formulation are only conceptual in nature, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley which is similar to their usage of Hessian.
+The Gibbs variant to averaging in weight space (Eqn.
+33, pg.
+22 of [HS97]) is similar to the averaging in Eqn.
+7 of our paper.
+
 We agree with the reviewer that the elaborate analysis of generalization of HS97 using the Gibbs formalism is a promising direction.
@@
 >> experiments are on a toy example
-Please note that our results on CIFAR-10 are close to the best result on non-augmented CIFAR-10 (6.55% error using ELU units) and certainly the best reported result for All-CNN-BN which is a medium-sized model with about 1.6 million weights.
+Please note that our results on CIFAR-10 are close to the best result on non-augmented CIFAR-10 (6.55% error using ELU units) and certainly the best reported results for the popular All-CNN network which is a medium-sized model with about 1.6 million weights.
 The largest model we experimented with is an LSTM on the PTB dataset with 66 million weights, please see the updates above for more details.
@@
 in Sec.
-2.
+2 and included due citations in Sec.
+1 and 3 as well.
 This paper builds upon their work and generalizes it to models with continuous variables, deep networks in particular.
-After posting the first version of this paper, we have collaborated with Baldassi et.
+After posting the first version of this paper, we have been collaborating with Baldassi et.
 al.
-on experiments with scoping techniques and they have co-authored this version of the paper.
+for experiments on scoping techniques and they have co-authored this version of the paper.
 
@@
 
-4. We set c from Thm. 3.12 of Hardt's paper to be 1 for clarity (c is the initial learning rate). Including it in the exponent does not change our analysis.
-
-------------------
-Areachair1 [Stefano should maybe edit and post that comment]:
-
-Re: You should make these connections clearer and more explicit
-We have discussed this in depth with sufficient technical details above.
-Using terms like "similar to motivations of Bayesian methods" and "using Bayesian methods in the inner loop" is generalizing to the point of absurdity.
-
-It is not accurate to term local entropy as "SVI with a moving prior" for a number of reasons.
-"SVI with a moving prior" has no theoretical basis.
-It is not connected to the log-likelihood of data, it is connected to local entropy instead.
-As we have discussed above, local entropy *resembles* SVI with a moving prior but only in an extremely non-rigorous and merely conceptual fashion.
-
-We have updated the experiments in our paper.
-We now plot the number of gradient evaluations on the x-axis and using scoping, we can in fact show that local entropy performs better in a smaller amount of time.
-As requested, we have also added a section comparing local-entropy with SGLD.
-
-Re: Exact Hessian-vector product is easy to compute
-We have discussed this above (see the comment regarding [P94]). Computing the "exact" HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch, which is expensive.
+>> Thm. 3 does not have the constant c in Hardt et. al. '15
+We set the constant c of Hardt et.
+al., to 1 in our statement to avoid confusion with our constant, with is also called c (the learning \eta_t \leq 1/t instead of the original \eta_t \leq c/t).
+This does not change the result qualitatively.

### 2017-01-12 doc/chaudhari.choromanska.ea.iclr17.tex
wip on rebuttal.txt

@@
 \href{mailto:yann@cs.nyu.edu}{yann@cs.nyu.edu},
-\href{mailto:carlobaldassi@gmail.com}{carlobaldassi@gmail.com}}\\[0.03in]
+\href{mailto:carlo.baldassi@polito.it}{carlo.baldassi@polito.it},}\\[0.03in]
 {\footnotesize
@@
 
-Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
+\pc{Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
 As the authors discuss, the effect of hyper-parameters for the constraints is intricately tied together and they are difficult to choose even for small problems.
@@
 The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{s:scoping}, this also provides significant computational advantages.
-The Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is however a very promising direction to understanding generalization in deep networks in the context of this paper.
-%, it bears supercial similarities to the uniform stability analysis~\citep{haussler1997mutual} we employ.
-%Recent work by~\citet{keskar2016large} is relevant in the context of the main theme of our paper.
-The authors estimate the loss in a neighborhood of the weights to argue that small batch size in SGD (i.e., larger gradient noise) generalizes better than large mini-batches and also results in significantly flatter minima.
+The Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is however a very promising direction to understanding generalization in deep networks in the context of this paper.}{While the motivations are exactly the same, the similarities with their formulation are only conceptual in nature, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley which is similar to their usage of Hessian.
+The Gibbs variant to averaging in weight space (Eqn.
+33, pg.
+22 of [HS97]) is similar to the averaging in Eqn.
+7 of our paper.}
 

### 2017-01-12 doc/chaudhari.choromanska.ea.iclr17.tex
minor

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability, \pc{under certain assumptions}{}.
+We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability, under certain assumptions.
 Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
@@
 We first provide a simple intuition for the concept of local entropy of an energy landscape.
-\pc{The discussion in this section builds upon the results of~\citet{baldassi2016unreasonable} and extends it for the case of continuous variables.}{} Consider a cartoon energy landscape in Fig.~\ref{fig:entropyfig} where the x-axis denotes the configuration space of the parameters.
+The discussion in this section builds upon the results of~\citet{baldassi2016unreasonable} and extends it for the case of continuous variables.
+Consider a cartoon energy landscape in Fig.~\ref{fig:entropyfig} where the x-axis denotes the configuration space of the parameters.
 We have constructed two local minima: a shallower although wider one at $x_{\trm{robust}}$ and a very sharp global minimum at $x_{\trm{non-robust}}$.
@@
 \end{definition}
-\pc{The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
-Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.}{talk about small gamma and large gamma valleys} Note that $-F(x, \g)$ has a \emph{global} minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
+The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
+Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
+Note that $-F(x, \g)$ has a \emph{global} minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
 
@@
 
-\pc{
 \begin{remark}
@@
 \end{remark}
-}{}
 

### 2017-01-12 doc/rebuttal.txt
added a few comments to AC in the rebuttal

@@
 
+--------------------
+Area chair:
+
+>> You should make the connections clearer and more explicit
+We have updated the paper with the above discussion about SVI and experiments comparing with SGLD in Appendix C.
+As our previous comment shows, it'd be inaccurate to term local entropy as "SVI with a moving prior" since the latter is not rigorously defined, the resemblance is quite conceptual in nature.
+
+>> change x-axis for error curves
+We have updated the experiments in the paper and now plot all error curves against the "effective number of epochs", i.e., the number of epochs for Entropy-SGD is multiplied by L (we set L=1 for SGD/Adam).
+Note that this is proportional to the number of back-props.
+Please see the "updates to the paper" comment above for more details.
+
+>> Exact HVP is easy to compute
+The comment regarding [P94] above discusses this. Computing the exact HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch, which is expensive.
+
+---------------------------
 REBUTTAL

### 2017-01-13 doc/rebuttal.txt
just polishing, minor updates

@@
 >> You should make the connections clearer and more explicit
-We have updated the paper with the above discussion about SVI and experiments comparing with SGLD in Appendix C.
-As our previous comment shows, it'd be inaccurate to term local entropy as "SVI with a moving prior" since the latter is not rigorously defined, the resemblance is quite conceptual in nature.
+We have updated the paper to include the above discussion about SVI and added experiments comparing with SGLD in Appendix C.
+As our previous comment shows, it'd be inaccurate to term local entropy as "SVI with a moving prior" since the latter is not rigorously defined, their resemblance is quite conceptual in nature.
 
@@
 
+(maybe can skip this, it is not relevant in light of faster expts.)
 >> Exact HVP is easy to compute
@@
 ---------------------------
-REBUTTAL
+MAIN REBUTTAL
 
@@
 We have incorporated all comments into our current draft.
-We first discuss the updates to the paper that address common questions raised by the reviewers.
-We have posted comments to each reviewer to address specific questions.
+We first discuss the updates to the paper and address common questions raised by the reviewers.
+We have also posted individual comments to the reviewers to address specific questions.
 
@@
 a) We have updated the experimental section of the paper with new experiments on MNIST, CIFAR-10 and two datasets on RNNs (PTB and char-LSTM).
-We also modified the algorithm to increase the value of the scoping parameter \gamma as training progresses instead of fixing it (Sec.
+We also modified the algorithm to introduce a technique called "scoping", this increases the scoping parameter \gamma as training progresses instead of fixing it (Sec.
 4.3).
@@
 
-We now plot the error curves in Figs.
+For a fair comparison of the training time, we now plot the error curves in Figs.
 4, 5 and 6 against the "effective number of epochs", i.e., the number of epochs of Entropy-SGD is multiplied by the number of Langevin iterations L (we set L=1 for SGD/Adam).
-Thus the x-axis is now a direct measure of the wall-clock time agnostic to the underlying implementation and proportional to the number of back-props as suggested, which makes the comparison fair.
+Thus the x-axis is thus a direct measure of the wall-clock time agnostic to the underlying implementation and is proportional to the number of back-props as suggested.
 
-We obtain significant speed-ups with respect to earlier results due to scoping of the parameter \gamma and the wall-clock training time for all our networks with Entropy-SGD is now comparable to SGD/Adam.
+We obtain significant speed-ups with respect to our earlier results due to scoping and the wall-clock training time for all our networks with Entropy-SGD is now comparable to SGD/Adam.
 In fact, Entropy-SGD is twice as fast as SGD on our experiments on RNNs and also obtains a better generalization error (cf.
@@
 
-c) Appendix C discusses the possible connections to variational inference (this is the same as the discussion with the AC below).
+c) Appendix C discusses the possible connections to variational inference (this is the same material as the discussion with the AC below).
 Sec.
 C.1 presents a comparison of local entropy and SGLD.
-We note here that our results using Entropy-SGD for both CNNs and RNNs are much better than SGLD is significantly smaller (~3x) wall-clock times.
+We note here that our results using Entropy-SGD for both CNNs and RNNs are much better than SGLD in significantly smaller (~3x) wall-clock times.
 
@@
 While smoothing the original loss function using convolutions or averaging gradient over perturbations of weights reduces the ruggedness of the energy landscape, it does not help with sharp, narrow valleys.
-Local entropy introduces a measure that focuses on wide local minima in the energy landscape which is the primary reason for its efficacy.
-Smoothing the loss function can also, for instance, generate an artificial local minimum between two sharp valleys separated by a tall energy barrier which is detrimental to generalization.
+Local entropy introduces a measure that focuses on wide local minima in the energy landscape (cf.
+Fig.
+2 which has a "global" minimum at a wide valley); this is the primary reason for its efficacy.
+Smoothing the loss function can also, for instance, generate an artificial local minimum between two close by sharp valleys which is detrimental to generalization.
 
@@
 Our analysis employs an assumption that the Hessian \nabla^2 f(x) does not have eigenvalues in the set [-2\gamma-c, c].
-This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.
+This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian in Fig.
 1 has a large fraction of its eigenvalues almost zero.
-Let us however remark that the bound on uniform stability in Thm.
+Let us note though that Fig.
+1 is plotted at a local minimum, the eigenspectrum is less sparse in the beginning of training, but quite similar to Fig.
+1.
+
+The bound on uniform stability in Thm.
 3 by Hardt et al.
@@
 9 remains qualitatively the same (with respect to the number of training iterations in particular) even if this assumption is violated to an extent before convergence happens.
-Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach of the current literature.
+Obtaining a rigorous generalization bound without this assumption requires a dynamical analysis of SGD and seems out of reach currently.
 

### 2017-01-13 doc/chaudhari.choromanska.ea.iclr17.tex
just polishing, minor updates

@@
 A related work in this context is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker from the ensemble average.
-Such an ensemble training procedure enables improved generalization by ensuring that different workers land in the same wide valley and indeed, it turns out to be closely related to the replica theoretic analysis of~\citep{baldassi2016unreasonable}.
+Such an ensemble training procedure enables improved generalization by ensuring that different workers land in the same wide valley and indeed, it turns out to be closely related to the replica theoretic analysis of~\citet{baldassi2016unreasonable}.
 
@@
 
-\pc{Motivated by the same final goal, viz., flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
+Motivated by the same final goal, viz.
+flat local minima, the authors in~\citet{hochreiter1997flat} introduce hard constraints on the training loss and the width of local minima and show using the Gibbs formalism~\citep{haussler1997mutual} that this leads to improved generalization.
 As the authors discuss, the effect of hyper-parameters for the constraints is intricately tied together and they are difficult to choose even for small problems.
-Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley) and also results in a more scalable algorithm.
+Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley).
 The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{s:scoping}, this also provides significant computational advantages.
-The Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is however a very promising direction to understanding generalization in deep networks in the context of this paper.}{While the motivations are exactly the same, the similarities with their formulation are only conceptual in nature, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley which is similar to their usage of Hessian.
-The Gibbs variant to averaging in weight space (Eqn.
-33, pg.
-22 of [HS97]) is similar to the averaging in Eqn.
-7 of our paper.}
+Among other conceptual similarities with our work, let us note that local entropy in a flat valley is a direct measure of the width of the valley which is similar to their usage of Hessian while the Gibbs variant to averaging in weight space (Eqn.
+33 of~\citet{hochreiter1997flat}) is similar to our Eqn.~\eqref{eq:entropysgd_grad}.
+Indeed, Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is a very promising direction to understanding generalization in deep networks~\citep{zhang2016understanding}.
 
-Another related area is homotopy continuation methods that convolve the loss function to solve sequentially refined optimization problems~\citep{allgower2012numerical,mobahi2015link} or methods that perturb the weights or activations to average the gradient~\citep{gulcehre2016mollifying} with an aim to smooth the rugged energy landscape.
+Homotopy continuation methods convolve the loss function to solve sequentially refined optimization problems~\citep{allgower2012numerical,mobahi2015link} or methods that perturb the weights or activations to average the gradient~\citep{gulcehre2016mollifying} with an aim to smooth the rugged energy landscape.
 Such a smoothing is however different from local entropy.
 For instance, the latter places more weight on wide local minima even if they are much shallower than the global minimum (cf.
-Fig.~\ref{fig:entropyfig}) which cannot be obtained by smoothing.
-In fact, smoothing can introduce an artifical minimum between two sharp valleys connected by a high energy barrier which is detrimental to generalization.
+Fig.~\ref{fig:entropyfig}); this effect cannot be obtained by smoothing.
+In fact, smoothing can introduce an artificial minimum between two close by sharp valleys which is detrimental to generalization.
 Please refer to Sec.~\ref{s:app:connection_variational} for a more elaborate discussion.
 In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
-Local entropy exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}; our experiments on the eigenspectrum of the Hessian are a step to showing the existence of such wide minima in modern deep networks.
+\todo{Local entropy, on the other hand, exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}, shallow neural networks in particular.}
 
@@
 
-\begin{definition}[Local entropy]
+\begin{definition}[\textbf{Local entropy}]
 \label{def:local_entropy}
@@
 The parameter $\g$ is used to focus the modified Gibbs distribution upon a local neighborhood of $x$ and we call it a ``scope'' henceforth.
+
+\paragraph{Effect on the energy landscape:}
 Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
-Note that $-F(x, \g)$ has a \emph{global} minimum near $x_{\textrm{robust}}$ which is exactly what we want; we expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well.
+We expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well and indeed the negative local entropy in Fig.~\ref{fig:entropyfig} has a \textbf{global} minimum near $x_{\textrm{robust}}$.
+For low values of $\g$, the energy landscape is significantly smoother and still maintains our this desired characteristic.
+As $\g$ increases, the local entropy energy landscape gets closer to the original energy landscape and they become equivalent at $\g \to \infty$.
+On the other hand, for very small values of $\g$, the local entropy energy landscape is almost uniform.
 
@@
 For large values of $\g$, the SGLD updates happen in a small neighborhood of the current parameters $x$ while low values of $\g$ allow the inner SGLD to explore further away from $x$.
-A ``reverse-annealing'' of the scope $\g$, i.e.
-increasing $\g$ as training progresses, thus has the effect of exploring the energy landscape on progressively finer scales.
-This process is called ``scoping''~\citep{baldassi2016unreasonable} and use a simple exponential schedule given by
+In the context of the discussion in Sec.~\ref{s:local_entropy}, a ``reverse-annealing'' of the scope $\g$, i.e.
+increasing $\g$ as training progresses has the effect of exploring the energy landscape on progressively finer scales.
+We call this process ``scoping'' which is similar to that of~\citet{baldassi2016unreasonable} and use a simple exponential schedule given by
 $$
@@
 Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} by~\citet{hardt2015train} assumes global conditions on the smoothness of the loss function; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
-Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach of the current literature.
+Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach currently.
 \end{remark}
@@
 The results in Fig.~\ref{fig:mnistfc_test} and Fig.~\ref{fig:lenet_test} show that $\entropysgd$ obtains a comparable generalization error: $1.37 \pm 0.03 \%$ and $0.50 \pm 0.01 \%$, for $\mnistfc$ and $\lenet$ respectively.
+While $\entropysgd$ trains slightly faster in wall-clock time for $\lenet$; it is marginally slower for $\mnistfc$ which is a small network and trains in about two minutes anyway.
 
@@
 Note that this again results in only $25$ effective epochs, i.e.
-half as much wall-clock time as Adam.
+half as much wall-clock time as SGD or Adam.
 We obtain a validation perplexity of $1.213 \pm 0.007$ and a test perplexity of $1.217 \pm 0.005$ over $4$ independent runs which is better than the baseline.
-We plot the error curves in Fig.~\ref{fig:char_valid} for this experiment.
+Fig.~\ref{fig:char_valid} shows the error curves for this experiment.
 
@@
 For comparison, the authors in~\citet{chen2015bridging} report $0.71\%$ error on MNIST on a slightly larger network.
+
 Our results with local entropy on RNNs are much better than those reported in~\citet{gan2016scalable} for SGLD.
@@
 
-Training Bayesian deep networks with SGLD, or other more sophisticated algorithms such as SGHMC~\citep{chen2014stochastic,neal2011mcmc} is difficult and the lack of such results in the literature corroborates our experimental results.
-Roughly speaking, local entropy is so effective because it operates on a transformation of the energy landscape that exploits local entropic effects.
+Training deep networks with SGLD, or other more sophisticated MCMC algorithms such as SGHMC, SGNHT etc.~\citep{chen2014stochastic,neal2011mcmc} to errors similar to those of SGD is difficult, and the lack of such results in the literature corroborates our experimental experience.
+Roughly speaking, local entropy is so effective because it operates on a transformation of the energy landscape that exploits entropic effects.
 Conventional MCMC techniques such as SGLD or Nose'-Hoover thermostats~\citep{ding2014bayesian} can only trade energy for entropy via the temperature parameter which does not allow the direct use of the geometric information of the energy landscape and does not help with narrow minima.

### 2017-01-13 doc/chaudhari.choromanska.ea.iclr17.tex
final pass over the writing

@@
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-We show that the local entropy objective is smoother than the original landscape and show improved generalization bounds over SGD using uniform stability, under certain assumptions.
+We show that the new objective has a smoother energy landscape and show improved generalization bounds over SGD using uniform stability, under certain assumptions.
 Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
@@
 \item positive eigenvalues (right inset) have a long tail with the largest one being almost $40$,
-\item negative eigenvalues (left inset), which are directions of descent that the optimizer missed, have a much faster decay (the largest negative eigenvalue is only $-0.4$).
+\item negative eigenvalues (left inset), which are directions of descent that the optimizer missed, have a much faster decay (the largest negative eigenvalue is only $-0.45$).
 \end{enumerate}
 Interestingly, this trend is not unique to this particular network.
-Rather, its qualitative properties are shared across a variety of network architectures, sizes, datasets or optimization algorithms (refer to Sec.~\ref{s:expt} for more experiments).
-Local minima that generalize well and are discovered by gradient descent lie in ``wide valleys'' of the energy landscape, rather than in sharp isolated minima.
+Rather, its qualitative properties are shared across a variety of network architectures, network sizes, datasets or optimization algorithms (refer to Sec.~\ref{s:expt} for more experiments).
+Local minima that generalize well and are discovered by gradient descent lie in ``wide valleys'' of the energy landscape, rather than in sharp, isolated minima.
 For an intuitive understanding of this phenomenon, imagine a Bayesian prior concentrated at the minimizer of the expected loss, the marginal likelihood of wide valleys under this prior is much higher than narrow, sharp valleys even if the latter are close to the global minimum in training loss.
@@
 
-The existence of multiple, almost equivalent, local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
+The existence of multiple --- almost equivalent --- local minima in deep networks has been predicted using a wide variety of theoretical analyses and empirical observations, e.g., papers such as~\citet{spinglass2015,DBLP:conf/colt/ChoromanskaLA15,chaudhari2015trivializing} that build upon results from statistical physics as also others such as~\citet{haeffele2015global} and~\citet{janzamin2015beating} that obtain similar results for matrix and tensor factorization problems.
 Although assumptions in these works are somewhat unrealistic in the context of deep networks used in practice, similar results are also true for linear networks which afford a more thorough analytical treatment~\citep{DBLP:journals/corr/SaxeMG13}.
@@
 
-\todo{
 Closer to the subject of this paper are results by~\citet{baldassi2015subdominant,baldassi2016unreasonable,baldassi2016multilevel} who show that the energy landscape of shallow networks with discrete weights is characterized by an exponential number of isolated minima and few very dense regions with lots of local minima close to each other.
@@
 The enabling technique in our case is using Langevin dynamics for estimating the gradient of local entropy, which can be done efficiently even for large deep networks using mini-batch updates.
-}
 
@@
 As the authors discuss, the effect of hyper-parameters for the constraints is intricately tied together and they are difficult to choose even for small problems.
-Our local entropy based objective naturally balances the energetic term (training loss) and the entropic term (width of the valley).
-The ``scope'' $\g$ affords more control on the optimization process than hard constraints and as we discuss in Sec.~\ref{s:scoping}, this also provides significant computational advantages.
-Among other conceptual similarities with our work, let us note that local entropy in a flat valley is a direct measure of the width of the valley which is similar to their usage of Hessian while the Gibbs variant to averaging in weight space (Eqn.
+Our local entropy based objective instead naturally balances the energetic term (training loss) and the entropic term (width of the valley).
+The role of $\g$ is clear as a focusing parameter (cf.
+Sec.~\ref{s:scoping}) and effectively exploiting this provides significant computational advantages.
+Among other conceptual similarities with our work, let us note that local entropy in a flat valley is a direct measure of the width of the valley which is similar to their usage of Hessian, while the Gibbs variant to averaging in weight space (Eqn.
 33 of~\citet{hochreiter1997flat}) is similar to our Eqn.~\eqref{eq:entropysgd_grad}.
-Indeed, Gibbs formalism used in the analysis of~\citet{hochreiter1997flat} is a very promising direction to understanding generalization in deep networks~\citep{zhang2016understanding}.
+Indeed, Gibbs formalism used in their analysis is a very promising direction to understanding generalization in deep networks~\citep{zhang2016understanding}.
 
-Homotopy continuation methods convolve the loss function to solve sequentially refined optimization problems~\citep{allgower2012numerical,mobahi2015link} or methods that perturb the weights or activations to average the gradient~\citep{gulcehre2016mollifying} with an aim to smooth the rugged energy landscape.
-Such a smoothing is however different from local entropy.
+Homotopy continuation methods convolve the loss function to solve sequentially refined optimization problems~\citep{allgower2012numerical,mobahi2015link}, similarly, methods that perturb the weights or activations to average the gradient~\citep{gulcehre2016mollifying} do so with an aim to smooth the rugged energy landscape.
+Such smoothing is however very different from local entropy.
 For instance, the latter places more weight on wide local minima even if they are much shallower than the global minimum (cf.
@@
 In fact, smoothing can introduce an artificial minimum between two close by sharp valleys which is detrimental to generalization.
-Please refer to Sec.~\ref{s:app:connection_variational} for a more elaborate discussion.
+Please refer to Appendix~\ref{s:app:connection_variational} for a more elaborate discussion.
 In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
-\todo{Local entropy, on the other hand, exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}, shallow neural networks in particular.}
+Local entropy, on the other hand, exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}, shallow neural networks in particular.
 
@@
 \paragraph{Effect on the energy landscape:}
-Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for the original energy landscape.
-We expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well and indeed the negative local entropy in Fig.~\ref{fig:entropyfig} has a \textbf{global} minimum near $x_{\textrm{robust}}$.
-For low values of $\g$, the energy landscape is significantly smoother and still maintains our this desired characteristic.
+Fig.~\ref{fig:entropyfig} shows the negative local entropy $-F(x, \g)$ for two different values of $\g$.
+We expect $x_{\textrm{robust}}$ to be more robust than $x_{\textrm{non-robust}}$ to perturbations of data or parameters and thus generalize well and indeed, the negative local entropy in Fig.~\ref{fig:entropyfig} has a \textbf{global} minimum near $x_{\textrm{robust}}$.
+For low values of $\g$, the energy landscape is significantly smoother than the original landscape and still maintains our desired characteristic, viz.
+global minimum at a wide valley.
 As $\g$ increases, the local entropy energy landscape gets closer to the original energy landscape and they become equivalent at $\g \to \infty$.
@@
 for the Gibbs distribution in~\eqref{eq:gibbs_modified}.
-Minimizing classical entropy however does not differentiate between flat regions that have very high loss versus dense regions that lie deeper in the energy landscape.
-For instance, in Fig.~\ref{fig:entropyfig}, classical entropy is smallest in the neighborhood of $x_{\textrm{candidate}}$ which is a large region with very high loss on the training dataset and is unlikely to generalize well.
+Minimizing classical entropy however does not differentiate between flat regions that have very high loss versus flat regions that lie deeper in the energy landscape.
+For instance in Fig.~\ref{fig:entropyfig}, classical entropy is smallest in the neighborhood of $x_{\textrm{candidate}}$ which is a large region with very high loss on the training dataset and is unlikely to generalize well.
 \ignore{We could minimize a modified loss function of the form $f(x) + \lambda S(x, \g)$ whose gradient can be computed to be
@@
 
-We now present the $\entropysgd$ algorithm that is a variation of SGD motivated from local entropy.
-Simply speaking, it minimizes the negative local entropy from Sec.~\ref{s:local_entropy}.
-This section discusses how the gradient of local entropy can be computed via Langevin dynamics; the resulting algorithm has a strong flavor of ``SGD-inside-SGD'': the outer SGD updates the parameters, while an inner SGD estimates the gradient of local entropy.
+Simply speaking, our $\entropysgd$ algorithm minimizes the negative local entropy from Sec.~\ref{s:local_entropy}.
+This section discusses how the gradient of local entropy can be computed via Langevin dynamics.
+The reader will see that the resulting algorithm has a strong flavor of ``SGD-inside-SGD'': the outer SGD updates the parameters, while an inner SGD estimates the gradient of local entropy.
 
@@
 \end{equation}
-where the objective $f(x, \xi_k)$ is typically, a non-convex function in both the weights $x$ and the sample $\xi_k$. The $\entropysgd$ algorithm instead solves the problem
+where the objective $f(x, \xi_k)$ is typically, a non-convex function in both the weights $x$ and the samples $\xi_k$. The $\entropysgd$ algorithm instead solves the problem
 \begin{equation}
@@
 In this paper, we use stochastic gradient Langevin dynamics (SGLD)~\citep{welling2011bayesian} that is an MCMC algorithm for drawing samples from a Bayesian posterior and scales to large datasets using mini-batch updates.
-Please see Sec.~\ref{s:app:langevin} in the Supplement for a brief overview of SGLD.
-For our application, as lines 3-6 of Alg.~\ref{alg:entropysgd} show, SGLD resembles a few iterations of SGD with additive gradient noise.
+Please see Appendix~\ref{s:app:langevin} for a brief overview of SGLD.
+For our application, as lines 3-6 of Alg.~\ref{alg:entropysgd} show, SGLD resembles a few iterations of SGD with a forcing term $-\g(x-x')$ and additive gradient noise.
 
@@
 
-Let us discuss the algorithm in a little more detail.
+Let us now discuss a few implementation details.
 Although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
@@
 We found that annealing $\eta'$ (for instance, setting it to be the same as the outer learning rate $\eta$) is detrimental; indeed a small learning rate leads to poor estimates of local entropy towards the end of training where they are most needed.
-The parameter $\e$ in SGLD on Line 5 is the thermal noise and we fix this to $\e = 10^{-4}$.
+The parameter $\e$ in SGLD on Line 5 is the thermal noise and we fix this to $\e \in [10^{-4},10^{-3}]$.
 Having thus fixed $L, \e$ and $\eta'$, an effective heuristic to tune the remaining parameter $\g$ is to match the magnitude of the gradient of the local entropy term, viz.
@@
 
-The scoping parameter $\g$ is fixed in Alg.~\ref{alg:entropysgd}.
+The scope $\g$ is fixed in Alg.~\ref{alg:entropysgd}.
 For large values of $\g$, the SGLD updates happen in a small neighborhood of the current parameters $x$ while low values of $\g$ allow the inner SGLD to explore further away from $x$.
@@
 Scoping of $\g$ unfortunately interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
-In practice, we therefore scale the local entropy gradient by $\g$ before the update and modify the line to read
+In practice, we therefore scale down the local entropy gradient by $\g$ before the weight update and modify the line to read
 $$
@@
 We can use much larger learning rates than SGD in our experiments because the local entropy gradient is less noisier than the original back-propagated gradient.
+This also enables very fast progress in the beginning with a smooth landscape of a small $\g$.
 
@@
 \end{lemma}
-Please see Sec.~\ref{s:app:proofs} for the proof. The local entropy objective is thus smoother than the original objective.
+Please see Appendix~\ref{s:app:proofs} for the proof. The local entropy objective is thus smoother than the original objective.
 %
@@
 We denote an optimization algorithm, viz., SGD or $\entropysgd$ by $A(\Xi)$, it is a function of the dataset $\Xi$ and outputs the parameters $x^*$ upon termination.
-Stability of the algorithm~\citep{bousquet2002stability} is then a notion of how much its output differs in loss upon being presented with two datasets $\Xi$ and $\Xi'$ that differ in at most one sample,
+Stability of the algorithm~\citep{bousquet2002stability} is then a notion of how much its output differs in loss upon being presented with two datasets, $\Xi$ and $\Xi'$, that differ in at most one sample:
 $$
@@
 \end{equation}
-which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$ if they both converge in $T$ parameter updates.
-However, from our experiments, $\entropysgd$ converges in roughly $L$ times fewer parameter updates than SGD, using this we can draw out the dependency on the number of Langevin iterations to get
-$$
-    \e_{\ \entropysgd} \lessapprox L^{-\f{c\ \b}{\g}}\ \e_{\ \textrm{SGD}}.
-$$
-As an aside, it is easy to see from the proof of Lem.~\ref{lem:smoothness_reduction} that for a convex loss function $f(x)$, adding the local entropy term does not change the minimizer of the original problem.
+which shows that $\entropysgd$ generalizes better than SGD for all $T > \a$ if they both converge after $T$ passes over the samples.
+
+Let us note that while the number of passes over the dataset for $\entropysgd$ and SGD are similar for our experiments on CNNs, $\entropysgd$ makes only half as many passes as SGD for our experiments on RNNs.
+As an aside, it is easy to see from the proof of Lem.~\ref{lem:smoothness_reduction} that for a convex loss function $f(x)$, the local entropy objective does not change the minimizer of the original problem.
 
@@
 This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian at a local minimum in Fig.~\ref{fig:lenet_hessian} has a large fraction of its eigenvalues almost zero.
-Let us however remark that the bound on uniform stability in Thm.~\ref{thm:generalization_bound_nonconvex} by~\citet{hardt2015train} assumes global conditions on the smoothness of the loss function; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
+Let us however remark that the result in Thm.~\ref{thm:generalization_bound_nonconvex} by~\citet{hardt2015train} assumes global conditions on the smoothness of the loss function; one imagines that Eqn.~\ref{eq:stability_bound} remains qualitatively the same (with respect to $T$ in particular) even if this assumption is violated to an extent before convergence happens.
 Obtaining a rigorous generalization bound without this assumption would require a dynamical analysis of SGD and seems out of reach currently.
@@
 %
-The right tails of the eigenspectra are much longer than the left tails.
+Away from the origin, the right tails of the eigenspectra are much longer than the left tails.
 Indeed, as discussed in numerous places in literature~\citep{Bray2007,dauphin2014identifying,spinglass2015}, SGD finds low-index critical points, i.e., optimizers with few negative eigenvalues of the Hessian.
@@
 
-We consider two prototypical networks: the first is a fully-connected network with two hidden layers of $1024$ units each which we denote as ``$\mnistfc$'' and the second is a convolutional neural network with the same size as $\lenet$ but with batch-normalization~\citep{ioffe2015batch}; both with a dropout of probability $0.5$ after each layer.
-As a baseline, we train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs to obtain an average error of $1.39 \pm 0.03\%$ and $0.51 \pm 0.01 \%$ for $\mnistfc$ and $\lenet$ respectively over $5$ independent runs.
+We consider two prototypical networks: the first, called $\mnistfc$, is a fully-connected network with two hidden layers of $1024$ units each and the second is a convolutional neural network with the same size as $\lenet$ but with batch-normalization~\citep{ioffe2015batch}; both use a dropout of probability $0.5$ after each layer.
+As a baseline, we train for $100$ epochs with Adam and a learning rate of $10^{-3}$ that drops by a factor of $5$ after every $30$ epochs to obtain an average error of $1.39 \pm 0.03\%$ and $0.51 \pm 0.01 \%$ for $\mnistfc$ and $\lenet$ respectively, over $5$ independent runs.
 
@@
 by multiplying the abscissae by a factor of $L$.
-(define $L = 1$ for SGD or Adam).
-This is thus a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
+(we define $L = 1$ for SGD or Adam).
+Modulo the time required for actually updating the parameters (which are fewer for $\entropysgd$), this is a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
 
@@
 We train on CIFAR-10 without data augmentation after performing global contrast normalization and ZCA whitening~\citep{goodfellow2013maxout}.
-We consider the All-CNN-C network of~\citet{springenberg2014striving} with a batch normalization layer after each convolutional layer and all other architectural parameters such as filter-sizes and weight decay are kept the same.
+We consider the All-CNN-C network of~\citet{springenberg2014striving} with the only difference that a batch normalization layer is added after each convolutional layer; all other architecture and hyper-parameters are kept the same.
 We train for $200$ epochs with SGD and Nesterov's momentum during which the initial learning rate of $0.1$ decreases by a factor of $5$ after every $60$ epochs.
@@
 Roughly, wide valleys favored by $\entropysgd$ are located deeper in the landscape with a lower empirical loss than local minima discovered by SGD where it presumably gets stuck.
-Such an interpretation is contrast to theoretical models of deep networks (cf.\@ Sec.~\ref{s:prior_work}) which predicts multiple equivalent local minima with the same loss.
+Such an interpretation is in contrast to theoretical models of deep networks (cf.\@ Sec.~\ref{s:prior_work}) which predicts multiple equivalent local minima with the same loss.
 Our work suggests that geometric properties of the energy landscape are crucial to generalize well and provides algorithmic approaches to exploit them.
@@
 The fundamental motivations of (stochastic) variational inference (SVI) and local entropy are similar: they both aim to generalize well by constructing a distribution on the weight space.
-In this section, we explore whether they are related and how one might reconcile the theoretical and algorithmic implications of local entropy objective with that of SVI.
+In this section, we explore whether they are related and how one might reconcile the theoretical and algorithmic implications of the local entropy objective with that of SVI.
 
 Let $\Xi$ denote the entire dataset, $z$ denote the weights of a deep neural network and $x$ be the parameters of a variational distribution $q_x(z)$.
-The Evidence Lower Bound (ELBO) can be then written as
+The Evidence Lower Bound (ELBO) can be then be written as
 \beq{
@@
 }
-where $p(z)$ denotes a parameter-free prior on the weights and controls how well the posited posterior $q_x(z)$ fits the data through their $\kl$-divergence.
+where $p(z)$ denotes a parameter-free prior on the weights and controls, through their $\kl$-divergence, how well the posited posterior $q_x(z)$ fits the data.
 Stochastic variational inference involves maximizing the right hand side of the above equation with respect to $x$ after choosing a suitable prior $p(z)$ and a family of distributions $q_x(z)$.
-These choices are typically dictated by the ease of sampling from $z \sim q_x(z)$, e.g.
+These choices are typically dictated by the ease of sampling $z \sim q_x(z)$, e.g.
 a mean-field model where $q_x(z)$ factorizes over $z$, and being able to compute the $\kl$-divergence term, e.g.
@@
 \aeq{
-    \log F(x, \g) &= \log \int_{z\ \in\ \reals^n}\ \exp \sqbrac{-f(z;\ \Xi) - \f{\g}{2}\ \norm{x-z}^2}\ dz \notag\\
+    \log F(x, \g) &= \log \int_{z\ \in\ \reals^n}\ \exp \sqbrac{-f(z;\ \Xi) - \f{\g}{2}\ \norm{x-z}^2}\ dz, \notag\\
     &\geq \int_{z\ \in\ \reals^n}\ \sqbrac{\log p(\Xi\ |\ z) - \f{\g}{2}\ \norm{x-z}^2}\ dz;
@@
 by an application of Jensen's inequality.
-It is thus clear that Eqn.~\eqref{eq:elbo} and~\eqref{eq:logF} are very different in general and one cannot choose a prior or a variational family that makes them equivalent and interpret local entropy as ELBO.
+It is thus clear that Eqn.~\eqref{eq:elbo} and~\eqref{eq:logF} are very different in general and one cannot choose a prior, or a variational family, that makes them equivalent and interpret local entropy as ELBO.
 
@@
 uniform distributions with $x$ as the mean and $p_x(z) \propto \exp \rbrac{-\f{\g}{2}\ \norm{x-z}^2}$.
-We emphasize that the distribution $p_x(z)$ depends on the parameters $x$ themselves and is thus, not really a prior.
+We emphasize that the distribution $p_x(z)$ depends on the parameters $x$ themselves and is thus, not really a prior, or one that can be derived using the ELBO.
 
@@
 We use stochastic gradient Langevin dynamics (cf.
-Sec.~\ref{s:app:langevin}) to estimate the gradient of local entropy in Alg.~\ref{alg:entropysgd}.
+Appendix~\ref{s:app:langevin}) to estimate the gradient of local entropy in Alg.~\ref{alg:entropysgd}.
 It is natural then, to ask the question whether vanilla SGLD performs as well as local entropy.
@@
 To this end, we compare the performance of SGLD on two prototypical networks: $\lenet$ on MNIST and $\allcnn$ on CIFAR-10.
-We follow the experiments in~\citep{welling2011bayesian} and~\citep{chen2015bridging} and set the learning rate schedule to be $\eta/(1 + t)^b$ where the initial learning rate $\eta$ and $b$ are hyper-parameters.
+We follow the experiments in~\citet{welling2011bayesian} and~\citet{chen2015bridging} and set the learning rate schedule to be $\eta/(1 + t)^b$ where the initial learning rate $\eta$ and $b$ are hyper-parameters.
 We make sure that other architectural aspects (dropout, batch-normalization) and regularization (weight decay) are consistent with the experiments in Sec.~\ref{s:expt}.
@@
 After a hyper-parameter search, we obtained a test error on $\lenet$ of $0.63 \pm 0.1\%$ after $300$ epochs and $9.89 \pm 0.11\%$ on $\allcnn$ after $500$ epochs.
-Even if one were to disregard the slow convergence of SGLD, its generalization error is much worse than our experimental results; we get $0.50 \pm 0.01 \%$ on $\lenet$ and $7.81 \pm 0.09 \%$ on $\allcnn$.
+Even if one were to disregard the slow convergence of SGLD, its generalization error is much worse than our experimental results; we get $0.50 \pm 0.01 \%$ on $\lenet$ and $7.81 \pm 0.09 \%$ on $\allcnn$ with $\entropysgd$.
 For comparison, the authors in~\citet{chen2015bridging} report $0.71\%$ error on MNIST on a slightly larger network.

### 2017-01-13 doc/chaudhari.choromanska.ea.iclr17.tex
added a table

@@
 In Sec.~\ref{ss:expt:universality}, we discuss experiments that suggest that the characteristics of the energy landscape around local minimal accessible by SGD are universal to deep architectures.
-This observation motivates the $\entropysgd$ algorithm considered in this paper and in Sec.~\ref{ss:expt:mnist} and~\ref{ss:expt:cifar}, we present experimental results on two standard image classification datasets, viz., MNIST and CIFAR-10.
+We then present experimental results on two standard image classification datasets, viz.
+MNIST and CIFAR-10 and two datasets for text prediction, viz.
+PTB and the text of War and Peace.
+Table~\ref{tab:expts} summarizes the results of these experiments on deep networks.
 
@@
 
-% {
-% \renewcommand{\arraystretch}{1.3}
-% \begin{table}[H]
-% \centering
-% \resizebox{0.8 \columnwidth}{!}
-% {%
-% \LARGE
-% \begin{tabular}{p{8cm} c c r}
-%     Model & Validation perplexity & Test perplexity & Time per effective epoch (s)\\  \hline
-%     $\ptblstm$ (SGD) & $81.43 \pm 0.2$ & $78.6 \pm 0.26$ & $0$\\
-%     $\ptblstm$ ($\entropysgd$) & $80.12 \pm 0.07$ & $77.66 \pm 0.17$ & $0$\\\hline
-%     $\charlstm$ (Adam) & $1.224 \pm 0.008$ & $1.226 \pm 0.01$ & $0$\\
-%     $\charlstm$ ($\entropyadam$) & $1.213 \pm 0.007$ & $1.217 \pm 0.005$ & $0$\\ \hline
-% \end{tabular}
-% }
-% \caption{\small $\entropysgd$ on RNNs}
-% \label{tab:rnns}
-% \end{table}
-% }
+{
+\setlength{\heavyrulewidth}{1.5pt}
+\renewcommand{\arraystretch}{1.4}
+\begin{table}[H]
+\centering
+\resizebox{0.9 \columnwidth}{!}
+{
+\large
+\begin{tabular}{p{4cm} | c c | c c}
+\toprule
+    \rowcolor{gray!15} Model & \multicolumn{2}{c}{$\entropysgd$} & \multicolumn{2}{c}{SGD / Adam}\\
+\toprule
+    & Error (\%) / Perplexity & Epochs & Error (\%) / Perplexity & Epochs\\
+    \rowcolor{gray!15} $\mnistfc$ & $1.37 \pm 0.03$ & $120$ & $1.39 \pm 0.03$ & $66$\\
+    $\lenet$ & $0.5 \pm 0.01$ & $80$ & $0.51 \pm 0.01$ & $100$\\
+    \rowcolor{gray!15} $\allcnn$ & $7.81 \pm 0.09$ & $115$ & $7.71 \pm 0.19$ & $180$\\
+    $\ptblstm$ & $77.656 \pm 0.171$ & $25$ & $78.6 \pm 0.26$ & $55$\\
+    \rowcolor{gray!15} $\charlstm$ & $1.217 \pm 0.005$ & $25$ & $1.226 \pm 0.01$ & $40$\\
+\bottomrule
+\end{tabular}
+}
+\caption{\small Experimental results: $\entropysgd$ vs. SGD / Adam}
+\label{tab:expts}
+\end{table}
+}
 

### 2017-01-14 doc/rebuttal.txt
final version uploaded

@@
 
---------------------
-Area chair:
-
->> You should make the connections clearer and more explicit
-We have updated the paper to include the above discussion about SVI and added experiments comparing with SGLD in Appendix C.
-As our previous comment shows, it'd be inaccurate to term local entropy as "SVI with a moving prior" since the latter is not rigorously defined, their resemblance is quite conceptual in nature.
-
->> change x-axis for error curves
-We have updated the experiments in the paper and now plot all error curves against the "effective number of epochs", i.e., the number of epochs for Entropy-SGD is multiplied by L (we set L=1 for SGD/Adam).
-Note that this is proportional to the number of back-props.
-Please see the "updates to the paper" comment above for more details.
-
-(maybe can skip this, it is not relevant in light of faster expts.)
->> Exact HVP is easy to compute
-The comment regarding [P94] above discusses this. Computing the exact HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch, which is expensive.
-
 ---------------------------
@@
 Updates
--------
+=====
 
 a) We have updated the experimental section of the paper with new experiments on MNIST, CIFAR-10 and two datasets on RNNs (PTB and char-LSTM).
-We also modified the algorithm to introduce a technique called "scoping", this increases the scoping parameter \gamma as training progresses instead of fixing it (Sec.
+
+b) We have modified the algorithm to introduce a technique called "scoping".
+This increases the scope parameter \gamma as training progresses instead of fixing it and has the effect of exploring the parameter space in the beginning of training (Sec.
 4.3).
@@
 
-For a fair comparison of the training time, we now plot the error curves in Figs.
+c) For a fair comparison of the training time, we now plot the error curves in Figs.
 4, 5 and 6 against the "effective number of epochs", i.e., the number of epochs of Entropy-SGD is multiplied by the number of Langevin iterations L (we set L=1 for SGD/Adam).
-Thus the x-axis is thus a direct measure of the wall-clock time agnostic to the underlying implementation and is proportional to the number of back-props as suggested.
+Thus the x-axis is a direct measure of the wall-clock time agnostic to the underlying implementation and is proportional to the number of back-props as suggested.
 
 We obtain significant speed-ups with respect to our earlier results due to scoping and the wall-clock training time for all our networks with Entropy-SGD is now comparable to SGD/Adam.
-In fact, Entropy-SGD is twice as fast as SGD on our experiments on RNNs and also obtains a better generalization error (cf.
+In fact, Entropy-SGD is almost twice as fast as SGD on our experiments on RNNs and also obtains a better generalization error (cf.
 Fig.
@@
 
-b) Improved exposition of the algorithm in Sec. 4.2 and discussion of experiments in Sec. 5.3, 5.4 that includes intuition for hyper-parameter tuning.
+Table 1 (page 11) summarizes the experimental section of the paper.
 
-c) Appendix C discusses the possible connections to variational inference (this is the same material as the discussion with the AC below).
+d) Improved exposition of the algorithm in Sec.
+4.2 that includes intuition for hyper-parameter tuning.
+We have expanded the discussion of experiments in Sec.
+5.3, 5.4 to provide more details and insights that relate to the energy landscape of deep networks.
+
+e) Appendix C discusses the possible connections to variational inference (this is the same material as the discussion with the AC below).
 Sec.
-C.1 presents a comparison of local entropy and SGLD.
-We note here that our results using Entropy-SGD for both CNNs and RNNs are much better than SGLD in significantly smaller (~3x) wall-clock times.
+C.1 presents an experimental comparison of local entropy vs.
+SGLD.
+We note here that our results using Entropy-SGD for both CNNs and RNNs are much better than vanilla SGLD in significantly smaller (~3-5x) wall-clock times.
 
 Response to the reviewers:
--------------------------
+================
 
-a) smoothing of the original loss vs. local entropy
+>> smoothing of the original loss vs. local entropy
 We discuss this in detail in the related work in Sec.
 2 and Appendix C.
-While smoothing the original loss function using convolutions or averaging gradient over perturbations of weights reduces the ruggedness of the energy landscape, it does not help with sharp, narrow valleys.
+While smoothing the original loss function using convolutions or averaging the gradient over perturbations of weights reduces the ruggedness of the energy landscape, it does not help with sharp, narrow valleys.
 Local entropy introduces a measure that focuses on wide local minima in the energy landscape (cf.
@@
 2 which has a "global" minimum at a wide valley); this is the primary reason for its efficacy.
-Smoothing the loss function can also, for instance, generate an artificial local minimum between two close by sharp valleys which is detrimental to generalization.
+Smoothing the loss function can also, for instance, generate an artificial local minimum between two close by sharp valleys, which is detrimental to generalization.
 
-b) unrealistic eigenvalue assumption in Sec. 4.3
+>> unrealistic eigenvalue assumption in Sec. 4.3
 We have clarified this point in Remark 4.
-Our analysis employs an assumption that the Hessian \nabla^2 f(x) does not have eigenvalues in the set [-2\gamma-c, c].
+Our analysis employs an assumption that the Hessian \nabla^2 f(x) does not have eigenvalues in the set [-2\gamma-c, c] for some c > 0.
 This is admittedly unrealistic, for instance, the eigenspectrum of the Hessian in Fig.
@@
 Let us note though that Fig.
-1 is plotted at a local minimum, the eigenspectrum is less sparse in the beginning of training, but quite similar to Fig.
-1.
+1 is plotted at a local minimum, from our experiments, the eigenspectrum is less sparse in the beginning of training.
 
-The bound on uniform stability in Thm.
+We would like to remark that the bound on uniform stability in Thm.
 3 by Hardt et al.
 assumes global conditions on the smoothness of the loss function; one imagines that Eqn.
-9 remains qualitatively the same (with respect to the number of training iterations in particular) even if this assumption is violated to an extent before convergence happens.
+9 remains qualitatively the same (in particular, with respect to the number of training iterations) even if this assumption is violated to an extent before convergence happens.
 Obtaining a rigorous generalization bound without this assumption requires a dynamical analysis of SGD and seems out of reach currently.
+
+--------------------
+Area chair:
+
+>> You should make the connections clearer and more explicit
+We have updated the paper to include the above discussion about SVI and added experiments comparing with SGLD in Appendix C (also see the discussion in the last paragraph of Sec.
+2).
+As already remarked in our previous comment, characterizing local entropy as "SVI with a moving prior" would be inaccurate as their relation is just conceptual and not rigorous.
+
+>> change x-axis for error curves
+We have updated the experiments in the paper and now plot all error curves against the "effective number of epochs", i.e., the number of epochs for Entropy-SGD is multiplied by L (we set L=1 for SGD/Adam).
+Note that this is proportional to the number of back-props.
+Please see the "updates to the paper" comment above for more details.
+
+>> Exact HVP is easy to compute
+The comment regarding [P94] above discusses this. Computing the exact HVP requires averaging over the entire dataset, i.e., ~4x the complexity of one epoch.
 
@@
 
+Please see the "updates to the paper" comment above for the response to the questions regarding (i) the eigenvalue assumption and, (ii) thorough experimental results.
+
 >> how to reconcile "Entropy-SGD obtains better generalization but results in lower cross-entropy loss"
-The confusion is probably caused by our omission, it should read "lower cross-entropy loss on the training set" (fixed now).
-This experimental observation suggests that there exist wide valleys that are deeper in the energy landscape that generalize well; (possibly) Entropy-SGD manages to find them while SGD gets stuck at a higher loss.
+The confusion is probably caused by our omission, it should read "...
+lower cross-entropy loss on the training set" (fixed now).
+This experimental observation suggests that there exist wide valleys that are deeper in the energy landscape that also generalize well; Entropy-SGD manages to find them while SGD gets stuck at a higher loss.
 
->> \rho = 0.01 on CIFAR-10 while \rho = 0 on MNIST
-We have modified the algorithm to only train with local entropy (i.e., \rho = 0 always). Using scoping we can obtain better results both in terms of generalization error and wall-clock time.
+>> \rho = 0.01 on CIFAR-10 but \rho = 0 on MNIST
+We have modified the algorithm to only train with the local entropy objective, i.e., \rho = 0 always (cf.
+Eqn.
+6).
+Using scoping we can obtain better results on all our networks both in terms of generalization error and wall-clock time.
 
 >> discuss similarities to Hochreiter and Schmidhuber '97 [HS97]
-We have modified our discussion of HS97 in Sec.
-2 to include this discussion: While the motivations are exactly the same, the similarities with their formulation are only conceptual in nature, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley which is similar to their usage of Hessian.
-The Gibbs variant to averaging in weight space (Eqn.
+We have expanded our discussion of [HS97] in Sec.
+2 to include this discussion: While the motivations are exactly the same, similarities with their exact formulation are only conceptual in nature, e.g., in a flat minimum, the local entropy is a direct measure of the width of the valley which is similar to their usage of Hessian.
+The Gibbs variant to averaging in weight space in their analysis (Eqn.
 33, pg.
@@
 
-We agree with the reviewer that the elaborate analysis of generalization of HS97 using the Gibbs formalism is a promising direction.
-In our case, we benefit from similar technical results introduced for uniform stability [BE02].
+We agree with the reviewer that the elaborate analysis of generalization of [HS97] using the Gibbs formalism is a promising direction.
+In our case, we benefit from similar elaborate and technical results introduced for uniform stability [BE02].
 
 >> experiments are on a toy example
-Please note that our results on CIFAR-10 are close to the best result on non-augmented CIFAR-10 (6.55% error using ELU units) and certainly the best reported results for the popular All-CNN network which is a medium-sized model with about 1.6 million weights.
-The largest model we experimented with is an LSTM on the PTB dataset with 66 million weights, please see the updates above for more details.
+Please note that our results on CIFAR-10 are without any data augmentation, the best result for this is 6.55% error using ELU units by [CUH15].
+To our knowledge, our baseline on CIFAR-10 (7.71 \pm 0.19% error with SGD), is the best reported result for the popular All-CNN-C network [S14], which is a medium-sized model with about 1.6 million weights.
+The largest model we have experimented with is an LSTM on the PTB dataset with 66 million weights and we achieve better test perplexity than the original authors [ZSV14] in half as much wall-clock time.
 
+[HS97] Hochreiter, S. and Schmidhuber, J. (1997), Flat Minima.
 [BE02] Olivier Bousquet and Andre Elisseeff. Stability and generalization. JMLR, 2002.
+[CUH15] Clevert, Djork-Arné, Thomas Unterthiner, and Sepp Hochreiter. "Fast and accurate deep network learning by exponential linear units (elus)." arXiv:1511.07289 (2015).
+[S14] Springenberg, Jost Tobias, et al. "Striving for simplicity: The all convolutional net." arXiv:1412.6806 (2014).
+[ZSV14] Zaremba, Wojciech, Ilya Sutskever, and Oriol Vinyals. "Recurrent neural network regularization." arXiv:1409.2329 (2014).
 
@@
 Csaba Szepesvari:
+
+Please see the "updates to the paper" comment above for the response to the questions regarding (i) the eigenvalue assumption, (ii) thorough experimental results and, (iii) smoothing the loss vs.
+local entropy.
 
@@
 
-Please see the "updates to the paper" comment above for the response to the questions regarding (i) eigenvalue assumption, (ii) more thorough experimental results and, (iii) smoothing the loss vs.
-local entropy.
-
 ------------------
@@
 >> experiments on RNNs
-Please see updates to the paper above (a), we have included results for word-level and character-level text prediction on PTB and char-LSTM with War and Peace.
-We not only obtain a better test perplexity than a competitive baseline with SGD, but also train in half as much wall-clock time.
+Thanks for suggesting this.
+We have included results for word-level and character-level text prediction on PTB and char-LSTM with War and Peace.
+We not only obtain a better test perplexity than a competitive baseline with SGD, but also train in about half as much wall-clock time.
+
+Please see the "updates to the paper" comment above for more details.
 
@@
 We set the constant c of Hardt et.
-al., to 1 in our statement to avoid confusion with our constant, with is also called c (the learning \eta_t \leq 1/t instead of the original \eta_t \leq c/t).
+al., to 1 in our statement to avoid confusion with our constant, with is also called c (our statement has the learning rate \eta_t \leq 1/t instead of their original \eta_t \leq c/t).
 This does not change the result qualitatively.
+
+Please see the "updates to the paper" comment above for the response to the questions regarding (i) the eigenvalue assumption, (ii) thorough experimental results and, (iii) smoothing the original loss vs.
+local entropy.

### 2017-01-14 doc/chaudhari.choromanska.ea.iclr17.tex
final version uploaded

@@
 
-\linespread{1.02}
+\linespread{1.0}
 
@@
 \textbf{Christian Borgs$^{5}$, Jennifer Chayes$^{5}$, 
+Levent Sagun$^{2}$, 
 Riccardo Zecchina$^{4}$}\\[0.05in]
-$^{1}$ Computer Science, University of California, Los Angeles.\\
-$^{2}$ Courant Institute of Mathematical Sciences, New York University.\\
-$^{3}$ Facebook AI Research, New York.\\
-$^{4}$ Politecnico di Torino.\\
-$^{5}$ Microsoft Research New England, Cambridge.\\ [0.05in]
+$^{1}$ Computer Science, University of California, Los Angeles\\
+$^{2}$ Courant Institute of Mathematical Sciences, New York University\\
+$^{3}$ Facebook AI Research, New York\\
+$^{4}$ Dipartimento Scienza Applicata e Tecnologia, Politecnico di Torino\\
+$^{5}$ Microsoft Research New England, Cambridge\\ [0.05in]
 {\footnotesize
@@
 \href{mailto:jchayes@microsoft.com}{jchayes@microsoft.com},
+\href{mailto:sagun@cims.nyu.edu}{sagun@cims.nyu.edu},
 \href{mailto:riccardo.zecchina@polito.it}{riccardo.zecchina@polito.it}
@@
 % uncomment these for submission!
-% \renewcommand{\pc}[2]{#1}
-% \renewcommand{\ss}[2]{#1}
-% \renewcommand{\ac}[2]{#1}
-% \renewcommand{\todo}[1]{#1}
+\renewcommand{\pc}[2]{#1}
+\renewcommand{\ss}[2]{#1}
+\renewcommand{\ac}[2]{#1}
+\renewcommand{\todo}[1]{#1}
+\renewcommand{\fix}[1]{#1}
 
@@
 Local extrema with low generalization error have a large proportion of almost-zero eigenvalues in the Hessian with very few positive or negative eigenvalues.
-We leverage upon this observation to construct a local entropy based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
+We leverage upon this observation to construct a local-entropy-based objective function that favors well-generalizable solutions lying in large flat regions of the energy landscape, while avoiding poorly-generalizable solutions located in the sharp valleys.
 Conceptually, our algorithm resembles two nested loops of SGD where we use Langevin dynamics in the inner loop to compute the gradient of the local entropy before each update of the weights.
-We show that the new objective has a smoother energy landscape and show improved generalization bounds over SGD using uniform stability, under certain assumptions.
+We show that the new objective has a smoother energy landscape and show improved generalization over SGD using uniform stability, under certain assumptions.
 Our experiments on convolutional and recurrent neural networks demonstrate that Entropy-SGD compares favorably to state-of-the-art techniques in terms of generalization error and training time.
@@
 
-\todo{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
+%\todo{Recent works on shallow networks with discrete weights~\citep{baldassi2015subdominant,baldassi2016unreasonable} have shown that the geometrical structure of the energy landscape is characterized by the existence of an exponentially large number of local minima, a large number of narrow isolated minima, and few very dense regions of minima.
 These dense states have been discovered analytically by adopting an appropriate local entropy measure, and have been shown to be both algorithmically accessible and to have good generalization properties in both random and real-world settings.
@@
 This paper presents a new optimization tool for deep learning designed to exploit the local geometric properties of the objective function.
-\todo{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.} Consider the histogram in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} with about $47,000$ weights (cf.\@ Sec.~\ref{ss:expt:universality}).
+%\todo{The existence of wide minima in deep networks is also suggested by the spectral analysis of the Hessian~\citep{sagun2016singularity}.}
+Consider the histogram we obtained in Fig.~\ref{fig:lenet_hessian} showing the spectrum of the Hessian at an extremum discovered by Adam~\citep{kingma2014adam} for a convolutional neural network on MNIST~\citep{lecun1998gradient} ($\approx 47,000$ weights, cf.\@ Sec.~\ref{ss:expt:universality}).
 It is evident that:
@@
 \item positive eigenvalues (right inset) have a long tail with the largest one being almost $40$,
-\item negative eigenvalues (left inset), which are directions of descent that the optimizer missed, have a much faster decay (the largest negative eigenvalue is only $-0.45$).
+\item negative eigenvalues (left inset), which are directions of descent that the optimizer missed, have a much faster decay (the largest negative eigenvalue is only $-0.46$).
 \end{enumerate}
@@
 This suggests that wide valleys should result in better generalization and, indeed, standard optimization algorithms in deep learning seem to discover exactly that --- without being explicitly tailored to do so.
-\todo{For a more detailed recent
-analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.}
+For another recent analysis of the Hessian, see the parallel work of~\citet{sagun2016singularity}.
 
 \begin{figure}[tbh]
-% \vspace*{-0.1in}
+\vspace*{-0.1in}
 \centering
@@
 These dense local minima can be shown to generalize well for random input data; more importantly, they are also accessible by efficient algorithms using a novel measure called ``robust ensemble'' that amplifies the weight of such dense regions.
-The authors use belief propagation to estimate local entropy for simpler models considered there.
+The authors use belief propagation to estimate local entropy for simpler models such as committee machines considered there.
 A related work in this context is EASGD~\citep{zhang2015deep} which trains multiple deep networks in parallel and modulates the distance of each worker from the ensemble average.
@@
 Fig.~\ref{fig:entropyfig}); this effect cannot be obtained by smoothing.
-In fact, smoothing can introduce an artificial minimum between two close by sharp valleys which is detrimental to generalization.
-Please refer to Appendix~\ref{s:app:connection_variational} for a more elaborate discussion.
+In fact, smoothing can introduce an artificial minimum between two nearby sharp valleys which is detrimental to generalization.
 In order to be effective, continuation techniques also require that minimizers of successively smaller convolutions of the loss function lie close to each other~\citep{DBLP:conf/icml/HazanLS16}; it is not clear whether this is true for deep networks.
-Local entropy, on the other hand, exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}, shallow neural networks in particular.
+Local entropy, on the other hand, exploits wide minima which have been shown to exist in a variety of learning problems~\citep{monasson1995weight,cocco1996analytical}.
+Please refer to Appendix~\ref{s:app:connection_variational} for a more elaborate discussion as well as possible connections to stochastic variational inference~\citep{blei2016variational}.
 
@@
 Let us now discuss a few implementation details.
-Although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $9$.
-In our experiments, we use Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
+Although we have written Alg.~\ref{alg:entropysgd} in the classical SGD setup, we can easily modify it to include techniques such as momentum and gradient pre-conditioning~\citep{duchi2011adaptive} by changing lines $5$ and $7$.
+In our experiments, we have used SGD with Nesterov's momentum~\citep{sutskever2013importance} and Adam for outer and inner loops with similar qualitative results.
 %
-We use exponential averaging to estimate $\mu$ in the SGLD loop (Line 6) with $\a = 0.75$ so as to put more weight on the later samples, this is akin to a burn-in in standard SGLD.
+We use exponential averaging to estimate $\mu$ in the SGLD loop (line 6) with $\a = 0.75$ so as to put more weight on the later samples, this is akin to a burn-in in standard SGLD.
 
@@
 We found that annealing $\eta'$ (for instance, setting it to be the same as the outer learning rate $\eta$) is detrimental; indeed a small learning rate leads to poor estimates of local entropy towards the end of training where they are most needed.
-The parameter $\e$ in SGLD on Line 5 is the thermal noise and we fix this to $\e \in [10^{-4},10^{-3}]$.
+The parameter $\e$ in SGLD on line 5 is the thermal noise and we fix this to $\e \in [10^{-4},10^{-3}]$.
 Having thus fixed $L, \e$ and $\eta'$, an effective heuristic to tune the remaining parameter $\g$ is to match the magnitude of the gradient of the local entropy term, viz.
@@
 
-Scoping of $\g$ unfortunately interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on Line 7 of Alg.~\ref{alg:entropysgd}.
+Scoping of $\g$ unfortunately interferes with the learning rate annealing that is popular in deep learning, this is a direct consequence of the update step on line 7 of Alg.~\ref{alg:entropysgd}.
 In practice, we therefore scale down the local entropy gradient by $\g$ before the weight update and modify the line to read
@@
 Our general strategy during hyper-parameter tuning is to set the initial scope $\g_0$ to be very small, pick a large value of $\eta$ and set $\g_1$ to be such that the magnitude of the local entropy gradient is comparable to that of SGD.
-We can use much larger learning rates than SGD in our experiments because the local entropy gradient is less noisier than the original back-propagated gradient.
+We can use much larger learning rates than SGD in our experiments because the local entropy gradient is less noisy than the original back-propagated gradient.
 This also enables very fast progress in the beginning with a smooth landscape of a small $\g$.
@@
 A large proportion of eigenvalues of the Hessian are very close to zero or positive with a very small (relative) magnitude.
-This is suggests that the local geometry of the energy landscape is almost flat at local minima discovered by gradient descent.
+This suggests that the local geometry of the energy landscape is almost flat at local minima discovered by gradient descent.
 This agrees with theoretical results such as~\citet{baldassi2016local} where the authors predict that flat regions of the landscape generalize better.
@@
 \centering
-    \begin{subfigure}[t]{0.4\textwidth}
+    \begin{subfigure}[t]{0.44\textwidth}
         \includegraphics[width=1.08\textwidth]{mnistfc_hessian.pdf}
@@
     \hspace{0.2in}
-    \begin{subfigure}[t]{0.4 \textwidth}
+    \begin{subfigure}[t]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{charlstm_hessian.pdf}
@@
         \centering
-        \begin{subfigure}[b]{0.4 \textwidth}
+        \begin{subfigure}[b]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dm.pdf}
@@
         \hspace{0.2in}
-        \begin{subfigure}[b]{0.4 \textwidth}
+        \begin{subfigure}[b]{0.44 \textwidth}
         \includegraphics[width=\textwidth]{allcnn_dp.pdf}
@@
 For both these networks, we train $\entropysgd$ for $5$ epochs with $L = 20$ and reduce the dropout probability ($0.15$ for $\mnistfc$ and $0.25$ for $\lenet$).
-The learning rate of the SGLD updates is fixed to $\eta' = 0.1$ while the outer loop's learning rate is set to $\eta=1$ and drops by a factor of $10$ after the second epoch; we use Nesterov's momentum for both the loops.
-The thermal noise in SGLD updates (Line 5 of Alg.~\ref{alg:entropysgd}) is set to $10^{-3}$.
+The learning rate of the SGLD updates is fixed to $\eta' = 0.1$ while the outer loop's learning rate is set to $\eta=1$ and drops by a factor of $10$ after the second epoch; we use Nesterov's momentum for both loops.
+The thermal noise in SGLD updates (line 5 of Alg.~\ref{alg:entropysgd}) is set to $10^{-3}$.
 We use an exponentially increasing value of $\g$ for scoping, the initial value of the scope is set to $\g = 10^{-4}$ and this increases by a factor of $1.001$ after each parameter update.
@@
 (we define $L = 1$ for SGD or Adam).
-Modulo the time required for actually updating the parameters (which are fewer for $\entropysgd$), this is a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
+Modulo the time required for the actual parameter updates (which are fewer for $\entropysgd$), this is a direct measure of wall-clock time, agnostic to the underlying hardware and implementation.
 
@@
 \centering
-\resizebox{0.9 \columnwidth}{!}
+\resizebox{0.85 \columnwidth}{!}
 {
@@
 \toprule
-    \rowcolor{gray!15} Model & \multicolumn{2}{c}{$\entropysgd$} & \multicolumn{2}{c}{SGD / Adam}\\
+    \rowcolor{gray!15} Model & \multicolumn{2}{c}{Entropy-SGD} & \multicolumn{2}{c}{SGD / Adam}\\
 \toprule
-    & Error (\%) / Perplexity & Epochs & Error (\%) / Perplexity & Epochs\\
+    & Error (\%) / Perplexity & Epochs & Error (\%) / Perplexity & Epochs\\[0.05in]
     \rowcolor{gray!15} $\mnistfc$ & $1.37 \pm 0.03$ & $120$ & $1.39 \pm 0.03$ & $66$\\
     $\lenet$ & $0.5 \pm 0.01$ & $80$ & $0.51 \pm 0.01$ & $100$\\
-    \rowcolor{gray!15} $\allcnn$ & $7.81 \pm 0.09$ & $115$ & $7.71 \pm 0.19$ & $180$\\
+    \rowcolor{gray!15} $\allcnn$ & $7.81 \pm 0.09$ & $160$ & $7.71 \pm 0.19$ & $180$\\
     $\ptblstm$ & $77.656 \pm 0.171$ & $25$ & $78.6 \pm 0.26$ & $55$\\
@@
 
-Local entropy in Def.~\eqref{def:local_entropy} is an expectation over the entire configuration space $x \in \reals^n$ and is hard to compute, we can however approximate its gradient using Markov chain Monte-Carlo (MCMC) techniques.
+Local entropy in Def.~\eqref{def:local_entropy} is an expectation over the entire configuration space $x \in \reals^n$ and is hard to compute; we can however approximate its gradient using Markov chain Monte-Carlo (MCMC) techniques.
 In this section, we briefly review stochastic gradient Langevin dynamics~\citep{welling2011bayesian} that is an MCMC algorithm designed to draw samples from a Bayesian posterior and scales to large datasets using mini-batch updates.
@@
 \end{equation}
-where $\e_t \sim \trm{N}(0, 1)$ is Gaussian noise and $\eta$ is the learning rate.
+where $\e_t \sim \trm{N}(0, \e^2)$ is Gaussian noise and $\eta$ is the learning rate.
 In this form, Langevin dynamics faces two major hurdles for applications to large datasets.
@@
 
-Let us note that there is a variety of increasingly sophisticated MCMC algorithms applicable to our problem, e.g., Hamiltonian Monte Carlo (SGHMC) by~\citet{chen2014stochastic} based on volume preserving flows in the ``parameter-momentum'' space, stochastic annealing thermostats (Santa) by~\citet{chen2015bridging} etc.
+Let us note that there is a variety of increasingly sophisticated MCMC algorithms applicable to our problem, e.g., Stochastic Gradient Hamiltonian Monte Carlo (SGHMC) by~\citet{chen2014stochastic} based on volume preserving flows in the ``parameter-momentum'' space, stochastic annealing thermostats (Santa) by~\citet{chen2015bridging} etc.
 We can also employ these techniques, although we use SGLD for ease of implementation.
@@
 
-Let us remark that methods in the deep learning literature that average the gradient through perturbations in the neighborhood of $x$~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term); such an averaging it thus different from local entropy.
+Let us remark that methods in the deep learning literature that average the gradient through perturbations in the neighborhood of $x$~\citep{mobahi2016training} or noisy activation functions~\citep{gulcehre2016mollifying} can be interpreted as computing the data likelihood in ELBO (without the KL-term); such an averaging is thus different from local entropy.
 

### 2017-01-14 hessian/hessianfig.py

@@ -31,22 +31,22 @@ plt.rc('figure', titlesize=fontsize)
 def load_data(n):
     e = []
     if n == 'lenet':
-        for i in glob.glob('../../results/hessian/lenet/*.npy'):
+        for i in glob.glob('../results/nov_expts/hessian/lenet/*.npy'):
             e.append(np.real(np.load(i)))
     elif n == 'mnistfc':
-        for i in glob.glob('../../results/hessian/mnistfc2/*.npy'):
+        for i in glob.glob('../results/nov_expts/hessian/mnistfc2/*.npy'):
             e.append(np.load(i)[:,0]*128)
     elif n == 'charlstm':
-        for i in glob.glob('../../results/hessian/char_lstm/*.npy'):
+        for i in glob.glob('../results/nov_expts/hessian/char_lstm/*.npy'):
             e.append(np.load(i)[:,0]*32)
     elif n == 'cifarfc':
-        for i in glob.glob('../../results/hessian/cifarfc/*.npy'):
+        for i in glob.glob('../results/nov_expts/hessian/cifarfc/*.npy'):
             e.append(np.load(i)[:,0])
     elif n == 'lenet_nonconvg':
-        for i in glob.glob('../../results/hessian/lenet_nonconvg/*.npy'):
+        for i in glob.glob('../results/nov_expts/hessian/lenet_nonconvg/*.npy'):
             e.append(np.real(np.load(i)))
     elif n == 'cifarconv':
-        for i in glob.glob('../../results/hessian/allcnn_fisher/*.optim_state.t7'):
+        for i in glob.glob('../results/nov_expts/hessian/allcnn_fisher/*.optim_state.t7'):
             t = thf.load(i)
             e.append( (t['v'] - t['m']**2)*256)
     else:

### 2017-01-25 hessian/landscapefig.py
minor, landscapefig.py

@@ -1,0 +1,61 @@
+import numpy as np
+import numpy.random as npr
+from scipy.signal import correlate, convolve, gaussian, exponential
+import matplotlib.pyplot as plt
+from scipy.misc import logsumexp
+import seaborn as sns
+import os, sys, argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('-s', '--save', action='store_true')
+opt = vars(parser.parse_args())
+
+plt.ion()
+sns.set_style('white')
+
+npr.seed(42)
+fontsize = 24
+
+plt.rc('font', size=fontsize)
+plt.rc('axes', titlesize=fontsize)
+plt.rc('axes', labelsize=fontsize)
+plt.rc('xtick', labelsize=fontsize)
+plt.rc('ytick', labelsize=fontsize)
+plt.rc('legend', fontsize=fontsize)
+plt.rc('figure', titlesize=fontsize)
+
+def plot_clusters():
+    ds = [1, 0.5, 0.1, 0.01]
+    n = 200
+    x = np.random.rand(n,2)*2 - 1
+    x = x[np.linalg.norm(x,axis=1) < 1]
+
+    xc1 = np.random.multivariate_normal([0.5,0.5], np.eye(2)*0.01, n/4)
+    xc2 = np.random.multivariate_normal([-0.25,0.75], np.eye(2)*0.005, n/4)
+    xc3 = np.random.multivariate_normal([0,-0.5], np.eye(2)*0.005, n/4)
+    xc = np.vstack((xc1,xc2,xc3))
+
+    for i in xrange(len(ds)):
+        _x = x[::int(1/ds[i])]
+        _xc = xc[::int(1/ds[i])]
+
+        fig = plt.figure(1, figsize=(6,6))
+        plt.clf()
+        ax = fig.add_subplot(111)
+        plt.plot(_x[:,0], _x[:,1], '.', marker='.', ms=20)
+        plt.plot(_xc[:,0], _xc[:,1], 'r.', marker='.', ms=20)
+        plt.plot(np.cos(np.linspace(0,2*np.pi,100)),np.sin(np.linspace(0,2*np.pi,100)),
+            'k-', lw=1)
+
+        plt.xlim([-1,1])
+        plt.ylim([-1,1])
+        plt.xticks([])
+        plt.yticks([])
+        plt.axis('off')
+
+        if opt['save']:
[…]

### 2017-01-25 doc/arxivize.tex
uploaded to arxiv

@@
     Title={Entropy-SGD: Biasing Gradient Descent Into Wide Valleysre},
-    Author={Pratik Chaudhari, Anna Choromanska, Stefano Soatto, Yann LeCun},
-    Keywords={deep learning, gradient descent, optimization, machine learning, neural networks}
+    Author={Pratik Chaudhari, Anna Choromanska, Stefano Soatto, Yann LeCun,Carlo Baldassi,Christian Borgs,Jennifer Chayes,Levent Sagun,Riccardo Zecchina},
+    Keywords={deep learning, gradient descent, optimization, machine learning, neural networks,statistical physics, local entropy}
   }

### 2017-02-11 doc/rebuttal.txt

@@
 local entropy.
+
+--------------------------------------------------
+
+>>(short questions)
+
+Congratulations to the authors for the high points on the updated PDF.
+I hope I am not very late in asking two short questions.
+Will you be opensourcing the code for the method & how does the method relate to the discussions in some other recent ICLR papers
+
+[1] Understanding deep learning requires rethinking generalization
+[2] Snapshot Ensembles: Train 1, Get M for Free
+[3] On Large-Batch Training for Deep Learning: Generalization Gap and Sharp Minima
+
+They are only tangentially related I know but I was curious how the motivation of the method fits into the generalization arguments of [1,3] and whether the ensembling trick of [2] can be used in conjunction with Entropy-SGD to reap benefits of both.
+
+-----------
+We plan to open-source the code and examples.
+
+We could speculate on how our method relates to the experiments of [1] but we would rather leave this to further investigation.
+
+[2] is quite different from the replica theoretic motivations of Entropy-SGD, which aims to create one good model.
+While it is possible to use ensembles of multiple networks with local scoping, the benefits might be offset by significantly longer test time.
+
+The metric in [3] is a proxy for local entropy (max instead of our integral).
+Entropy-SGD is a specialized algorithm to lead to flat minima, achieving the same effect using parameters such as batch-size, dropout, batch-normalization, data augmentation etc.
+is quite hard.
